---
title: "Real Data Analysis"
author: "Leiwen Gao"
date: "8/30/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=FALSE, warning = FALSE, message = FALSE)
```

Here are reproducible codes for results in Section 4 (Analysis of SEER dataset with four cancers).

### 1. Load California map and packages
```{r}
##### Libraries for sptial information in map, map projection and map ploting
library(maps)
#Import California map with counties
ca.county = map("county","california", fill=TRUE, plot=FALSE)
library(mapproj)
library(spdep)
library(maptools)
library(classInt)
library(RColorBrewer)
library(ggpubr)

##### Libraries for matrix, distribution and math computation
library(mvtnorm)
library(matrixStats)
library(blockmatrix)
library(Matrix)
library(invgamma)
library(mnormt)
library(MASS)
library(Hmisc) #rcorr
library(gtools) #permutation

##### Libraries for JAGS
library(rjags)
library(R2jags)

##### Draw plots
library(ggplot2)

##### Data management
library(tidyr)

##### Import and export documents
library(knitr)
library(readr)

##### Bayesian inference
library(LaplacesDemon)

##### Model selection
library(bridgesampling)

```


### 2. Dataset exploration and visualization
#### 2.1 Import dataset (county-level) extracted from SEER
Cancers: lung, esophageal, larynx and colorectal

Outcome: Standard incidence ratios (SIR$_{id} = Y_{id}/E_{id}$) in the years from 2012 to 2016.

$Y_{id}$ - observed counts of incidence for each cancer $d$ in each county $i$

$E_{id}$ - expected number of cases for each cancer $d$ in each county $i$

Covariates (percentage): adult cigarette smoking rates (smoking), older than 65 years old (age), percentages of unemployed residents (unemployed)

```{r}
#Import covariates
covariates <- read.csv("data/covariates.csv")
race <- read.csv("data/race.csv")
sex <- read.csv("data/sex.csv")
insurance <- read.csv("data/insurance.csv")
smoking <- read.csv("data/smoking.csv")
smoking$smoking <- as.numeric(substr(smoking$Cigarette.Smoking.Rate., 1,4))

#Import age-adjusted incidence rates for 4 cancers in California
rate_lung = rate_5y %>% filter(Site.code == "Lung and Bronchus")
rate_esophagus = rate_5y %>% filter(Site.code == "Esophagus")
rate_larynx = rate_5y %>% filter(Site.code == "Larynx")
rate_colrect = rate_5y %>% filter(Site.code == "Colon and Rectum")

################## Data construction ###########

## Data

# Covariates in percentage
county_attribute = covariates[substr(covariates$State_county,1,2) == "CA",]
county_attribute$state = extract_numeric(county_attribute$State_county)
county_attribute1 = data.frame(county_attribute[order(county_attribute$state),])
county_attribute1$V_Persons_age_18_ACS_2012_2016 = as.numeric(county_attribute1$V_Persons_age_18_ACS_2012_2016)/100
county_attribute1$V_Persons_age_65_ACS_2012_2016 = as.numeric(county_attribute1$V_Persons_age_65_ACS_2012_2016)/100
county_attribute1$VHighschooleducationACS2012201 = as.numeric(county_attribute1$VHighschooleducationACS2012201)/100
county_attribute1$VFamiliesbelowpovertyACS201220 = as.numeric(county_attribute1$VFamiliesbelowpovertyACS201220)/100
county_attribute1$V_Unemployed_ACS_2012_2016 = as.numeric(county_attribute1$V_Unemployed_ACS_2012_2016)/100

race1 = race[substr(race$State_county,1,2) == "CA"&race$Race_recode_White_Black_Other=="Black",]
sex1 = sex[substr(sex$State_county,1,2) == "CA"&sex$Sex=="Male",]
insurance1 = insurance[substr(insurance$State_county,1,2) == "CA"&insurance$Insurance_Recode_2007=="Uninsured",]

rate_lung1 = cbind(rate_lung, smoking$smoking, county_attribute1[,2:6], race1$Row_Percent, sex1$Row_Percent,insurance1$Row_Percent)
colnames(rate_lung1) = c("county", "O_count", "E_count", "standard_ratio", "site", "smoking", "young","old", "highschool", "poverty", "unemployed", "black", "male", "uninsured")
rate_esophagus1 = cbind(rate_esophagus, smoking$smoking, county_attribute1[,2:6], race1$Row_Percent, sex1$Row_Percent,insurance1$Row_Percent)
colnames(rate_esophagus1) = c("county", "O_count", "E_count", "standard_ratio", "site", "smoking", "young","old", "highschool", "poverty", "unemployed", "black", "male", "uninsured")
rate_larynx1 = cbind(rate_larynx, smoking$smoking, county_attribute1[,2:6], race1$Row_Percent, sex1$Row_Percent,insurance1$Row_Percent)
colnames(rate_larynx1) = c("county", "O_count", "E_count", "standard_ratio", "site", "smoking", "young","old", "highschool", "poverty", "unemployed", "black", "male", "uninsured")
rate_colrect1 = cbind(rate_colrect, smoking$smoking, county_attribute1[,2:6], race1$Row_Percent, sex1$Row_Percent,insurance1$Row_Percent)
colnames(rate_colrect1) = c("county", "O_count", "E_count", "standard_ratio", "site", "smoking", "young","old", "highschool", "poverty", "unemployed", "black", "male", "uninsured")
```

#### 2.2 Combine cancer data with the California map for each county

To plot the cancer data, we need to combine the data with the map polygon
```{r}
#County information
county.ID <- sapply(strsplit(ca.county$names, ","), function(x) x[2])
ca.poly = map2SpatialPolygons(ca.county, IDs=county.ID)
ca.poly$rate_lung = rate_lung$standard_ratio
ca.poly$rate_esophagus = rate_esophagus$standard_ratio
ca.poly$rate_larynx = rate_larynx$standard_ratio
ca.poly$rate_colrect = rate_colrect$standard_ratio
ca.poly$smoking = smoking$smoking

ca.coords = coordinates(ca.poly)

```


#### 2.3 Plots for raw incidence data and important covariates

Figure 3: Maps of standardized incidence ratios (SIR) for lung, esophageal, larynx and colocrectal cancer in California, 2012 - 2016
```{r, fig.height=7, fig.width=10}
##################### Plot SIR in areal map ##################
brks_fit_lung = quantile(ca.poly$rate_lung, c(0, 0.2, 0.4, 0.6, 0.8, 1))
brks_fit_esophagus = quantile(ca.poly$rate_esophagus, c(0, 0.2, 0.4, 0.6, 0.8, 1))
brks_fit_larynx = quantile(ca.poly$rate_larynx, c(0, 0.2, 0.4, 0.6, 0.8, 1))
brks_fit_colrect = quantile(ca.poly$rate_colrect, c(0, 0.2, 0.4, 0.6, 0.8, 1))


color.pallete = rev(brewer.pal(5,"RdBu"))
class.rate_lung = classIntervals(var=ca.poly$rate_lung, n=5, style="fixed", 
                                 fixedBreaks=brks_fit_lung, dataPrecision=4)
class.rate_esophagus = classIntervals(var=ca.poly$rate_esophagus, n=5, style="fixed", 
                                      fixedBreaks=brks_fit_esophagus, dataPrecision=4)
class.rate_larynx = classIntervals(var=ca.poly$rate_larynx, n=5, style="fixed", 
                                   fixedBreaks=brks_fit_larynx, dataPrecision=4)
class.rate_colrect = classIntervals(var=ca.poly$rate_colrect, n=5, style="fixed", 
                                    fixedBreaks=brks_fit_colrect, dataPrecision=4)
color.code.rate_lung = findColours(class.rate_lung, color.pallete)
color.code.rate_esophagus = findColours(class.rate_esophagus, color.pallete)
color.code.rate_larynx = findColours(class.rate_larynx, color.pallete)
color.code.rate_colrect = findColours(class.rate_colrect, color.pallete)


par(mfrow=c(2,2), oma = c(0,0,4,0) + 0.1, mar = c(0,0,2,0) + 0.1)

plot(ca.poly, col = color.code.rate_lung, main = "Lung cancer")

plot(ca.poly, col = color.code.rate_esophagus, main = "Esophageal cancer")

plot(ca.poly, col = color.code.rate_larynx, main = "larynx cancer")

plot(ca.poly, col = color.code.rate_colrect, main = "colorectal cancer")
leg.txt = c("0-20%", "20%-40%", "40%-60%", "60%-80%", "80%-100%")

par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0), new = TRUE)
plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
legend("center", legend=leg.txt, xpd = TRUE, cex=1.5, bty="n", horiz = FALSE, 
       fill = color.pallete)

```

Figure 11: Maps of county-level covariates: adult cigarette smoking rates (left) and unemployed residents (right).
```{r, fig.height=5, fig.width=10}
ca.poly$unemployed = rate_lung1$unemployed
color.pallete = rev(brewer.pal(5,"RdBu"))
brks_fit_smoking = quantile(ca.poly$smoking, c(0, 0.2, 0.4, 0.6, 0.8, 1))
class.smoking = classIntervals(var=ca.poly$smoking, n=5, style="fixed",
                                  fixedBreaks=brks_fit_smoking, dataPrecision=4)
color.code.smoking = findColours(class.smoking, color.pallete)


brks_fit_unemployed = quantile(ca.poly$unemployed, c(0, 0.2, 0.4, 0.6, 0.8, 1))
class.unemployed = classIntervals(var=ca.poly$unemployed, n=5, style="fixed",
                                  fixedBreaks=brks_fit_unemployed, dataPrecision=4)
color.code.unemployed = findColours(class.unemployed, color.pallete)

par(mfrow=c(1,2), oma = c(0,0,4,0) + 0.1, mar = c(0,0,1,0) + 0.1)

plot(ca.poly, col = color.code.smoking)
leg.txt = c("6.70 - 11.04", "11.04 - 12.60", "12.60 - 15.30", "15.30 - 18.10", "18.10 - 25.50")
legend("bottomleft", title="Smoking (%)", legend=leg.txt, xpd = TRUE, cex=1.25, bty="n", horiz = FALSE,
       fill = color.pallete)

plot(ca.poly, col = color.code.unemployed)
leg.txt = c("3.92 - 7.00", "7.00 - 8.68", "8.68 - 10.28", "10.28 - 11.72", "11.72 - 17.09")
legend("bottomleft", title="Unemployed (%)", legend=leg.txt, xpd = TRUE, cex=1.25, bty="n", horiz = FALSE,
       fill = color.pallete)


```

#### 2.4 Create adjacency matrix for spatial information in directed graph (ordered regions)

The matrix ``Adj'' is the adjacency matrix for counties in California. We reorder the counties starting from northwest to southeast for directed graph. Then we put all the islands (counties with no parent neighbors) at the beginning. There are $58$ counties in California.

```{r}
ca.neighbors = poly2nb(ca.poly)
n=length(ca.neighbors)

# original adjacency matrix
Adj=sapply(ca.neighbors,function(x,n) {v=rep(0,n);v[x]=1;v},n)
colnames(Adj)=county.ID
ca.coord = coordinates(ca.poly)
ca.latrange=round(quantile(ca.coord[,2],c(0.25,0.75)))
ca.albersproj=mapproject(ca.coord[,1],ca.coord[,2],projection = "albers",param=ca.latrange)

# reorder counties from northwest to southeast
perm=order(ca.albersproj$x-ca.albersproj$y)
Adj_new=Adj[perm,perm]

n=nrow(Adj_new)
ni=rowSums(Adj_new)
maxn=max(ni)
neimat=matrix(0,n,maxn)
neighbors=lapply(1:n,function(x) which(Adj_new[x,]==1))
dneighbors=sapply(2:n,function(i) intersect(neighbors[[i]],1:(i-1)))
dni=sapply(dneighbors,length)
original_perm = 1:58
index2=c(1,which(dni==0)+1)

# final permutation for counties: put all the islands (counties with no parent neighbors) at the beginning
final_perm=c(original_perm[perm][index2],
             original_perm[perm][-index2])
colnames(Adj)[final_perm]
Minc = Adj[final_perm,final_perm]
n=nrow(Minc)
ni=rowSums(Minc)
maxn=max(ni)
neimat=matrix(0,n,maxn)
neighbors=lapply(1:n,function(x) which(Minc[x,]==1))
#directed or parent neighbors
dneighbors=sapply(2:n,function(i) intersect(neighbors[[i]],1:(i-1)))
dni=sapply(dneighbors,length)
nmax=max(dni)
cni=cumsum(dni)
dneimat=sapply(dneighbors, function(nei,nmax,n) c(nei,rep(n+1,nmax+1-length(nei))),nmax,n)
udnei=unlist(dneighbors)
ni_wo = sapply(neighbors,length)
cni_wo = cumsum(ni_wo)
udnei_wo = unlist(neighbors)
cn = c(0, cni)
ns = dni

region = seq(1:n)
cn = c(0, cni)
ns = dni
index = list()
for(i in 1:(n-2)){
  index[[i]] = region[-(udnei[(cn[i+1] + 1):(cn[i+1] + ns[i+1])])]
}
index1 = unlist(index)
```

#### 2.5 Pairs of neighboring counties

Here are a total of 139 pairs of neighboring counties in California.
```{r}
# Neighboring counties information
neighborvec0 = NULL
neighbor_list0 = NULL
neighbor_name = NULL
for(i in 1:(n-1)){
  for(j in (i+1):n){
    if(Adj[i,j] == 1){
      neighborvec0 = c(neighborvec0, paste(i, ",", j, sep=""))
      neighbor_list0 = rbind(neighbor_list0, c(i,j))
      neighbor_name = c(neighbor_name, paste(colnames(Adj)[i], ", ", colnames(Adj)[j], sep=""))
    }
  }
}
neighbor_name
```

#### 2.6 Exploratory analysis for spatial dependence and association across cancers
##### 2.6.1  spatial dependence: Moran's I
To explore the spatial association for each disease, we calculate Moran’s I based upon rth order neighbors for each cancer and plot the areal correlogram.

Figure 4: Moran’s I of rth order neighbors for lung, esophageal, larynx and colorectum cancer
```{r, fig.height=5, fig.width=7}
# distance matrix
projmat=cbind(ca.albersproj$x,ca.albersproj$y)
dmat=as.matrix(dist(projmat))
# Moran's I function
moranI <- function(y, A){
  n = length(y)
  nom_sum = 0
  den_sum = 0
  for(i in 1:n){
    den_sum = den_sum + (y[i]-mean(y))^2
    for(j in 1:n){
      nom_sum = nom_sum + A[i,j]*(y[i]-mean(y))*(y[j]-mean(y))
    }
  }
  return(n*nom_sum/sum(A)/den_sum)
}

lung_moran = c()
esophagus_moran = c()
larynx_moran = c()
colrect_moran = c()
for(lag in 1:11){
  A_1 <- as.matrix((dmat <= 0.01*lag & dmat > 0.01*(lag-1))*1)
  diag(A_1) = 0
  lung_moran[lag] <- as.numeric(moranI(rate_lung$standard_ratio, A_1))
  esophagus_moran[lag] <- as.numeric(moranI(rate_esophagus$standard_ratio, A_1))
  larynx_moran[lag] <- as.numeric(moranI(rate_larynx$standard_ratio, A_1))
  colrect_moran[lag] <- as.numeric(moranI(rate_colrect$standard_ratio, A_1))
}

moran_value = c(lung_moran, esophagus_moran, larynx_moran, colrect_moran)
cancer = c(rep("Lung", 11), rep("Esophageal", 11), rep("Larynx", 11), rep("Colorectum", 11))
df = data.frame(cancer)
df$moran_value = moran_value
df$r = rep(1:11, 4)
df$cancer = factor(df$cancer, levels = c("Lung", "Esophageal", "Larynx", "Colorectum"))

ggplot(df, aes(r, moran_value)) + geom_point() +
  ylab("Moran's I") + facet_wrap(~cancer) + theme_bw() +
  scale_x_continuous(breaks = 1:11) +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        plot.title = element_text(size=11))
```

##### 2.5.2  Association across cancers: Pearson's correlation matrix
```{r}
cancer_ratio = cbind(rate_lung$standard_ratio, rate_esophagus$standard_ratio, rate_larynx$standard_ratio, rate_colrect$standard_ratio)
colnames(cancer_ratio) = c("Lung and Bronchus", "Esophageal", "Larynx", "Colon and Rectum")
cor_matrix <- rcorr(cancer_ratio)
cor_matrix
```

##### 2.5.3 Figure 5: Boundaries (in red) selected as the first 70 pairs with largest differences for lung, esophageal, larynx and colorectal cancer.
```{r}
bd_diffl1 =  NULL
bd_diffl2 =  NULL
bd_diffl3 =  NULL
bd_diffl4 =  NULL
for(i in 1:(n-1)){
  for(j in (i+1):n){
    if(Adj[i,j] == 1){
      bd_diff1 = abs(rate_lung$standard_ratio[i] - rate_lung$standard_ratio[j])
      names(bd_diff1) = paste(colnames(Adj)[i], ", ", colnames(Adj)[j], sep="")
      bd_diffl1 = c(bd_diffl1, bd_diff1)
      
      bd_diff2 = abs(rate_esophagus$standard_ratio[i] - rate_esophagus$standard_ratio[j])
      names(bd_diff2) = paste(colnames(Adj)[i], ", ", colnames(Adj)[j], sep="")
      bd_diffl2 = c(bd_diffl2, bd_diff2)
      
      bd_diff3 = abs(rate_larynx$standard_ratio[i] - rate_larynx$standard_ratio[j])
      names(bd_diff3) = paste(colnames(Adj)[i], ", ", colnames(Adj)[j], sep="")
      bd_diffl3 = c(bd_diffl3, bd_diff3)
      
      bd_diff4 = abs(rate_colrect$standard_ratio[i] - rate_colrect$standard_ratio[j])
      names(bd_diff4) = paste(colnames(Adj)[i], ", ", colnames(Adj)[j], sep="")
      bd_diffl4 = c(bd_diffl4, bd_diff4)
    }
  }
}
raw_th1 <- sort(bd_diffl1, decreasing = T)[70]
raw_th2 <- sort(bd_diffl2, decreasing = T)[70]
raw_th3 <- sort(bd_diffl3, decreasing = T)[70]
raw_th4 <- sort(bd_diffl4, decreasing = T)[70]

raw_diff1 = as.numeric(bd_diffl1 >= raw_th1)
raw_diff2 = as.numeric(bd_diffl2 >= raw_th2)
raw_diff3 = as.numeric(bd_diffl3 >= raw_th3)
raw_diff4 = as.numeric(bd_diffl4 >= raw_th4)

neighbor_list_diff1 <- neighbor_list0[raw_diff1 == 1, ]
neighbor_list_diff2 <- neighbor_list0[raw_diff2 == 1, ]
neighbor_list_diff3 <- neighbor_list0[raw_diff3 == 1, ]
neighbor_list_diff4 <- neighbor_list0[raw_diff4 == 1, ]
```

```{r}
# Libraries for map with boundaires
if (!require(gpclib)) install.packages("gpclib", type="source")
gpclibPermit()
p = list()
p_est=list()

library(plyr)
ca.poly$lung_SIR = rate_lung$standard_ratio
ca.poly$esophagus_SIR = rate_esophagus$standard_ratio
ca.poly$larynx_SIR = rate_larynx$standard_ratio
ca.poly$colrect_SIR = rate_colrect$standard_ratio

ca.poly@data$id <- rownames(ca.poly@data)
ca.poly.f <- fortify(ca.poly, region = "id")
ca.poly.df <- join(ca.poly.f, ca.poly@data, by = "id")

# Correct boundary data for the map
path=list()
for(i in 1: nrow(neighbor_list0)){
  r1 <- ca.poly.df[ca.poly.df$id %in% neighbor_list0[i,1],1:2]
  r2 <- ca.poly.df[ca.poly.df$id %in% neighbor_list0[i,2],1:2]
  edges <- generics::intersect(r1, r2)
  path[[i]] = edges
}
path[[2]][nrow(path[[2]])+1,] <- path[[2]][1,]
path[[2]] <- path[[2]][-1,]
path[[11]][nrow(path[[11]])+1,] <- path[[11]][1,]
path[[11]] <- path[[11]][-1,]
path[[19]][nrow(path[[19]])+1,] <- path[[19]][1,]
path[[19]] <- path[[19]][-1,]
path[[25]][nrow(path[[25]])+1,] <- path[[25]][1,]
path[[25]] <- path[[25]][-1,]
path[[27]][nrow(path[[27]])+1,] <- path[[27]][1,]
path[[27]] <- path[[27]][-1,]
path[[31]][nrow(path[[31]])+1,] <- path[[31]][1,]
path[[31]] <- path[[31]][-1,]
path[[39]][nrow(path[[39]])+1,] <- path[[39]][1,]
path[[39]] <- path[[39]][-1,]
path[[43]][nrow(path[[43]])+1,] <- path[[43]][1,]
path[[43]] <- path[[43]][-1,]
path[[46]][nrow(path[[46]])+1,] <- path[[46]][1,]
path[[46]] <- path[[46]][-1,]
path[[64]][nrow(path[[64]])+1,] <- path[[64]][1,]
path[[64]] <- path[[64]][-1,]
path[[75]][nrow(path[[75]])+1,] <- path[[75]][1,]
path[[75]] <- path[[75]][-1,]
path[[76]][nrow(path[[76]])+1,] <- path[[76]][1,]
path[[76]] <- path[[76]][-1,]
path[[81]][nrow(path[[81]])+1,] <- path[[81]][1,]
path[[81]] <- path[[81]][-1,]
path[[100]][nrow(path[[100]])+1,] <- path[[100]][1,]
path[[100]] <- path[[100]][-1,]
path[[125]][nrow(path[[125]])+1,] <- path[[125]][1,]
path[[125]] <- path[[125]][-1,]
path[[130]][nrow(path[[130]])+1,] <- path[[130]][1,]
path[[130]] <- path[[130]][-1,]
path[[133]][nrow(path[[133]])+1,] <- path[[133]][1,]
path[[133]] <- path[[133]][-1,]
```

```{r}
breaks1 = quantile(ca.poly$lung_SIR, c(0.2, 0.4, 0.6, 0.8))
breaks2 = quantile(ca.poly$esophagus_SIR, c(0.2, 0.4, 0.6, 0.8))
breaks3 = quantile(ca.poly$larynx_SIR, c(0.2, 0.4, 0.6, 0.8))
breaks4 = quantile(ca.poly$colrect_SIR, c(0.2, 0.4, 0.6, 0.8))


color = rev(brewer.pal(5,"RdBu"))

ch_edge1 <- bd_diffl1[which(raw_diff1 == 1)]
edge_plot1 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group, fill = lung_SIR),
               #color = ifelse(ca.poly.df$zero_pos == 1, "white", "black"),
               #alpha = ifelse(ca.poly.df$zero_pos == 1, 0.5, 1),
               color = "black",
               alpha = 0.6
               #fill = "white"
               #linetype = ifelse(ca.poly.df$zero_pos == 1, "dotted", "solid")
  ) + 
  scale_fill_stepsn(colors = color,  
                    breaks = breaks1)

for(i in which(raw_diff1 == 1)){
  edge_plot1 = edge_plot1 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red", 
                                      size = ((bd_diffl1[i]- min(ch_edge1))/(max(ch_edge1) - min(ch_edge1)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=16, face="bold",hjust = 0.5),
          legend.title = element_blank()) + 
    #ggtitle("Lung (T = 75)")
    ggtitle(paste("Lung"))
}


ch_edge2 <- bd_diffl2[which(raw_diff2 == 1)]
edge_plot2 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group, fill = esophagus_SIR),
               #color = ifelse(ca.poly.df$zero_pos == 1, "white", "black"),
               #alpha = ifelse(ca.poly.df$zero_pos == 1, 0.5, 1),
               color = "black",
               alpha = 0.6
               #fill = "white"
               #linetype = ifelse(ca.poly.df$zero_pos == 1, "dotted", "solid")
  ) + 
  scale_fill_stepsn(colors = color,  
                    breaks = breaks2)
for(i in which(raw_diff2 == 1)){
  edge_plot2 = edge_plot2 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red", 
                                      size = ((bd_diffl2[i]- min(ch_edge2))/(max(ch_edge2) - min(ch_edge2)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=16, face="bold",hjust = 0.5),
          legend.title = element_blank()) + 
    ggtitle(paste("Esophageal"))
}

ch_edge3 <- bd_diffl3[which(raw_diff3 == 1)]
edge_plot3 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group, fill = larynx_SIR),
               #color = ifelse(ca.poly.df$zero_pos == 1, "white", "black"),
               #alpha = ifelse(ca.poly.df$zero_pos == 1, 0.5, 1),
               color = "black",
               alpha = 0.6
               #fill = "white"
               #linetype = ifelse(ca.poly.df$zero_pos == 1, "dotted", "solid")
  ) + 
  scale_fill_stepsn(colors = color,  
                    breaks = breaks3)
for(i in which(raw_diff3 == 1)){
  edge_plot3 = edge_plot3 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red", 
                                      size = ((bd_diffl3[i]- min(ch_edge3))/(max(ch_edge3) - min(ch_edge3)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=16, face="bold",hjust = 0.5),
          legend.title = element_blank()) + 
    #ggtitle("Larynx (T = 59)")
    ggtitle(paste("Larynx"))
}

ch_edge4 <- bd_diffl4[which(raw_diff4 == 1)]
edge_plot4 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group, fill = colrect_SIR),
               #color = ifelse(ca.poly.df$zero_pos == 1, "white", "black"),
               #alpha = ifelse(ca.poly.df$zero_pos == 1, 0.5, 1),
               color = "black",
               alpha = 0.6
               #fill = "white"
               #linetype = ifelse(ca.poly.df$zero_pos == 1, "dotted", "solid")
  ) + 
  scale_fill_stepsn(colors = color,  
                    breaks = breaks4)
for(i in which(raw_diff4 == 1)){
  edge_plot4 = edge_plot4 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red", 
                                      size = ((bd_diffl4[i]- min(ch_edge4))/(max(ch_edge4) - min(ch_edge4)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=16, face="bold",hjust = 0.5),
          legend.title = element_blank()) + 
    #ggtitle("Colorectal (T = 67)")
    ggtitle(paste("Colorectal"))
}

ggarrange(edge_plot1, edge_plot2, edge_plot3, edge_plot4, nrow = 2, ncol = 2)

```

### 3. Hierarchical models for difference boundary detection of the incidence of four cancers
We apply two joint models (MDAGAR and MCAR) and two independent-disease models (DAGAR$_{ind}$ and CAR$_{ind}$) for the multivariate difference boundary detection. We only include intercept for fixed effects and exclude covariates in this section.

The two joint models are implemented using Metropolis within Gibbs Sampler in R, and two independent-disease models are implemented using rjags. For MDAGAR models, we detect difference boundaries for each cancer individually, shared difference boundaries, and cross-cancer boundaries. For MCAR model, we only show difference boundary detection for each cancer individually. For DAGAR$_{ind}$ and CAR$_{ind}$ models, the difference boundaries for each cancer individually and shared difference boundaries are detected to compare with joint models. We also calculate D scores to compare the model fitting for four models.

For each model, it takes up to 24 hrs to run.

#### 3.1 Data in models without covariates

We run models for incidence data without covariates (only include intercept)
```{r}
Y1 = rate_lung1$O_count[final_perm]
Y2 = rate_esophagus1$O_count[final_perm]
Y3 = rate_larynx1$O_count[final_perm]
Y4 = rate_colrect1$O_count[final_perm]

E1 = rate_lung1$E_count[final_perm]
E2 = rate_esophagus1$E_count[final_perm]
E3 = rate_larynx1$E_count[final_perm]
E4 = rate_colrect1$E_count[final_perm]

X1 = as.matrix(cbind(1,rate_lung1[,6:14]))[final_perm,]
X2 = as.matrix(cbind(1,rate_esophagus1[,6:14]))[final_perm,]
X3 = as.matrix(cbind(1,rate_larynx1[,6:14]))[final_perm,]
X4 = as.matrix(cbind(1,rate_colrect1[,6:14]))[final_perm,]

Y = c(Y1,Y2,Y3,Y4)
E = c(E1, E2, E3, E4)
X = as.matrix(bdiag(bdiag(X1[,1], X2[,1]), bdiag(X3[,1],X4[,1])))
```


#### 3.2 Joint models
Functions for Joint models
```{r}
# compute stick-breaking weights
makeprobs<-function(v){
  m<-length(v) 
  p <-matrix(0,m,m)
  for(j in 2:m){
    p[1:(j-1),j]<-1
  }
  probs<-exp(log(v)+log(1-v)%*%p)
  probs[m]<-1-sum(probs[1:(m-1)])
  probs
}

# compute u weights from F_r and prob 
makeu<- function(F_r,probs){
  
  m1=length(F_r)
  m2=length(probs)
  u=rep(0,m1)
  for (k in 1:m1){
    for (l in 1:m2){
      if (sum(probs[1:(l-1)])<F_r[k] && F_r[k]<sum(probs[1:l])){
        u[k]=l
      }
      if (u[k]==0){
        u[k]=1
      }
    }
  }
  return(u)
}

# Jacobian matrix for updating A
Jacob_A <- function(A){
  prod = 1
  for(i in 1:nrow(A)){
    prod = prod*A[i,i]^(nrow(A)-i+1)
  }
  return(2^nrow(A)*prod)
}
```

##### 3.2.1 Multivariate Joint DAGAR (MDAGAR) Model: Implement Metropolis within Gibbs Sampler for MCMC updating 
```{r}
# Compute presicion matrices of DAGAR 
Dinv_new <- function(Rho, n, cn, ns, udnei,q){
  Tau <- list()
  B <- list()
  invD <- list()
  for(i in 1:q){
    Tau[[i]] <- diag(n)
    B[[i]] <- matrix(0, n, n)
    for(j in 3:n){
      Tau[[i]][j,j] <- (1 + (ns[j-1] - 1) * Rho[i]^2) / (1 - Rho[i]^2)
      for(k in (udnei[(cn[j-1] + 1):(cn[j-1] + ns[j-1])])){
        B[[i]][j,k] <- Rho[i] / (1 + (ns[j-1] - 1) * Rho[i]^2)
      }
    }
    invD[[i]] <- t(diag(n) - B[[i]]) %*% Tau[[i]] %*% (diag(n) - B[[i]])
  }
  return(invD)
}

#######################################
### MDAGAR model with no covariates ###
#######################################

# Metropolis within Gibbs Sampler for MCMC updating 
ARDP_joint_diff<-function(y=Y, x1=as.matrix(X1[,1]), x2=as.matrix(X2[,1]), x3 = as.matrix(X3[,1]), x4 = as.matrix(X4[,1]),
                          X = X, E = E, Minc, alpha=1, q = 4, n.atoms=15, runs=10000, burn=1000){
  #y:       data
  #x:       covariates
  #n.atoms: number of atoms in the mixture dist.
  #theta:      the theta's (iid from baseline) in the model
  #alpha:     v~beta(1,alpha)
  #u:       the index indicator of spatial random effect
  #rho:     sptatial autocorrelation parameter in DAGAR
  #Minc:       0-1 adjacency matrix
  #V_r:     covariance matrix of joint MDAGAR
  #Q:       presicion matrix of DAGAR
  #r:       random effects following DAGAR
  #F_r:     Marginal CDF of r
  #taued:   presicion in Baseline of DP for disease d
  #taus:    precision for theta
  
  
  
  
  nq<-length(y)
  n <- nq / q
  p<-ncol(X)
  y1 <- y[1:n]
  y2 <- y[(n+1):(2*n)]
  y3 <- y[(2*n+1):(3*n)]
  y4 <- y[(3*n+1):(4*n)]
  
  E1 <- E[1:n]
  E2 <- E[(n+1):(2*n)]
  E3 <- E[(2*n+1):(3*n)]
  E4 <- E[(3*n+1):(4*n)]
  
  sigmasq_beta = 10000
  keepbeta=matrix(0,runs,p)
  keepphi=matrix(0,runs,nq)
  keeptheta=matrix(0,runs,n.atoms)
  keepA=matrix(0,runs,10)
  keepu=matrix(0,runs,nq)
  keeprho1=keeprho2=keeprho3=keeprho4=keeptaus = rep(0,runs)
  keepv=matrix(0,runs,n.atoms)
  keepr=matrix(0,runs,nq)
  keepFr=matrix(0,runs,nq)
  
  
  #initial values
  
  theta=rep(0,n.atoms)
  
  beta1<-rep(0,ncol(x1))
  beta2<-rep(0,ncol(x2))
  beta3<-rep(0,ncol(x3))
  beta4<-rep(0,ncol(x4))
  beta = c(beta1, beta2, beta3, beta4)
  
  taus = 1
  
  c=2
  d=0.1
  d2=1
  v<-rep(.1,n.atoms)
  probs=makeprobs(v)
  
  #A matrix
  rho=c(0.5, 0.5, 0.5, 0.5)
  eta = log(rho/(1-rho))
  A = matrix(0, q, q)
  for(i in 1:q){
    for(j in 1:i){
      if(j == i){
        A[i, j] = exp(rnorm(1))
      }else{
        A[i, j] = rnorm(1)
      }
    }
  }
  #Sigma = A%*%t(A)
  
  Q = Dinv_new(Rho = rho, n, cn, ns, udnei, q)
  invQ1 = solve(Q[[1]])
  invQ2 = solve(Q[[2]])
  invQ3 = solve(Q[[3]])
  invQ4 = solve(Q[[4]])
  
  #f1 = rmvnorm(1, rep(0, n), invQ1)
  #f2 = rmvnorm(1, rep(0, n), invQ2)
  #f = c(f1, f2)
  kprod = kronecker(A, diag(n))
  invQ = as.matrix(bdiag(bdiag(invQ1, invQ2), bdiag(invQ3, invQ4)))
  #r = kprod %*% f
  
  Vr = as.matrix(forceSymmetric(kprod %*% invQ %*% t(kprod)))
  r = rmvnorm(1, rep(0, nq), Vr)
  
  
  F_r=pnorm(r,0,sqrt(diag(Vr)))
  u=makeu(F_r,probs)
  phi <- theta[u]
  
  nu = 2
  R = 0.1 * diag(q)
  
  
  acceptr=acceptv=acceptrho=acceptA=0
  acceptbeta1 = acceptbeta2=acceptbeta3=acceptbeta4 = 0
  accepttheta = 0
  
  #sumtp<-summu<-summu2<-rep(0,n)
  count<-afterburn<-0
  burn = burn + 1
  #iter = 1
  #runs = 1000
  for(iter in 1:runs){
    
    if(iter %% 100 == 0){
      print(iter)
      print(acceptA/(iter-1))
      print(acceptr/nq/(iter-1))
      print(acceptrho/(iter-1))
      print(acceptv/n.atoms/(iter-1))
      print(accepttheta/n.atoms/(iter-1))
      print(acceptbeta1/(iter-1))
      print(acceptbeta2/(iter-1))
      print(acceptbeta3/(iter-1))
      print(acceptbeta4/(iter-1))
    }
    ######################
    ###   update beta  ###
    ######################
    # update beta (intercept only model)
    
    pro_beta1=rnorm(1,beta1,0.02)
    MHrate1=sum(-E1*exp(pro_beta1*x1+theta[u][1:n])+y1*(pro_beta1*x1+theta[u][1:n]))-
      sum(-E1*exp(beta1*x1+theta[u][1:n])+y1*(beta1*x1+theta[u][1:n]))
    
    if(runif(1,0,1)<exp(MHrate1)){
      beta1<-pro_beta1
      acceptbeta1=acceptbeta1+1
    }
    
    pro_beta2=rnorm(1,beta2,0.16)
    MHrate2=sum(-E2*exp(pro_beta2*x2+theta[u][(n+1):(2*n)])+y2*(pro_beta2*x2+theta[u][(n+1):(2*n)]))-
      sum(-E2*exp(beta2*x2+theta[u][(n+1):(2*n)])+y2*(beta2*x2+theta[u][(n+1):(2*n)]))
    
    if(runif(1,0,1)<exp(MHrate2)){
      beta2<-pro_beta2
      acceptbeta2=acceptbeta2+1
    }
    
    pro_beta3=rnorm(1,beta3,0.05)
    MHrate3=sum(-E3*exp(pro_beta3*x3+theta[u][(2*n+1):(3*n)])+y3*(pro_beta3*x3+theta[u][(2*n+1):(3*n)]))-
      sum(-E3*exp(beta3*x3+theta[u][(2*n+1):(3*n)])+y3*(beta3*x3+theta[u][(2*n+1):(3*n)]))
    
    if(runif(1,0,1)<exp(MHrate3)){
      beta3<-pro_beta3
      acceptbeta3=acceptbeta3+1
    }
    
    pro_beta4=rnorm(1,beta4,0.02)
    MHrate4=sum(-E4*exp(pro_beta4*x4+theta[u][(3*n+1):(4*n)])+y4*(pro_beta4*x4+theta[u][(3*n+1):(4*n)]))-
      sum(-E4*exp(beta4*x4+theta[u][(3*n+1):(4*n)])+y4*(beta4*x4+theta[u][(3*n+1):(4*n)]))
    
    if(runif(1,0,1)<exp(MHrate4)){
      beta4<-pro_beta4
      acceptbeta4=acceptbeta4+1
    }
    
    beta <- c(beta1, beta2, beta3, beta4)
    
    
    #########################
    ###   update theta    ###
    #########################
    
    u1 = u[1:n]
    u2 = u[(n+1):(2*n)]
    u3 = u[(2*n+1):(3*n)]
    u4 = u[(3*n+1):(4*n)]
    
    
    
    for (j in 1:n.atoms){
      
      count[j]=sum(u==j)
      
      if (count[j]==0){
        theta[j]=rnorm(1,0,sqrt(1/taus))
      }else{
        pro_theta=rnorm(1,theta[j],0.06)
        
        MHrate<-sum(-(E[u==j])*exp(X[u==j,] %*% beta+pro_theta)+(y[u==j])*(X[u==j,] %*% beta+pro_theta))-taus/2*pro_theta^2-
          sum(-(E[u==j])*exp(X[u==j,] %*% beta+theta[j])+(y[u==j])*(X[u==j,] %*% beta+theta[j]))+taus/2*theta[j]^2
        if(runif(1,0,1)<exp(MHrate)){
          theta[j]<-pro_theta
          accepttheta=accepttheta+1
        }
      }
    }
    
    
    
    ######################
    ###   update r     ###
    ######################
    
    #pro_r=t(rmvnorm(1,r,sigma=covcar2,method="chol"))
    
    
    #tauvec = c(rep(taue1, n), rep(taue2, n), rep(taue3, n), rep(taue4, n))
    
    #pro_r = rmvnorm(1, r, 0.01*diag(nq))
    
    for (k in 1:nq){
      pro_r=r;pro_Fr=F_r;pro_u=u
      pro_r[k]=rnorm(1,r[k],1.8)
      pro_Fr[k]=pnorm(pro_r[k],0,sqrt(Vr[k,k]))
      pro_u[k]=makeu(pro_Fr[k],probs)
      
      MH=dmvnorm(pro_r,mean=rep(0, nq),sigma=Vr,log=T)+dpois(y[k],E[k]*exp(as.numeric(X[k,]%*%beta)+theta[pro_u[k]]),log=T)-
        dmvnorm(r,mean=rep(0, nq),sigma=Vr,log=T)-dpois(y[k],E[k]*exp(as.numeric(X[k,]%*%beta)+theta[u[k]]),log=T)
      #MH=dmvnorm(pro_r,mean=rep(0, nq),sigma=Vr,log=T)+dmvnorm(t(as.matrix(y1)),mean=as.vector(x1%*%t(beta1))+theta[pro_u[1:n]],sigma=1/taue1*diag(n),log=T)+
      # dmvnorm(t(as.matrix(y2)),mean=as.vector(x2%*%t(beta2))+theta[pro_u[(n+1):(2*n)]],sigma=1/taue2*diag(n),log=T)-
      #dmvnorm(r,mean=rep(0, nq),sigma=Vr,log=T)-dmvnorm(t(as.matrix(y1)),mean=as.vector(x1%*%t(beta1))+theta[u[1:n]],sigma=1/taue1*diag(n),log=T)-
      #dmvnorm(t(as.matrix(y2)),mean=as.vector(x2%*%t(beta2))+theta[u[(n+1):(2*n)]],sigma=1/taue2*diag(n),log=T)
      if(runif(1,0,1)<exp(MH)){
        r[k]=pro_r[k]
        F_r[k]=pro_Fr[k]
        u[k]=pro_u[k]
        acceptr=acceptr+1
      }
    }
    
    
    ######################
    ###   update   v   ###
    ######################
    
    
    #pro_v=v
    for (j in 1:n.atoms){
      #0.9
      pro_v=v
      pro_v[j]<-rnorm(1,v[j],0.06)
      while (pro_v[j]<0 || pro_v[j]>1) {pro_v[j]<-rnorm(1,v[j],0.06)}
      #if(pro_v[j]>0 & pro_v[j]<1){
      pro_probs=makeprobs(pro_v)
      pro_u=makeu(F_r,pro_probs)
      
      MH=log(dbeta(pro_v[j],1,alpha))+sum(dpois(y,E*exp(X%*%beta+theta[pro_u]),log=T))-
        log(dbeta(v[j],1,alpha))-sum(dpois(y,E*exp(X%*%beta+theta[u]),log=T))
      
      
      if(runif(1,0,1)<exp(MH)){
        v[j]=pro_v[j]
        probs=pro_probs
        u=pro_u
        acceptv=acceptv+1
      }
      # }
    }
    
    
    ######################
    ###   update taus  ###
    ######################
    
    
    taus=rgamma(1,shape=1/2*n.atoms+c,rate=sum(theta^2)/2+d2)
    
    
    ######################
    ###   update rho  ###
    ######################
    
    
    pro_eta1 = rnorm(1, eta[1], 0.4)
    pro_eta2 = rnorm(1, eta[2], 0.4)
    pro_eta3 = rnorm(1, eta[3], 0.4)
    pro_eta4 = rnorm(1, eta[4], 0.4)
    pro_eta = c(pro_eta1, pro_eta2, pro_eta3, pro_eta4)
    pro_rho = exp(pro_eta)/(1+exp(pro_eta))
    pro_Q = Dinv_new(Rho = pro_rho, n, cn, ns, udnei, q)
    pro_invQ1 = solve(pro_Q[[1]])
    pro_invQ2 = solve(pro_Q[[2]])
    pro_invQ3 = solve(pro_Q[[3]])
    pro_invQ4 = solve(pro_Q[[4]])
    
    kprod = kronecker(A, diag(n))
    pro_invQ = as.matrix(bdiag(bdiag(pro_invQ1, pro_invQ2), bdiag(pro_invQ3, pro_invQ4)))
    pro_Vr = kprod %*% pro_invQ %*% t(kprod)
    
    MH = dmvnorm(r,mean=rep(0, nq),sigma=as.matrix(forceSymmetric(pro_Vr)),log=T) +
      log(pro_rho[1]) + log(1-pro_rho[1]) + log(pro_rho[2]) + log(1-pro_rho[2]) +
      log(pro_rho[3]) + log(1-pro_rho[3]) + log(pro_rho[4]) + log(1-pro_rho[4]) -
      dmvnorm(r,mean=rep(0, nq),sigma=Vr,log=T) - log(rho[1]) - log(1-rho[1]) - log(rho[2]) - log(1-rho[2]) -
      log(rho[3]) - log(1-rho[3]) - log(rho[4]) - log(1-rho[4])
    
    if(runif(1,0,1)<exp(MH)){
      eta = pro_eta
      rho = pro_rho
      Vr = as.matrix(forceSymmetric(pro_Vr))
      acceptrho=acceptrho+1
    }
    
    ######################
    ###   update A  ###
    ######################
    #0.06
    pro_A = matrix(0, q, q)
    for(i in 1:q){
      for(j in 1:i){
        if(j == i){
          pro_A[i, j] = exp(rnorm(1, log(A[i, j]), 0.04))
          #pro_A[i, j] = rlnorm(1, log(A[i, j]), 0.1)
        }else{
          pro_A[i, j] = rnorm(1, A[i, j], 0.03)
        }
      }
    }
    
    Q = Dinv_new(Rho = rho, n, cn, ns, udnei, q=4)
    invQ1 = solve(Q[[1]])
    invQ2 = solve(Q[[2]])
    invQ3 = solve(Q[[3]])
    invQ4 = solve(Q[[4]])
    Sigma =  A %*% t(A)
    
    pro_kprod = kronecker(pro_A, diag(n))
    invQ = as.matrix(bdiag(bdiag(invQ1, invQ2), bdiag(invQ3, invQ4)))
    pro_Vr = pro_kprod %*% as.matrix(forceSymmetric(invQ)) %*% t(pro_kprod)
    pro_Sigma =  pro_A %*% t(pro_A)
    
    lpA = -(nu+4)/2*logdet(Sigma) - 1/2*sum(diag(nu*R%*%solve(Sigma))) + log(Jacob_A(A)) + sum(log(diag(A)))
    pro_lpA = -(nu+4)/2*logdet(pro_Sigma) - 1/2*sum(diag(nu*R%*%solve(pro_Sigma))) + log(Jacob_A(pro_A)) + sum(log(diag(pro_A)))
    MH = dmvnorm(r,mean=rep(0, nq),sigma=as.matrix(forceSymmetric(pro_Vr)),log=T) + pro_lpA -
      dmvnorm(r,mean=rep(0, nq),sigma=Vr,log=T) - lpA
    
    if(runif(1,0,1)<exp(MH)){
      A = pro_A
      Vr = as.matrix(forceSymmetric(pro_Vr))
      acceptA=acceptA+1
    }
    
    #kprod = kronecker(A, diag(n))
    #Vr = as.matrix(forceSymmetric(kprod %*% invQ %*% t(kprod)))
    
    ######################
    ### record samples ###
    ######################
    
    keeptheta[iter,] = theta
    keepphi[iter,] = theta[u]
    
    keepA[iter,] = as.vector(A)[-c(5,9,10,13:15)]
    
    keepbeta[iter,]=beta
    keeptaus[iter]=taus
    keeprho1[iter]=rho[1]
    keeprho2[iter]=rho[2]
    keeprho3[iter]=rho[3]
    keeprho4[iter]=rho[4]
    keepu[iter,]=u
    keepv[iter,]=v
    keepr[iter,]=r
    keepFr[iter,]=F_r
    
    #    cat("iteration = ", i, "acceptance rate of r = ", acceptr/(n*i),"acceptance rate of v = ",acceptv/(n.atoms*i), "\n")
    
  }
  
  list(beta=keepbeta[burn:runs,], A = keepA[burn:runs,],phi=keepphi[burn:runs,],theta=keeptheta[burn:runs,],u=keepu[burn:runs,],
       v=keepv[burn:runs,],r=keepr[burn:runs,],Fr=keepFr[burn:runs,], taus=keeptaus[burn:runs],
       rho1=keeprho1[burn:runs], rho2=keeprho2[burn:runs], rho3=keeprho3[burn:runs], rho4=keeprho4[burn:runs])
}

set.seed(123)
mcmc_samples=ARDP_joint_diff(y=Y, x1=as.matrix(X1[,1]), x2=as.matrix(X2[,1]), x3 = as.matrix(X3[,1]), x4 = as.matrix(X4[,1]), 
                             X = X, E=E, Minc, alpha=1, q = 4, n.atoms=15, runs=30000, burn=20000)

#Estimates of parameters
sample.mcmc = cbind(mcmc_samples$taus, mcmc_samples$rho1, mcmc_samples$rho2, mcmc_samples$rho3, mcmc_samples$rho4)

```

###### I. Estimates for $\beta$ and $\rho_d$ in Table 5
```{r}
mysummary = function(invector) {
  c(mean(invector), quantile(invector, .025), quantile(invector,.975))
}
# Estimates for intercepts
round(t(apply(mcmc_samples$beta, 2, mysummary)), 3)
# Estimates for $rho_d$
round(mysummary(mcmc_samples$rho1), 3)
round(mysummary(mcmc_samples$rho2), 3)
round(mysummary(mcmc_samples$rho3), 3)
round(mysummary(mcmc_samples$rho4), 3)
```

###### II. Obtain spatial randome effects in original spatial order
```{r}
phis <- mcmc_samples$phi
phis1 <- phis[,1:58][,order(final_perm)]
phis2 <- phis[,59:116][,order(final_perm)]
phis3 <- phis[,117:174][,order(final_perm)]
phis4 <- phis[,175:232][,order(final_perm)]
phis_origin <- cbind(phis1, phis2, phis3, phis4)
```

###### III. Difference boundaries for each cancer individually
We calculate posterior probabilities: $P(\phi_{id} \neq \phi_{jd} | y), \quad i \sim j$
```{r}
vij_samples1 <- t(apply(phis1, 1, function(x){
  diff_b1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]]){
      diff_b1 <- c(diff_b1, 1)
    }else{
      diff_b1 <- c(diff_b1, 0)
    }
  }
  return(diff_b1)
}))

vij_samples2 <- t(apply(phis2, 1, function(x){
  diff_b1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]]){
      diff_b1 <- c(diff_b1, 1)
    }else{
      diff_b1 <- c(diff_b1, 0)
    }
  }
  return(diff_b1)
}))

vij_samples3 <- t(apply(phis3, 1, function(x){
  diff_b1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]]){
      diff_b1 <- c(diff_b1, 1)
    }else{
      diff_b1 <- c(diff_b1, 0)
    }
  }
  return(diff_b1)
}))

vij_samples4 <- t(apply(phis4, 1, function(x){
  diff_b1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]]){
      diff_b1 <- c(diff_b1, 1)
    }else{
      diff_b1 <- c(diff_b1, 0)
    }
  }
  return(diff_b1)
}))

# probabilities: P(phi_id != phi_jd), i~j
pvij1 <- apply(vij_samples1, 2, mean)
pvij2 <- apply(vij_samples2, 2, mean)
pvij3 <- apply(vij_samples3, 2, mean)
pvij4 <- apply(vij_samples4, 2, mean)

names(pvij1) <- neighbor_name
names(pvij2) <- neighbor_name
names(pvij3) <- neighbor_name
names(pvij4) <- neighbor_name

```

####### a. Figure 6 (a): Estimated FDR curves plotted against the number of selected difference boundaries for four cancers using MDAGAR
```{r}
T_edge1 = seq(0, 135, 1)[-1]
threshold1 = sort(pvij1, decreasing = TRUE)[T_edge1]
threshold2 = sort(pvij2, decreasing = TRUE)[T_edge1]
threshold3 = sort(pvij3, decreasing = TRUE)[T_edge1]
threshold4 = sort(pvij4, decreasing = TRUE)[T_edge1]
FDR_est1 <- rep(0, length(T_edge1))
FDR_est2 <- rep(0, length(T_edge1))
FDR_est3 <- rep(0, length(T_edge1))
FDR_est4 <- rep(0, length(T_edge1))


for(i in 1:length(threshold1)){
  th1 <- threshold1[i]
  est_diff1 <- as.numeric(pvij1 >= th1)
  FDR_est1[i] <- sum((1-pvij1) * est_diff1)  / sum(est_diff1)
  
  th2 <- threshold2[i]
  est_diff2 <- as.numeric(pvij2 >= th2)
  FDR_est2[i] <- sum((1-pvij2) * est_diff2)  / sum(est_diff2)
  
  th3 <- threshold3[i]
  est_diff3 <- as.numeric(pvij3 >= th3)
  FDR_est3[i] <- sum((1-pvij3) * est_diff3)  / sum(est_diff3)
  
  th4 <- threshold4[i]
  est_diff4 <- as.numeric(pvij4 >= th4)
  FDR_est4[i] <- sum((1-pvij4) * est_diff4)  / sum(est_diff4)
}

plot(FDR_est1,type="l", lty = 1, xlab="Number of edges selected", ylab = "Estimated FDR", ylim=c(0,0.5))
lines(FDR_est2,type="l", lty = 2)
lines(FDR_est3,type="l", lty = 3)
lines(FDR_est4,type="l", lty = 4)
legend("topleft", legend=c("Lung", "Esophageal", "Layrnx", "Colorectal"),
       lty=1:4, cex=0.7)
```

####### b. Table 3: Names of adjacent counties that have significant boundary effect for each cancer when $\delta = 0.05$ 
```{r}
# Set threshold for FDR: 0.05
alpha_n = 0.05
T1 = sum(FDR_est1<=alpha_n)
T2 = sum(FDR_est2<=alpha_n)
T3 = sum(FDR_est3<=alpha_n)
T4 = sum(FDR_est4<=alpha_n)

name_diff1 <- names(pvij1[order(pvij1,decreasing = T)][1:T1])
name_diff2 <- names(pvij2[order(pvij2,decreasing = T)][1:T2])
name_diff3 <- names(pvij3[order(pvij3,decreasing = T)][1:T3])
name_diff4 <- names(pvij4[order(pvij4,decreasing = T)][1:T4])

# Lung
name_diff1
# Esophageal
name_diff2
# Larynx
name_diff3
# Colorectal
name_diff4
```

Number $1-35$ for lung cancer, $1-4$ for esophageal cancer and $1-12$ for colorectal cancer are ranked by initial letters with $P(\phi_{id} \neq \phi_{jd} | y) = 1$.

####### c. Difference boundary detection
By setting threshold for FDR as $0.05$, we get difference boundaries detected for each cancer individually.

Figure 7 (a): Difference boundaries (highlighted in red) detected by MAGAR in SIR map for four cancers individually when $\delta = 0.05$.
```{r, fig.height=7, fig.width=10}

# Difference boundary for each cancer with FDR <= 0.05
est_diff1 <- as.numeric(pvij1 >= threshold1[T1])

est_diff2 <- as.numeric(pvij2 >= threshold2[T2])

est_diff3 <- as.numeric(pvij3 >= threshold3[T3])

est_diff4 <- as.numeric(pvij4 >= threshold4[T4])

# Lung cancer
ch_edge1 <- pvij1[which(est_diff1 == 1)]
edge_plot1 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group, fill = lung_SIR),
               color = "black",
               alpha = 0.6
  ) +
  scale_fill_stepsn(colors = color,
                    breaks = breaks1)

for(i in which(est_diff1 == 1)){
  edge_plot1 = edge_plot1 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                      size = ((pvij1[i]- min(ch_edge1))/(max(ch_edge1) - min(ch_edge1)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=16, face="bold",hjust = 0.5),
          legend.title = element_blank()) +
    ggtitle(paste("Lung (T = ", T1, ")", sep=""))
}

# Esophageal cancer
ch_edge2 <- pvij2[which(est_diff2 == 1)]
edge_plot2 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group, fill = esophagus_SIR),
               color = "black",
               alpha = 0.6
  ) +
  scale_fill_stepsn(colors = color,
                    breaks = breaks2)
for(i in which(est_diff2 == 1)){
  edge_plot2 = edge_plot2 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                      size = ((pvij2[i]- min(ch_edge2))/(max(ch_edge2) - min(ch_edge2)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=16, face="bold",hjust = 0.5),
          legend.title = element_blank()) +
    ggtitle(paste("Esophageal (T = ", T2, ")", sep=""))
}

# Larynx cancer
ch_edge3 <- pvij3[which(est_diff3 == 1)]
edge_plot3 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group, fill = larynx_SIR),
               color = "black",
               alpha = 0.6
  ) +
  scale_fill_stepsn(colors = color,
                    breaks = breaks3)
for(i in which(est_diff3 == 1)){
  edge_plot3 = edge_plot3 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                      size = ((pvij3[i]- min(ch_edge3))/(max(ch_edge3) - min(ch_edge3)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=16, face="bold",hjust = 0.5),
          legend.title = element_blank()) +
    ggtitle(paste("Larynx (T = ", T3, ")", sep=""))
}

# Colorectal cancer
ch_edge4 <- pvij4[which(est_diff4 == 1)]
edge_plot4 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group, fill = colrect_SIR),
               color = "black",
               alpha = 0.6
  ) +
  scale_fill_stepsn(colors = color,
                    breaks = breaks4)
for(i in which(est_diff4 == 1)){
  edge_plot4 = edge_plot4 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                      size = ((pvij4[i]- min(ch_edge4))/(max(ch_edge4) - min(ch_edge4)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=16, face="bold",hjust = 0.5),
          legend.title = element_blank()) +
    ggtitle(paste("Colorectal (T = ", T4, ")", sep=""))
}


ggarrange(edge_plot1, edge_plot2, edge_plot3, edge_plot4, nrow = 2, ncol = 2) 
```

###### IV. Shared difference boundary for each pair of cancers
We calculate posterior probabilities: $P(\phi_{id} \neq \phi_{jd}, \phi_{id'} \neq \phi_{jd'} | y),\quad i \sim j,\quad d \neq d'$
```{r}
vij_samplesc12 <- t(apply(cbind(phis1,phis2), 1, function(x){
  diff_bc1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]] & x[neighbor_list0[i,1]+58] != x[neighbor_list0[i,2]+58]){
      diff_bc1 <- c(diff_bc1, 1)
    }else{
      diff_bc1 <- c(diff_bc1, 0)
    }
  }
  return(diff_bc1)
}))

vij_samplesc13 <- t(apply(cbind(phis1,phis3), 1, function(x){
  diff_bc1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]] & x[neighbor_list0[i,1]+58] != x[neighbor_list0[i,2]+58]){
      diff_bc1 <- c(diff_bc1, 1)
    }else{
      diff_bc1 <- c(diff_bc1, 0)
    }
  }
  return(diff_bc1)
}))

vij_samplesc14 <- t(apply(cbind(phis1,phis4), 1, function(x){
  diff_bc1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]] & x[neighbor_list0[i,1]+58] != x[neighbor_list0[i,2]+58]){
      diff_bc1 <- c(diff_bc1, 1)
    }else{
      diff_bc1 <- c(diff_bc1, 0)
    }
  }
  return(diff_bc1)
}))

vij_samplesc23 <- t(apply(cbind(phis2,phis3), 1, function(x){
  diff_bc1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]] & x[neighbor_list0[i,1]+58] != x[neighbor_list0[i,2]+58]){
      diff_bc1 <- c(diff_bc1, 1)
    }else{
      diff_bc1 <- c(diff_bc1, 0)
    }
  }
  return(diff_bc1)
}))

vij_samplesc24 <- t(apply(cbind(phis2,phis4), 1, function(x){
  diff_bc1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]] & x[neighbor_list0[i,1]+58] != x[neighbor_list0[i,2]+58]){
      diff_bc1 <- c(diff_bc1, 1)
    }else{
      diff_bc1 <- c(diff_bc1, 0)
    }
  }
  return(diff_bc1)
}))

vij_samplesc34 <- t(apply(cbind(phis3,phis4), 1, function(x){
  diff_bc1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]] & x[neighbor_list0[i,1]+58] != x[neighbor_list0[i,2]+58]){
      diff_bc1 <- c(diff_bc1, 1)
    }else{
      diff_bc1 <- c(diff_bc1, 0)
    }
  }
  return(diff_bc1)
}))

# probabilities: P(phi_id != phi_jd, phi_id' != phi_jd'), i ~ j, d != d'
pvijc12 <- apply(vij_samplesc12, 2, mean)
pvijc13 <- apply(vij_samplesc13, 2, mean)
pvijc14 <- apply(vij_samplesc14, 2, mean)
pvijc23 <- apply(vij_samplesc23, 2, mean)
pvijc24 <- apply(vij_samplesc24, 2, mean)
pvijc34 <- apply(vij_samplesc34, 2, mean)

# Estimated FDR
thresholdc12 = sort(pvijc12, decreasing = TRUE)[T_edge1]
thresholdc13 = sort(pvijc13, decreasing = TRUE)[T_edge1]
thresholdc14 = sort(pvijc14, decreasing = TRUE)[T_edge1]
thresholdc23 = sort(pvijc23, decreasing = TRUE)[T_edge1]
thresholdc24 = sort(pvijc24, decreasing = TRUE)[T_edge1]
thresholdc34 = sort(pvijc34, decreasing = TRUE)[T_edge1]

FDR_estc12 <- rep(0, length(T_edge1))
FDR_estc13 <- rep(0, length(T_edge1))
FDR_estc14 <- rep(0, length(T_edge1))
FDR_estc23 <- rep(0, length(T_edge1))
FDR_estc24 <- rep(0, length(T_edge1))
FDR_estc34 <- rep(0, length(T_edge1))


for(i in 1:length(threshold1)){
  thc12 <- thresholdc12[i]
  est_diffc12 <- as.numeric(pvijc12 >= thc12)
  FDR_estc12[i] <- sum((1-pvijc12) * est_diffc12)  / sum(est_diffc12)
  
  thc13 <- thresholdc13[i]
  est_diffc13 <- as.numeric(pvijc13 >= thc13)
  FDR_estc13[i] <- sum((1-pvijc13) * est_diffc13)  / sum(est_diffc13)
  
  thc14 <- thresholdc14[i]
  est_diffc14 <- as.numeric(pvijc14 >= thc14)
  FDR_estc14[i] <- sum((1-pvijc14) * est_diffc14)  / sum(est_diffc14)
  
  thc23 <- thresholdc23[i]
  est_diffc23 <- as.numeric(pvijc23 >= thc23)
  FDR_estc23[i] <- sum((1-pvijc23) * est_diffc23)  / sum(est_diffc23)
  
  thc24 <- thresholdc24[i]
  est_diffc24 <- as.numeric(pvijc24 >= thc24)
  FDR_estc24[i] <- sum((1-pvijc24) * est_diffc24)  / sum(est_diffc24)
  
  thc34 <- thresholdc34[i]
  est_diffc34 <- as.numeric(pvijc34 >= thc34)
  FDR_estc34[i] <- sum((1-pvijc34) * est_diffc34)  / sum(est_diffc34)
}
```

Use the same threshold as above and identify shared boundaries

Figure 8: Shared difference boundaries (highlighted in red) detected by MDAGAR for each pair of cancers in SIR map when $\delta = 0.05$.
```{r, fig.height=7, fig.width=10}
alpha_nc = 0.05
T12c = sum(FDR_estc12<=alpha_nc)
T13c = sum(FDR_estc13<=alpha_nc)
T14c = sum(FDR_estc14<=alpha_nc)
T23c = sum(FDR_estc23<=alpha_nc)
T24c = sum(FDR_estc24<=alpha_nc)
T34c = sum(FDR_estc34<=alpha_nc)


est_diffc12 <- as.numeric(pvijc12 >= thresholdc12[T12c])
neighbor_list_diffc12 <- neighbor_list0[est_diffc12 == 1, ]

est_diffc13 <- as.numeric(pvijc13 >= thresholdc13[T13c])
neighbor_list_diffc13 <- neighbor_list0[est_diffc13 == 1, ]

est_diffc14 <- as.numeric(pvijc14 >= thresholdc14[T14c])
neighbor_list_diffc14 <- neighbor_list0[est_diffc14 == 1, ]

est_diffc23 <- as.numeric(pvijc23 >= thresholdc23[T23c])
neighbor_list_diffc23 <- neighbor_list0[est_diffc23 == 1, ]

est_diffc24 <- as.numeric(pvijc24 >= thresholdc24[T24c])
neighbor_list_diffc24 <- neighbor_list0[est_diffc24 == 1, ]

est_diffc34 <- as.numeric(pvijc34 >= thresholdc34[T34c])
neighbor_list_diffc34 <- neighbor_list0[est_diffc34 == 1, ]

ch_edge12 <- pvijc12[est_diffc12==1]
edge_plotc12 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group),
               color = "black",
               fill = "white"
  )
for(i in which(est_diffc12==1)){
  edge_plotc12 = edge_plotc12 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                          size = ((pvijc12[i]- min(ch_edge12))/(max(ch_edge12) - min(ch_edge12)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=14, face="bold",hjust = 0.5)) +
    ggtitle(paste("Lung, Esophageal (T = ", T12c, ")", sep=""))
}

ch_edge13 <- pvijc13[est_diffc13==1]
edge_plotc13 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group),
               color = "black",
               fill = "white"
  )
for(i in which(est_diffc13==1)){
  edge_plotc13 = edge_plotc13 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                          size = ((pvijc13[i]- min(ch_edge13))/(max(ch_edge13) - min(ch_edge13)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=14, face="bold",hjust = 0.5)) +
    ggtitle(paste("Lung, Layrnx (T = ", T13c, ")", sep=""))
}

ch_edge14 <- pvijc14[est_diffc14==1]
edge_plotc14 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group),
               color = "black",
               fill = "white"
  )
for(i in which(est_diffc14==1)){
  edge_plotc14 = edge_plotc14 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                          size = ((pvijc14[i]- min(ch_edge14))/(max(ch_edge14) - min(ch_edge14)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=14, face="bold",hjust = 0.5)) +
    ggtitle(paste("Lung, Colorectal (T = ", T14c, ")", sep=""))
}

ch_edge23 <- pvijc23[est_diffc23==1]
edge_plotc23 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group),
               color = "black",
               fill = "white"
  )
for(i in which(est_diffc23==1)){
  edge_plotc23 = edge_plotc23 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                          size = ((pvijc23[i]- min(ch_edge23))/(max(ch_edge23) - min(ch_edge23)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=14, face="bold",hjust = 0.5)) +
    ggtitle(paste("Esophageal, Layrnx (T = ", T23c, ")", sep=""))
}

ch_edge24 <- pvijc24[est_diffc24==1]
edge_plotc24 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group),
               color = "black",
               fill = "white"
  )
for(i in which(est_diffc24==1)){
  edge_plotc24 = edge_plotc24 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                          size = ((pvijc24[i]- min(ch_edge24))/(max(ch_edge24) - min(ch_edge24)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=14, face="bold",hjust = 0.5)) +
    ggtitle(paste("Esophageal, Colorectal (T = ", T24c, ")", sep=""))
}

ch_edge34 <- pvijc34[est_diffc34==1]
edge_plotc34 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group),
               color = "black",
               fill = "white"
  )
for(i in which(est_diffc34==1)){
  edge_plotc34 = edge_plotc34 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                          size = ((pvijc34[i]- min(ch_edge34))/(max(ch_edge34) - min(ch_edge34)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=14, face="bold",hjust = 0.5)) +
    ggtitle(paste("Larynx, Colorectal (T = ", T34c, ")", sep=""))
}

ggarrange(edge_plotc12, edge_plotc13, edge_plotc14, 
          edge_plotc23, edge_plotc24, edge_plotc34, nrow = 2, ncol = 3)
```

###### V. Mutual cross-disease difference boundary
We calculate posterior probabilities: $P(\phi_{id} \neq \phi_{jd'}, \phi_{id'} \neq \phi_{jd}| y),\quad i \sim j, \quad i < j,\quad d \neq d'$

```{r}
vij_samples12 <- t(apply(cbind(phis1,phis2), 1, function(x){
  diff_b1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]+58] &
       x[neighbor_list0[i,1]+58] != x[neighbor_list0[i,2]]){
      diff_b1 <- c(diff_b1, 1)
    }else{
      diff_b1 <- c(diff_b1, 0)
    }
  }
  return(diff_b1)
}))

vij_samples13 <- t(apply(cbind(phis1,phis3), 1, function(x){
  diff_b1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]+58] &
       x[neighbor_list0[i,1]+58] != x[neighbor_list0[i,2]]){
      diff_b1 <- c(diff_b1, 1)
    }else{
      diff_b1 <- c(diff_b1, 0)
    }
  }
  return(diff_b1)
}))

vij_samples14 <- t(apply(cbind(phis1,phis4), 1, function(x){
  diff_b1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]+58] &
       x[neighbor_list0[i,1]+58] != x[neighbor_list0[i,2]]){
      diff_b1 <- c(diff_b1, 1)
    }else{
      diff_b1 <- c(diff_b1, 0)
    }
  }
  return(diff_b1)
}))


vij_samples23 <- t(apply(cbind(phis2,phis3), 1, function(x){
  diff_b1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]+58] &
       x[neighbor_list0[i,1]+58] != x[neighbor_list0[i,2]]){
      diff_b1 <- c(diff_b1, 1)
    }else{
      diff_b1 <- c(diff_b1, 0)
    }
  }
  return(diff_b1)
}))

vij_samples24 <- t(apply(cbind(phis2,phis4), 1, function(x){
  diff_b1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]+58] &
       x[neighbor_list0[i,1]+58] != x[neighbor_list0[i,2]]){
      diff_b1 <- c(diff_b1, 1)
    }else{
      diff_b1 <- c(diff_b1, 0)
    }
  }
  return(diff_b1)
}))

vij_samples34 <- t(apply(cbind(phis3,phis4), 1, function(x){
  diff_b1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]+58] &
       x[neighbor_list0[i,1]+58] != x[neighbor_list0[i,2]]){
      diff_b1 <- c(diff_b1, 1)
    }else{
      diff_b1 <- c(diff_b1, 0)
    }
  }
  return(diff_b1)
}))

# probabilities: P(phi_id != phi_jd', phi_id' != phi_jd), i ~ j, i < j, d != d'
pvij12 <- apply(vij_samples12, 2, mean)
pvij13 <- apply(vij_samples13, 2, mean)
pvij14 <- apply(vij_samples14, 2, mean)
pvij23 <- apply(vij_samples23, 2, mean)
pvij24 <- apply(vij_samples24, 2, mean)
pvij34 <- apply(vij_samples34, 2, mean)

# Estimated FDR
threshold12 = sort(pvij12, decreasing = TRUE)[T_edge1]
threshold13 = sort(pvij13, decreasing = TRUE)[T_edge1]
threshold14 = sort(pvij14, decreasing = TRUE)[T_edge1]
threshold23 = sort(pvij23, decreasing = TRUE)[T_edge1]
threshold24 = sort(pvij24, decreasing = TRUE)[T_edge1]
threshold34 = sort(pvij34, decreasing = TRUE)[T_edge1]

FDR_est12 <- rep(0, length(T_edge1))
FDR_est13 <- rep(0, length(T_edge1))
FDR_est14 <- rep(0, length(T_edge1))
FDR_est23 <- rep(0, length(T_edge1))
FDR_est24 <- rep(0, length(T_edge1))
FDR_est34 <- rep(0, length(T_edge1))

for(i in 1:length(threshold1)){
  th12 <- threshold12[i]
  est_diff12<- as.numeric(pvij12 >= th12)
  FDR_est12[i] <- sum((1-pvij12) * est_diff12)  / sum(est_diff12)
  
  th13 <- threshold13[i]
  est_diff13 <- as.numeric(pvij13 >= th13)
  FDR_est13[i] <- sum((1-pvij13) * est_diff13)  / sum(est_diff13)
  
  th14 <- threshold14[i]
  est_diff14 <- as.numeric(pvij14 >= th14)
  FDR_est14[i] <- sum((1-pvij14) * est_diff14)  / sum(est_diff14)
  
  th23 <- threshold23[i]
  est_diff23 <- as.numeric(pvij23 >= th23)
  FDR_est23[i] <- sum((1-pvij23) * est_diff23)  / sum(est_diff23)
  
  th24 <- threshold24[i]
  est_diff24 <- as.numeric(pvij24 >= th24)
  FDR_est24[i] <- sum((1-pvij24) * est_diff24)  / sum(est_diff24)
  
  th34 <- threshold34[i]
  est_diff34 <- as.numeric(pvij34 >= th34)
  FDR_est34[i] <- sum((1-pvij34) * est_diff34)  / sum(est_diff34)
}
```

Use the same threshold as above and identify cross-disease boundaries

Figure 9: Mutual cross-cancer difference boundaries (highlighted in red) detected by MDAGAR for each pair of cancers in SIR map when $\delta = 0.05$.
```{r, fig.height=7, fig.width=10}
alpha_n1 = 0.05
T12=sum(FDR_est12<=alpha_n1)
T13=sum(FDR_est13<=alpha_n1)
T14=sum(FDR_est14<=alpha_n1)
T23=sum(FDR_est23<=alpha_n1)
T24=sum(FDR_est24<=alpha_n1)
T34=sum(FDR_est34<=alpha_n1)

est_diff12 <- as.numeric(pvij12 >= threshold12[T12])
est_diff13 <- as.numeric(pvij13 >= threshold13[T13])
est_diff14 <- as.numeric(pvij14 >= threshold23[T14])
est_diff23 <- as.numeric(pvij23 >= threshold23[T23])
est_diff24 <- as.numeric(pvij24 >= threshold24[T24])
est_diff34 <- as.numeric(pvij34 >= threshold34[T34])

ch_edge12 <- pvij12[which(est_diff12 == 1)]
edge_plot12 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group),
               color = "black",
               fill = "white"
  )
for(i in which(est_diff12==1)){
  edge_plot12 = edge_plot12 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                        size = ((pvij12[i]- min(ch_edge12))/(max(ch_edge12) - min(ch_edge12)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=14, face="bold",hjust = 0.5)) +
    ggtitle(paste("Lung vs. Esophageal (T = ", T12, ")", sep=""))
}

ch_edge13 <- pvij13[which(est_diff13 == 1)]
edge_plot13 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group),
               color = "black",
               fill = "white"
  )
for(i in which(est_diff13==1)){
  edge_plot13 = edge_plot13 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                        size = ((pvij13[i]- min(ch_edge13))/(max(ch_edge13) - min(ch_edge13)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=14, face="bold",hjust = 0.5)) +
    ggtitle(paste("Lung vs. Layrnx (T = ", T13, ")", sep=""))
}

ch_edge14 <- pvij14[which(est_diff14 == 1)]
edge_plot14 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group),
               color = "black",
               fill = "white"
  )
for(i in which(est_diff14==1)){
  edge_plot14 = edge_plot14 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                        size = ((pvij14[i]- min(ch_edge14))/(max(ch_edge14) - min(ch_edge14)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=14, face="bold",hjust = 0.5)) +
    ggtitle(paste("Lung vs. Colorectal (T = ", T14, ")", sep=""))
}

ch_edge23 <- pvij23[which(est_diff23 == 1)]
edge_plot23 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group),
               color = "black",
               fill = "white"
  )
for(i in which(est_diff23==1)){
  edge_plot23 = edge_plot23 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                        size = ((pvij23[i]- min(ch_edge23))/(max(ch_edge23) - min(ch_edge23)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=14, face="bold",hjust = 0.5)) +
    ggtitle(paste("Esophageal vs. Layrnx (T = ", T23, ")", sep=""))
}

ch_edge24 <- pvij24[which(est_diff24 == 1)]
edge_plot24 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group),
               color = "black",
               fill = "white"
  )
for(i in which(est_diff24==1)){
  edge_plot24 = edge_plot24 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                        size = ((pvij24[i]- min(ch_edge24))/(max(ch_edge24) - min(ch_edge24)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=14, face="bold",hjust = 0.5)) +
    ggtitle(paste("Esophageal vs. Colorectal (T = ", T24, ")", sep=""))
}

ch_edge34 <- pvij34[which(est_diff34 == 1)]
edge_plot34 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group),
               color = "black",
               fill = "white"
  )
for(i in which(est_diff34==1)){
  edge_plot34 = edge_plot34 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                        size = ((pvij34[i]- min(ch_edge34))/(max(ch_edge34) - min(ch_edge34)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=14, face="bold",hjust = 0.5)) +
    ggtitle(paste("Larynx vs. Colorectal (T = ", T34, ")", sep=""))
}

ggarrange(edge_plot12, edge_plot13, edge_plot14, 
          edge_plot23, edge_plot24, edge_plot34, nrow = 2, ncol = 3)
```

###### VI. Model fitting: D score in Table 4
```{r}
Y_rep1 <- matrix(0, nrow = n, ncol = 10000)
Y_rep2 <- matrix(0, nrow = n, ncol = 10000)
Y_rep3 <- matrix(0, nrow = n, ncol = 10000)
Y_rep4 <- matrix(0, nrow = n, ncol = 10000)
for(i in 1:10000){
  
  beta1 <- mcmc_samples$beta[i,1]
  beta2 <- mcmc_samples$beta[i,2]
  beta3 <- mcmc_samples$beta[i,3]
  beta4 <- mcmc_samples$beta[i,4]
  
  W1_mcmc = mcmc_samples$phi[i,1:58]
  W2_mcmc = mcmc_samples$phi[i,59:116]
  W3_mcmc = mcmc_samples$phi[i,117:174]
  W4_mcmc = mcmc_samples$phi[i,175:232]
  
  Y_rep1[,i] <- rpois(n, E1 * exp(X1[,1] * beta1 +  W1_mcmc)) / E1
  Y_rep2[,i] <- rpois(n, E2 * exp(X2[,1] * beta2 +  W2_mcmc)) / E2
  Y_rep3[,i] <- rpois(n, E3 * exp(X3[,1] * beta3 +  W3_mcmc)) / E3
  Y_rep4[,i] <- rpois(n, E4 * exp(X4[,1] * beta4 +  W4_mcmc)) / E4
  
}
mu_rep1 = rowMeans(Y_rep1)
mu_rep2 = rowMeans(Y_rep2)
mu_rep3 = rowMeans(Y_rep3)
mu_rep4 = rowMeans(Y_rep4)

var_rep1 = rowVars(Y_rep1)
var_rep2 = rowVars(Y_rep2)
var_rep3 = rowVars(Y_rep3)
var_rep4 = rowVars(Y_rep4)

G.latent1 = sum((Y1/E1 - mu_rep1)^2)
P.latent1 = sum(var_rep1)
D.latent1 = G.latent1 + P.latent1

G.latent2 = sum((Y2/E2 - mu_rep2)^2)
P.latent2 = sum(var_rep2)
D.latent2 = G.latent2 + P.latent2

G.latent3 = sum((Y3/E3 - mu_rep3)^2)
P.latent3 = sum(var_rep3)
D.latent3 = G.latent3 + P.latent3

G.latent4 = sum((Y4/E4 - mu_rep4)^2)
P.latent4 = sum(var_rep4)
D.latent4 = G.latent4 + P.latent4

D_DAGAR = c(D.latent1, D.latent2, D.latent3, D.latent4)
D_DAGAR
sum(D_DAGAR)
```


##### 3.2.2 Multivariate CAR (MCAR) Model: Implement Metropolis within Gibbs Sampler for MCMC updating 
```{r}
# Matrix for precision matrix of CAR
D=matrix(0,n,n)
for(i in 1:n) D[i,i]=ni[i]

#######################################
### MCAR model with no covariates ###
#######################################

# Metropolis within Gibbs Sampler for MCMC updating 

ARDP_joint_diff_car<-function(y=Y, x1=as.matrix(X1[,1]), x2=as.matrix(X2[,1]), x3 = as.matrix(X3[,1]), x4 = as.matrix(X4[,1]),
                              X = X, E = E, Minc, alpha=1, q = 4, n.atoms=15, runs=10000, burn=1000){
  #y:       data
  #x:       covariates
  #n.atoms: number of atoms in the mixture dist.
  #theta:      the theta's (iid from baseline) in the model
  #alpha:     v~beta(1,alpha)
  #u:       the index indicator of spatial random effect
  #rho:     sptatial autocorrelation parameter in DAGAR
  #Minc:       0-1 adjacency matrix
  #V_r:     covariance matrix of joint MDAGAR
  #Q:       presicion matrix of DAGAR
  #r:       random effects following DAGAR
  #F_r:     Marginal CDF of r
  #taued:   presicion in Baseline of DP for disease d
  #taus:    precision for theta
  
  
  
  
  nq<-length(y)
  n <- nq / q
  p<-ncol(X)
  y1 <- y[1:n]
  y2 <- y[(n+1):(2*n)]
  y3 <- y[(2*n+1):(3*n)]
  y4 <- y[(3*n+1):(4*n)]
  
  E1 <- E[1:n]
  E2 <- E[(n+1):(2*n)]
  E3 <- E[(2*n+1):(3*n)]
  E4 <- E[(3*n+1):(4*n)]
  
  sigmasq_beta = 10000
  keepbeta=matrix(0,runs,p)
  keepphi=matrix(0,runs,nq)
  keeptheta=matrix(0,runs,n.atoms)
  keepA=matrix(0,runs,10)
  keepu=matrix(0,runs,nq)
  keeprho1=keeprho2=keeprho3=keeprho4=keeptaus = rep(0,runs)
  keepv=matrix(0,runs,n.atoms)
  keepr=matrix(0,runs,nq)
  keepFr=matrix(0,runs,nq)
  
  
  #initial values
  
  theta=rep(0,n.atoms)
  
  beta1<-rep(0,ncol(x1))
  beta2<-rep(0,ncol(x2))
  beta3<-rep(0,ncol(x3))
  beta4<-rep(0,ncol(x4))
  beta = c(beta1, beta2, beta3, beta4)
  
  taus = 1
  
  c=2
  d=0.1
  d2=1
  v<-rep(.1,n.atoms)
  probs=makeprobs(v)
  
  #A matrix
  rho=c(0.5, 0.5, 0.5, 0.5)
  eta = log(rho/(1-rho))
  A = matrix(0, q, q)
  for(i in 1:q){
    for(j in 1:i){
      if(j == i){
        A[i, j] = exp(rnorm(1))
      }else{
        A[i, j] = rnorm(1)
      }
    }
  }
  #Sigma = A%*%t(A)
  
  invQ1 = solve(D - rho[1] * Minc)
  invQ2 = solve(D - rho[2] * Minc)
  invQ3 = solve(D - rho[3] * Minc)
  invQ4 = solve(D - rho[4] * Minc)
  
  #f1 = rmvnorm(1, rep(0, n), invQ1)
  #f2 = rmvnorm(1, rep(0, n), invQ2)
  #f = c(f1, f2)
  kprod = kronecker(A, diag(n))
  invQ = as.matrix(bdiag(bdiag(invQ1, invQ2), bdiag(invQ3, invQ4)))
  #r = kprod %*% f
  
  Vr = as.matrix(forceSymmetric(kprod %*% invQ %*% t(kprod)))
  r = rmvnorm(1, rep(0, nq), Vr)
  
  
  F_r=pnorm(r,0,sqrt(diag(Vr)))
  u=makeu(F_r,probs)
  phi <- theta[u]
  
  nu = 2
  R = 0.1 * diag(q)
  
  
  acceptr=acceptv=acceptrho=acceptA=0
  acceptbeta1 = acceptbeta2=acceptbeta3=acceptbeta4 = 0
  accepttheta = 0
  
  #sumtp<-summu<-summu2<-rep(0,n)
  count<-afterburn<-0
  burn = burn + 1
  #iter = 1
  #runs = 1000
  for(iter in 1:runs){
    
    if(iter %% 100 == 0){
      print(iter)
      print(acceptA/(iter-1))
      print(acceptr/nq/(iter-1))
      print(acceptrho/(iter-1))
      print(acceptv/n.atoms/(iter-1))
      print(accepttheta/n.atoms/(iter-1))
      print(acceptbeta1/(iter-1))
      print(acceptbeta2/(iter-1))
      print(acceptbeta3/(iter-1))
      print(acceptbeta4/(iter-1))
    }
    ######################
    ###   update beta  ###
    ######################
    # update beta (intercept only model)
    
    pro_beta1=rnorm(1,beta1,0.02)
    MHrate1=sum(-E1*exp(pro_beta1*x1+theta[u][1:n])+y1*(pro_beta1*x1+theta[u][1:n]))-
      sum(-E1*exp(beta1*x1+theta[u][1:n])+y1*(beta1*x1+theta[u][1:n]))
    
    if(runif(1,0,1)<exp(MHrate1)){
      beta1<-pro_beta1
      acceptbeta1=acceptbeta1+1
    }
    
    pro_beta2=rnorm(1,beta2,0.05)
    MHrate2=sum(-E2*exp(pro_beta2*x2+theta[u][(n+1):(2*n)])+y2*(pro_beta2*x2+theta[u][(n+1):(2*n)]))-
      sum(-E2*exp(beta2*x2+theta[u][(n+1):(2*n)])+y2*(beta2*x2+theta[u][(n+1):(2*n)]))
    
    if(runif(1,0,1)<exp(MHrate2)){
      beta2<-pro_beta2
      acceptbeta2=acceptbeta2+1
    }
    
    pro_beta3=rnorm(1,beta3,0.05)
    MHrate3=sum(-E3*exp(pro_beta3*x3+theta[u][(2*n+1):(3*n)])+y3*(pro_beta3*x3+theta[u][(2*n+1):(3*n)]))-
      sum(-E3*exp(beta3*x3+theta[u][(2*n+1):(3*n)])+y3*(beta3*x3+theta[u][(2*n+1):(3*n)]))
    
    if(runif(1,0,1)<exp(MHrate3)){
      beta3<-pro_beta3
      acceptbeta3=acceptbeta3+1
    }
    
    pro_beta4=rnorm(1,beta4,0.02)
    MHrate4=sum(-E4*exp(pro_beta4*x4+theta[u][(3*n+1):(4*n)])+y4*(pro_beta4*x4+theta[u][(3*n+1):(4*n)]))-
      sum(-E4*exp(beta4*x4+theta[u][(3*n+1):(4*n)])+y4*(beta4*x4+theta[u][(3*n+1):(4*n)]))
    
    if(runif(1,0,1)<exp(MHrate4)){
      beta4<-pro_beta4
      acceptbeta4=acceptbeta4+1
    }
    
    beta <- c(beta1, beta2, beta3, beta4)
    
    
    #########################
    ###   update theta    ###
    #########################
    
    u1 = u[1:n]
    u2 = u[(n+1):(2*n)]
    u3 = u[(2*n+1):(3*n)]
    u4 = u[(3*n+1):(4*n)]
    
    
    
    for (j in 1:n.atoms){
      
      count[j]=sum(u==j)
      
      if (count[j]==0){
        theta[j]=rnorm(1,0,sqrt(1/taus))
      }else{
        pro_theta=rnorm(1,theta[j],0.06)
        
        MHrate<-sum(-(E[u==j])*exp(X[u==j,] %*% beta+pro_theta)+(y[u==j])*(X[u==j,] %*% beta+pro_theta))-taus/2*pro_theta^2-
          sum(-(E[u==j])*exp(X[u==j,] %*% beta+theta[j])+(y[u==j])*(X[u==j,] %*% beta+theta[j]))+taus/2*theta[j]^2
        if(runif(1,0,1)<exp(MHrate)){
          theta[j]<-pro_theta
          accepttheta=accepttheta+1
        }
      }
    }
    
    
    
    ######################
    ###   update r     ###
    ######################
    
    #pro_r=t(rmvnorm(1,r,sigma=covcar2,method="chol"))
    
    
    #tauvec = c(rep(taue1, n), rep(taue2, n), rep(taue3, n), rep(taue4, n))
    
    #pro_r = rmvnorm(1, r, 0.01*diag(nq))
    
    for (k in 1:nq){
      pro_r=r;pro_Fr=F_r;pro_u=u
      pro_r[k]=rnorm(1,r[k],1.7)
      pro_Fr[k]=pnorm(pro_r[k],0,sqrt(Vr[k,k]))
      pro_u[k]=makeu(pro_Fr[k],probs)
      
      MH=dmvnorm(pro_r,mean=rep(0, nq),sigma=Vr,log=T)+dpois(y[k],E[k]*exp(as.numeric(X[k,]%*%beta)+theta[pro_u[k]]),log=T)-
        dmvnorm(r,mean=rep(0, nq),sigma=Vr,log=T)-dpois(y[k],E[k]*exp(as.numeric(X[k,]%*%beta)+theta[u[k]]),log=T)
      #MH=dmvnorm(pro_r,mean=rep(0, nq),sigma=Vr,log=T)+dmvnorm(t(as.matrix(y1)),mean=as.vector(x1%*%t(beta1))+theta[pro_u[1:n]],sigma=1/taue1*diag(n),log=T)+
      # dmvnorm(t(as.matrix(y2)),mean=as.vector(x2%*%t(beta2))+theta[pro_u[(n+1):(2*n)]],sigma=1/taue2*diag(n),log=T)-
      #dmvnorm(r,mean=rep(0, nq),sigma=Vr,log=T)-dmvnorm(t(as.matrix(y1)),mean=as.vector(x1%*%t(beta1))+theta[u[1:n]],sigma=1/taue1*diag(n),log=T)-
      #dmvnorm(t(as.matrix(y2)),mean=as.vector(x2%*%t(beta2))+theta[u[(n+1):(2*n)]],sigma=1/taue2*diag(n),log=T)
      if(runif(1,0,1)<exp(MH)){
        r[k]=pro_r[k]
        F_r[k]=pro_Fr[k]
        u[k]=pro_u[k]
        acceptr=acceptr+1
      }
    }
    
    
    ######################
    ###   update   v   ###
    ######################
    
    
    #pro_v=v
    for (j in 1:n.atoms){
      #0.9
      pro_v=v
      pro_v[j]<-rnorm(1,v[j],0.06)
      while (pro_v[j]<0 || pro_v[j]>1) {pro_v[j]<-rnorm(1,v[j],0.06)}
      #if(pro_v[j]>0 & pro_v[j]<1){
      pro_probs=makeprobs(pro_v)
      pro_u=makeu(F_r,pro_probs)
      
      MH=log(dbeta(pro_v[j],1,alpha))+sum(dpois(y,E*exp(X%*%beta+theta[pro_u]),log=T))-
        log(dbeta(v[j],1,alpha))-sum(dpois(y,E*exp(X%*%beta+theta[u]),log=T))
      
      
      if(runif(1,0,1)<exp(MH)){
        v[j]=pro_v[j]
        probs=pro_probs
        u=pro_u
        acceptv=acceptv+1
      }
      # }
    }
    
    
    ######################
    ###   update taus  ###
    ######################
    
    
    taus=rgamma(1,shape=1/2*n.atoms+c,rate=sum(theta^2)/2+d2)
    
    
    ######################
    ###   update rho  ###
    ######################
    
    
    pro_eta1 = rnorm(1, eta[1], 1.1)
    pro_eta2 = rnorm(1, eta[2], 1.1)
    pro_eta3 = rnorm(1, eta[3], 1.1)
    pro_eta4 = rnorm(1, eta[4], 1.1)
    pro_eta = c(pro_eta1, pro_eta2, pro_eta3, pro_eta4)
    pro_rho = exp(pro_eta)/(1+exp(pro_eta))
    
    pro_invQ1 = solve(D - pro_rho[1] * Minc)
    pro_invQ2 = solve(D - pro_rho[2] * Minc)
    pro_invQ3 = solve(D - pro_rho[3] * Minc)
    pro_invQ4 = solve(D - pro_rho[4] * Minc)
    
    kprod = kronecker(A, diag(n))
    pro_invQ = as.matrix(bdiag(bdiag(pro_invQ1, pro_invQ2), bdiag(pro_invQ3, pro_invQ4)))
    pro_Vr = kprod %*% pro_invQ %*% t(kprod)
    
    MH = dmvnorm(r,mean=rep(0, nq),sigma=as.matrix(forceSymmetric(pro_Vr)),log=T) +
      log(pro_rho[1]) + log(1-pro_rho[1]) + log(pro_rho[2]) + log(1-pro_rho[2]) +
      log(pro_rho[3]) + log(1-pro_rho[3]) + log(pro_rho[4]) + log(1-pro_rho[4]) -
      dmvnorm(r,mean=rep(0, nq),sigma=Vr,log=T) - log(rho[1]) - log(1-rho[1]) - log(rho[2]) - log(1-rho[2]) -
      log(rho[3]) - log(1-rho[3]) - log(rho[4]) - log(1-rho[4])
    
    if(runif(1,0,1)<exp(MH)){
      eta = pro_eta
      rho = pro_rho
      Vr = as.matrix(forceSymmetric(pro_Vr))
      acceptrho=acceptrho+1
    }
    
    ######################
    ###   update A  ###
    ######################
    #0.06
    pro_A = matrix(0, q, q)
    for(i in 1:q){
      for(j in 1:i){
        if(j == i){
          pro_A[i, j] = exp(rnorm(1, log(A[i, j]), 0.05))
          #pro_A[i, j] = rlnorm(1, log(A[i, j]), 0.1)
        }else{
          pro_A[i, j] = rnorm(1, A[i, j], 0.05)
        }
      }
    }
    
    invQ1 = solve(D - rho[1] * Minc)
    invQ2 = solve(D - rho[2] * Minc)
    invQ3 = solve(D - rho[3] * Minc)
    invQ4 = solve(D - rho[4] * Minc)
    
    Sigma =  A %*% t(A)
    
    pro_kprod = kronecker(pro_A, diag(n))
    invQ = as.matrix(bdiag(bdiag(invQ1, invQ2), bdiag(invQ3, invQ4)))
    pro_Vr = pro_kprod %*% as.matrix(forceSymmetric(invQ)) %*% t(pro_kprod)
    pro_Sigma =  pro_A %*% t(pro_A)
    
    lpA = -(nu+4)/2*logdet(Sigma) - 1/2*sum(diag(nu*R%*%solve(Sigma))) + log(Jacob_A(A)) + sum(log(diag(A)))
    pro_lpA = -(nu+4)/2*logdet(pro_Sigma) - 1/2*sum(diag(nu*R%*%solve(pro_Sigma))) + log(Jacob_A(pro_A)) + sum(log(diag(pro_A)))
    MH = dmvnorm(r,mean=rep(0, nq),sigma=as.matrix(forceSymmetric(pro_Vr)),log=T) + pro_lpA -
      dmvnorm(r,mean=rep(0, nq),sigma=Vr,log=T) - lpA
    
    if(runif(1,0,1)<exp(MH)){
      A = pro_A
      Vr = as.matrix(forceSymmetric(pro_Vr))
      acceptA=acceptA+1
    }
    
    #kprod = kronecker(A, diag(n))
    #Vr = as.matrix(forceSymmetric(kprod %*% invQ %*% t(kprod)))
    
    ######################
    ### record samples ###
    ######################
    
    keeptheta[iter,] = theta
    keepphi[iter,] = theta[u]
    
    keepA[iter,] = as.vector(A)[-c(5,9,10,13:15)]
    
    keepbeta[iter,]=beta
    keeptaus[iter]=taus
    keeprho1[iter]=rho[1]
    keeprho2[iter]=rho[2]
    keeprho3[iter]=rho[3]
    keeprho4[iter]=rho[4]
    keepu[iter,]=u
    keepv[iter,]=v
    keepr[iter,]=r
    keepFr[iter,]=F_r
    
    #    cat("iteration = ", i, "acceptance rate of r = ", acceptr/(n*i),"acceptance rate of v = ",acceptv/(n.atoms*i), "\n")
    
  }
  
  list(beta=keepbeta[burn:runs,], A = keepA[burn:runs,],phi=keepphi[burn:runs,],theta=keeptheta[burn:runs,],u=keepu[burn:runs,],
       v=keepv[burn:runs,],r=keepr[burn:runs,],Fr=keepFr[burn:runs,], taus=keeptaus[burn:runs],
       rho1=keeprho1[burn:runs], rho2=keeprho2[burn:runs], rho3=keeprho3[burn:runs], rho4=keeprho4[burn:runs])
}
set.seed(123)
mcmc_samples=ARDP_joint_diff_car(y=Y, x1=as.matrix(X1[,1]), x2=as.matrix(X2[,1]), x3 = as.matrix(X3[,1]), x4 = as.matrix(X4[,1]), 
                                 X = X, E=E, Minc, alpha=1, q = 4, n.atoms=15, runs=30000, burn=20000)
#Estimates of parameters
sample.mcmc = cbind(mcmc_samples$taus, mcmc_samples$rho1, mcmc_samples$rho2, mcmc_samples$rho3, mcmc_samples$rho4)
```
###### I. Obtain spatial randome effects in original spatial order
```{r}
phis <- mcmc_samples$phi
phis1 <- phis[,1:58][,order(final_perm)]
phis2 <- phis[,59:116][,order(final_perm)]
phis3 <- phis[,117:174][,order(final_perm)]
phis4 <- phis[,175:232][,order(final_perm)]
phis_origin <- cbind(phis1, phis2, phis3, phis4)
```

###### II. Difference boundaries for each cancer individually
We calculate posterior probabilities: $P(\phi_{id} \neq \phi_{jd} | y), \quad i \sim j$
```{r}
vij_samples1 <- t(apply(phis1, 1, function(x){
  diff_b1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]]){
      diff_b1 <- c(diff_b1, 1)
    }else{
      diff_b1 <- c(diff_b1, 0)
    }
  }
  return(diff_b1)
}))

vij_samples2 <- t(apply(phis2, 1, function(x){
  diff_b1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]]){
      diff_b1 <- c(diff_b1, 1)
    }else{
      diff_b1 <- c(diff_b1, 0)
    }
  }
  return(diff_b1)
}))

vij_samples3 <- t(apply(phis3, 1, function(x){
  diff_b1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]]){
      diff_b1 <- c(diff_b1, 1)
    }else{
      diff_b1 <- c(diff_b1, 0)
    }
  }
  return(diff_b1)
}))

vij_samples4 <- t(apply(phis4, 1, function(x){
  diff_b1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]]){
      diff_b1 <- c(diff_b1, 1)
    }else{
      diff_b1 <- c(diff_b1, 0)
    }
  }
  return(diff_b1)
}))

# probabilities: P(phi_id != phi_jd), i~j
pvij1 <- apply(vij_samples1, 2, mean)
pvij2 <- apply(vij_samples2, 2, mean)
pvij3 <- apply(vij_samples3, 2, mean)
pvij4 <- apply(vij_samples4, 2, mean)

names(pvij1) <- neighbor_name
names(pvij2) <- neighbor_name
names(pvij3) <- neighbor_name
names(pvij4) <- neighbor_name

```

####### a. Figure 6 (b): Estimated FDR curves plotted against the number of selected difference boundaries for four cancers using MCAR
```{r}
T_edge1 = seq(0, 135, 1)[-1]
threshold1 = sort(pvij1, decreasing = TRUE)[T_edge1]
threshold2 = sort(pvij2, decreasing = TRUE)[T_edge1]
threshold3 = sort(pvij3, decreasing = TRUE)[T_edge1]
threshold4 = sort(pvij4, decreasing = TRUE)[T_edge1]
FDR_est1 <- rep(0, length(T_edge1))
FDR_est2 <- rep(0, length(T_edge1))
FDR_est3 <- rep(0, length(T_edge1))
FDR_est4 <- rep(0, length(T_edge1))


for(i in 1:length(threshold1)){
  th1 <- threshold1[i]
  est_diff1 <- as.numeric(pvij1 >= th1)
  FDR_est1[i] <- sum((1-pvij1) * est_diff1)  / sum(est_diff1)
  
  th2 <- threshold2[i]
  est_diff2 <- as.numeric(pvij2 >= th2)
  FDR_est2[i] <- sum((1-pvij2) * est_diff2)  / sum(est_diff2)
  
  th3 <- threshold3[i]
  est_diff3 <- as.numeric(pvij3 >= th3)
  FDR_est3[i] <- sum((1-pvij3) * est_diff3)  / sum(est_diff3)
  
  th4 <- threshold4[i]
  est_diff4 <- as.numeric(pvij4 >= th4)
  FDR_est4[i] <- sum((1-pvij4) * est_diff4)  / sum(est_diff4)
}

plot(FDR_est1,type="l", lty = 1, xlab="Number of edges selected", ylab = "Estimated FDR", ylim=c(0,0.5))
lines(FDR_est2,type="l", lty = 2)
lines(FDR_est3,type="l", lty = 3)
lines(FDR_est4,type="l", lty = 4)
legend("topleft", legend=c("Lung", "Esophageal", "Layrnx", "Colorectal"),
       lty=1:4, cex=0.7)
```

####### b. Names of adjacent counties that have significant boundary effects from the MCAR model for each cancer when $\delta = 0.05$
```{r}
# Set threshold for FDR: 0.025
alpha_n = 0.05
T1 = sum(FDR_est1<=alpha_n)
T2 = sum(FDR_est2<=alpha_n)
T3 = sum(FDR_est3<=alpha_n)
T4 = sum(FDR_est4<=alpha_n)

name_diff1 <- names(pvij1[order(pvij1,decreasing = T)][1:T1])
name_diff2 <- names(pvij2[order(pvij2,decreasing = T)][1:T2])
name_diff3 <- names(pvij3[order(pvij3,decreasing = T)][1:T3])
name_diff4 <- names(pvij4[order(pvij4,decreasing = T)][1:T4])
```

####### c. Difference boundary detection
By setting FDR threshold the same as MDAGAR ($0.05$), we get difference boundaries detected for each cancer individually.

Figure 7 (b): Difference boundaries (highlighted in red) detected by MCAR in SIR map for four cancers individually when $\delta = 0.05$.
```{r, fig.height=7, fig.width=10}
# Difference boundary for each cancer with FDR <= 0.05
est_diff1 <- as.numeric(pvij1 >= threshold1[T1])

est_diff2 <- as.numeric(pvij2 >= threshold2[T2])

est_diff3 <- as.numeric(pvij3 >= threshold3[T3])

est_diff4 <- as.numeric(pvij4 >= threshold4[T4])

# Lung cancer
ch_edge1 <- pvij1[which(est_diff1 == 1)]
edge_plot1 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group, fill = lung_SIR),
               color = "black",
               alpha = 0.6
  ) +
  scale_fill_stepsn(colors = color,
                    breaks = breaks1)

for(i in which(est_diff1 == 1)){
  edge_plot1 = edge_plot1 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                      size = ((pvij1[i]- min(ch_edge1))/(max(ch_edge1) - min(ch_edge1)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=16, face="bold",hjust = 0.5),
          legend.title = element_blank()) +
    ggtitle(paste("Lung (T = ", T1, ")", sep=""))
}

# Esophageal cancer
ch_edge2 <- pvij2[which(est_diff2 == 1)]
edge_plot2 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group, fill = esophagus_SIR),
               color = "black",
               alpha = 0.6
  ) +
  scale_fill_stepsn(colors = color,
                    breaks = breaks2)
for(i in which(est_diff2 == 1)){
  edge_plot2 = edge_plot2 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                      size = ((pvij2[i]- min(ch_edge2))/(max(ch_edge2) - min(ch_edge2)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=16, face="bold",hjust = 0.5),
          legend.title = element_blank()) +
    ggtitle(paste("Esophageal (T = ", T2, ")", sep=""))
}

# Larynx cancer
ch_edge3 <- pvij3[which(est_diff3 == 1)]
edge_plot3 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group, fill = larynx_SIR),
               color = "black",
               alpha = 0.6
  ) +
  scale_fill_stepsn(colors = color,
                    breaks = breaks3)
for(i in which(est_diff3 == 1)){
  edge_plot3 = edge_plot3 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                      size = ((pvij3[i]- min(ch_edge3))/(max(ch_edge3) - min(ch_edge3)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=16, face="bold",hjust = 0.5),
          legend.title = element_blank()) +
    ggtitle(paste("Larynx (T = ", T3, ")", sep=""))
}

# Colorectal cancer
ch_edge4 <- pvij4[which(est_diff4 == 1)]
edge_plot4 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group, fill = colrect_SIR),
               color = "black",
               alpha = 0.6
  ) +
  scale_fill_stepsn(colors = color,
                    breaks = breaks4)
for(i in which(est_diff4 == 1)){
  edge_plot4 = edge_plot4 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                      size = ((pvij4[i]- min(ch_edge4))/(max(ch_edge4) - min(ch_edge4)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=16, face="bold",hjust = 0.5),
          legend.title = element_blank()) +
    ggtitle(paste("Colorectal (T = ", T4, ")", sep=""))
}

ggarrange(edge_plot1, edge_plot2, edge_plot3, edge_plot4, nrow = 2, ncol = 2) 


```

###### III. Model fitting: D score in Table 4
```{r}
Y_rep1 <- matrix(0, nrow = n, ncol = 10000)
Y_rep2 <- matrix(0, nrow = n, ncol = 10000)
Y_rep3 <- matrix(0, nrow = n, ncol = 10000)
Y_rep4 <- matrix(0, nrow = n, ncol = 10000)
for(i in 1:10000){
  
  beta1 <- mcmc_samples$beta[i,1]
  beta2 <- mcmc_samples$beta[i,2]
  beta3 <- mcmc_samples$beta[i,3]
  beta4 <- mcmc_samples$beta[i,4]
  
  W1_mcmc = mcmc_samples$phi[i,1:58]
  W2_mcmc = mcmc_samples$phi[i,59:116]
  W3_mcmc = mcmc_samples$phi[i,117:174]
  W4_mcmc = mcmc_samples$phi[i,175:232]
  
  Y_rep1[,i] <- rpois(n, E1 * exp(X1[,1] * beta1 +  W1_mcmc)) / E1
  Y_rep2[,i] <- rpois(n, E2 * exp(X2[,1] * beta2 +  W2_mcmc)) / E2
  Y_rep3[,i] <- rpois(n, E3 * exp(X3[,1] * beta3 +  W3_mcmc)) / E3
  Y_rep4[,i] <- rpois(n, E4 * exp(X4[,1] * beta4 +  W4_mcmc)) / E4
  
}
mu_rep1 = rowMeans(Y_rep1)
mu_rep2 = rowMeans(Y_rep2)
mu_rep3 = rowMeans(Y_rep3)
mu_rep4 = rowMeans(Y_rep4)

var_rep1 = rowVars(Y_rep1)
var_rep2 = rowVars(Y_rep2)
var_rep3 = rowVars(Y_rep3)
var_rep4 = rowVars(Y_rep4)

G.latent1 = sum((Y1/E1 - mu_rep1)^2)
P.latent1 = sum(var_rep1)
D.latent1 = G.latent1 + P.latent1

G.latent2 = sum((Y2/E2 - mu_rep2)^2)
P.latent2 = sum(var_rep2)
D.latent2 = G.latent2 + P.latent2

G.latent3 = sum((Y3/E3 - mu_rep3)^2)
P.latent3 = sum(var_rep3)
D.latent3 = G.latent3 + P.latent3

G.latent4 = sum((Y4/E4 - mu_rep4)^2)
P.latent4 = sum(var_rep4)
D.latent4 = G.latent4 + P.latent4

D_CAR = c(D.latent1, D.latent2, D.latent3, D.latent4)
D_CAR
sum(D_CAR)
```

You can also apply other analysis as MDAGAR here for more details.

#### 3.3 Independent-disease Models implemented in rjags

##### 3.3.1 DAGAR$_{ind}$ Model
```
sink("MDAGAR_ARDP_ind.txt")
cat("
    model
    {
    
    for (i in 1:k)
    {
    log(mu[i]) <- log(E[i]) + X[i,] %*% beta + phi[i]
    Y[i] ~ dpois(mu[i])
    phi[i] <- inprod(theta, u[i,])
    Fr[i] <- pnorm(r[i], 0, 1/invQ1[i,i])
    for(h in 2:H){
    u[i, h] <- ifelse(Fr[i] > sum_pi[h-1] && Fr[i] < sum_pi[h], 1, 0)
    }
    u[i,1] <- ifelse(sum(u[i, 2:H])==0, 1, 0)
    }
    
    for (i in (k+1):(2*k))
    {
    log(mu[i]) <- log(E[i]) + X[i,] %*% beta + phi[i]
    Y[i] ~ dpois(mu[i])
    phi[i] <- inprod(theta, u[i,])
    Fr[i] <- pnorm(r[i], 0, 1/invQ2[i-k,i-k])
    for(h in 2:H){
    u[i, h] <- ifelse(Fr[i] > sum_pi[h-1] && Fr[i] < sum_pi[h], 1, 0)
    }
    u[i,1] <- ifelse(sum(u[i, 2:H])==0, 1, 0)
    }
    
    for (i in (2*k+1):(3*k))
    {
    log(mu[i]) <- log(E[i]) + X[i,] %*% beta + phi[i]
    Y[i] ~ dpois(mu[i])
    phi[i] <- inprod(theta, u[i,])
    Fr[i] <- pnorm(r[i], 0, 1/invQ3[i-2*k,i-2*k])
    for(h in 2:H){
    u[i, h] <- ifelse(Fr[i] > sum_pi[h-1] && Fr[i] < sum_pi[h], 1, 0)
    }
    u[i,1] <- ifelse(sum(u[i, 2:H])==0, 1, 0)
    }
    
    for (i in (3*k+1):(4*k))
    {
    log(mu[i]) <- log(E[i]) + X[i,] %*% beta + phi[i]
    Y[i] ~ dpois(mu[i])
    phi[i] <- inprod(theta, u[i,])
    Fr[i] <- pnorm(r[i], 0, 1/invQ4[i-3*k,i-3*k])
    for(h in 2:H){
    u[i, h] <- ifelse(Fr[i] > sum_pi[h-1] && Fr[i] < sum_pi[h], 1, 0)
    }
    u[i,1] <- ifelse(sum(u[i, 2:H])==0, 1, 0)
    }
    
    
    pi[1] <- V[1]
    prod[1] <- 1 - pi[1]
    for (h in 2:H){
    prod[h] <- prod[h-1] * (1-V[h])
    pi[h] <- V[h] * prod[h-1]
    }
    
    for (h in 1:H){
    theta[h] ~ dnorm(0,taus)
    V[h] ~ dbeta(1,alpha)
    sum_pi[h] <- sum(pi[1:h])
    }
    
   for(i in 1:2){
       Tau1[i] <- 1
       Tau2[i] <- 1
       Tau3[i] <- 1
       Tau4[i] <- 1
       r1[i] ~ dnorm(0, tau1*Tau1[i])
       r2[i] ~ dnorm(0, tau2*Tau2[i])
       r3[i] ~ dnorm(0, tau3*Tau3[i])
       r4[i] ~ dnorm(0, tau4*Tau4[i])
   }
   for (i in 3:k){
       Tau1[i] <- (1 + (ns[i-1] - 1) * rho1^2) / (1 - rho1^2)
       Tau2[i] <- (1 + (ns[i-1] - 1) * rho2^2) / (1 - rho2^2)
       Tau3[i] <- (1 + (ns[i-1] - 1) * rho3^2) / (1 - rho3^2)
       Tau4[i] <- (1 + (ns[i-1] - 1) * rho4^2) / (1 - rho4^2)

       b1[i] <- rho1 / (1 + (ns[i-1] - 1) * rho1^2)
       b2[i] <- rho2 / (1 + (ns[i-1] - 1) * rho2^2)
       b3[i] <- rho3 / (1 + (ns[i-1] - 1) * rho3^2)
       b4[i] <- rho4 / (1 + (ns[i-1] - 1) * rho4^2)
      
      for(j in 1:ns[i-1]){
         t1[i,j] <- r1[udnei[(cn[i-1] + j)]] * b1[i]
         t2[i,j] <- r2[udnei[(cn[i-1] + j)]] * b2[i]
         t3[i,j] <- r3[udnei[(cn[i-1] + j)]] * b3[i]
         t4[i,j] <- r4[udnei[(cn[i-1] + j)]] * b4[i]
      }
      for(j in (ns[i-1]+1):mns){
        t1[i,j] <- 0
        t2[i,j] <- 0
        t3[i,j] <- 0
        t4[i,j] <- 0
      }
      r1[i] ~ dnorm(sum(t1[i,]), tau1*Tau1[i])
      r2[i] ~ dnorm(sum(t2[i,]), tau2*Tau2[i])
      r3[i] ~ dnorm(sum(t3[i,]), tau3*Tau3[i])
      r4[i] ~ dnorm(sum(t4[i,]), tau4*Tau4[i])
   }
    
   for(i in 1:2){
      Taum1[i, i] <- 1
      Taum2[i, i] <- 1
      Taum3[i, i] <- 1
      Taum4[i, i] <- 1
    
      for(j in 1:k){
         B1[i,j] <- 0
         B2[i,j] <- 0
         B3[i,j] <- 0
         B4[i,j] <- 0
      }
    }
    for(l in 1:(k-1)){
       for(m in (l+1):k){
          Taum1[l,m] <- 0
          Taum2[l,m] <- 0
          Taum3[l,m] <- 0
          Taum4[l,m] <- 0
        }
    }

    Taum1[2,1] <- 0
    Taum2[2,1] <- 0
    Taum3[2,1] <- 0
    Taum4[2,1] <- 0
    for (i in 3:k) {
        Taum1[i,i] <- Tau1[i]
        Taum2[i,i] <- Tau2[i]
        Taum3[i,i] <- Tau3[i]
        Taum4[i,i] <- Tau4[i]
   
      for(m in 1:(i-1)){
        Taum1[i,m] <- 0
        Taum2[i,m] <- 0
        Taum3[i,m] <- 0
        Taum4[i,m] <- 0
       }

      for(j in (udnei[(cn[i-1] + 1):(cn[i-1] + ns[i-1])])){
        B1[i,j] <- b1[i]
        B2[i,j] <- b2[i]
        B3[i,j] <- b3[i]
        B4[i,j] <- b4[i]
      }
    
     for(j in index1[((k)*(i-3)-cn[i-1]+1) : ((k)*(i-3)-cn[i-1] + (k - ns[i-1]))]){
        B1[i,j] <- 0
        B2[i,j] <- 0
        B3[i,j] <- 0
        B4[i,j] <- 0
     }
  }
     
    Q1 <- tau1 * t(Ik - B1) %*% Taum1 %*% (Ik - B1)
    Q2 <- tau2 * t(Ik - B2) %*% Taum2 %*% (Ik - B2)
    Q3 <- tau3 * t(Ik - B3) %*% Taum3 %*% (Ik - B3)
    Q4 <- tau4 * t(Ik - B4) %*% Taum4 %*% (Ik - B4)
    r[1:k] <- r1[1:k]
    r[(k+1):(2*k)] <- r2[1:k]
    r[(2*k+1):(3*k)] <- r3[1:k]
    r[(3*k+1):(4*k)] <- r4[1:k]
    
    invQ1 <- inverse(Q1)
    invQ2 <- inverse(Q2)
    invQ3 <- inverse(Q3)
    invQ4 <- inverse(Q4)
    
    rho1 ~ dunif(0, 0.98)
    rho2 ~ dunif(0, 0.98)
    rho3 ~ dunif(0, 0.98)
    rho4 ~ dunif(0, 0.98)
    tau1 ~ dgamma(2, 0.1)
    tau2 ~ dgamma(2, 0.1)
    tau3 ~ dgamma(2, 0.1)
    tau4 ~ dgamma(2, 0.1)
    
    taus ~ dgamma(2,1)
    beta[1:ncolumn] ~ dmnorm(rep(0,ncolumn), (0.0001*I));
    }
    ", fill = TRUE)
sink()

```

```{r eval=FALSE}
model.data1 <- list(k = n,  I = diag(ncol(X)), X = X, Y = Y, E = E, Ik = diag(n), index1 = index1,
                    alpha = 1, ncolumn = ncol(X),  H = 15,  ns = dni, udnei = udnei, cn = c(0, cni))
model.inits <- rep(list(list(rho1 = 0.1, rho2 = 0.1, rho3 = 0.1, rho4 = 0.1, tau1 = 1, tau2 = 1, tau3 = 1, tau4 = 1,  
                             taus = 1, beta = rep(0, ncol(X)))),2)
model.param <- c("beta", "rho1", "rho2", "rho3", "rho4", "tau1", "tau2", "tau3", "tau4", 
                 "taus", "phi", "u")
set.seed(123)
mcmc_samples <- jags(model.data1, model.inits, model.param, "MDAGAR_ARDP_ind.txt",
                   n.chains = 2, n.iter = 25000,n.burnin = 20000, n.thin = 1)
# Parameter estimates
sample.mcmc = mcmc_samples$BUGSoutput$sims.matrix[,c(246, 238:245)]
```

###### I. Obtain spatial randome effects in original spatial order
```{r eval=FALSE}
phis <- mcmc_samples$BUGSoutput$sims.matrix[,6:237]
phis1 <- phis[,1:58][,order(final_perm)]
phis2 <- phis[,59:116][,order(final_perm)]
phis3 <- phis[,117:174][,order(final_perm)]
phis4 <- phis[,175:232][,order(final_perm)]
phis_origin <- cbind(phis1, phis2, phis3, phis4)
```

###### II. Difference boundaries for each cancer individually
We calculate posterior probabilities: $P(\phi_{id} \neq \phi_{jd}), \quad i \sim j$
```{r eval=FALSE}
vij_samples1 <- t(apply(phis1, 1, function(x){
  diff_b1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]]){
      diff_b1 <- c(diff_b1, 1)
    }else{
      diff_b1 <- c(diff_b1, 0)
    }
  }
  return(diff_b1)
}))

vij_samples2 <- t(apply(phis2, 1, function(x){
  diff_b1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]]){
      diff_b1 <- c(diff_b1, 1)
    }else{
      diff_b1 <- c(diff_b1, 0)
    }
  }
  return(diff_b1)
}))

vij_samples3 <- t(apply(phis3, 1, function(x){
  diff_b1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]]){
      diff_b1 <- c(diff_b1, 1)
    }else{
      diff_b1 <- c(diff_b1, 0)
    }
  }
  return(diff_b1)
}))

vij_samples4 <- t(apply(phis4, 1, function(x){
  diff_b1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]]){
      diff_b1 <- c(diff_b1, 1)
    }else{
      diff_b1 <- c(diff_b1, 0)
    }
  }
  return(diff_b1)
}))

# probabilities: P(phi_id != phi_jd), i~j
pvij1 <- apply(vij_samples1, 2, mean)
pvij2 <- apply(vij_samples2, 2, mean)
pvij3 <- apply(vij_samples3, 2, mean)
pvij4 <- apply(vij_samples4, 2, mean)

names(pvij1) <- neighbor_name
names(pvij2) <- neighbor_name
names(pvij3) <- neighbor_name
names(pvij4) <- neighbor_name

```

####### a. Estimated FDR curves
```{r eval=FALSE}
T_edge1 = seq(0, 135, 1)[-1]
threshold1 = sort(pvij1, decreasing = TRUE)[T_edge1]
threshold2 = sort(pvij2, decreasing = TRUE)[T_edge1]
threshold3 = sort(pvij3, decreasing = TRUE)[T_edge1]
threshold4 = sort(pvij4, decreasing = TRUE)[T_edge1]
FDR_est1 <- rep(0, length(T_edge1))
FDR_est2 <- rep(0, length(T_edge1))
FDR_est3 <- rep(0, length(T_edge1))
FDR_est4 <- rep(0, length(T_edge1))


for(i in 1:length(threshold1)){
  th1 <- threshold1[i]
  est_diff1 <- as.numeric(pvij1 >= th1)
  FDR_est1[i] <- sum((1-pvij1) * est_diff1)  / sum(est_diff1)
  
  th2 <- threshold2[i]
  est_diff2 <- as.numeric(pvij2 >= th2)
  FDR_est2[i] <- sum((1-pvij2) * est_diff2)  / sum(est_diff2)
  
  th3 <- threshold3[i]
  est_diff3 <- as.numeric(pvij3 >= th3)
  FDR_est3[i] <- sum((1-pvij3) * est_diff3)  / sum(est_diff3)
  
  th4 <- threshold4[i]
  est_diff4 <- as.numeric(pvij4 >= th4)
  FDR_est4[i] <- sum((1-pvij4) * est_diff4)  / sum(est_diff4)
}

plot(FDR_est1,type="l", lty = 1, xlab="Number of edges selected", ylab = "Estimated FDR", ylim=c(0,0.5))
lines(FDR_est2,type="l", lty = 2)
lines(FDR_est3,type="l", lty = 3)
lines(FDR_est4,type="l", lty = 4)
legend("topleft", legend=c("Lung", "Esophageal", "Layrnx", "Colorectal"),
       lty=1:4, cex=0.7)
```

####### b. Difference boundary detection
By setting FDR threshold as $0.1$, we get difference boundaries detected for each cancer individually.
```{r}
# Set threshold for FDR: 0.1
alpha_n = 0.1
T1 = sum(FDR_est1<=alpha_n)
T2 = sum(FDR_est2<=alpha_n)
T3 = sum(FDR_est3<=alpha_n)
T4 = sum(FDR_est4<=alpha_n)

# Difference boundary for each cancer with FDR <= 0.025
est_diff1 <- as.numeric(pvij1 >= threshold1[T1])

est_diff2 <- as.numeric(pvij2 >= threshold2[T2])

est_diff3 <- as.numeric(pvij3 >= threshold3[T3])

est_diff4 <- as.numeric(pvij4 >= threshold4[T4])

# Lung cancer
ch_edge1 <- pvij1[which(est_diff1 == 1)]
edge_plot1 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group, fill = lung_SIR),
               color = "black",
               alpha = 0.6
  ) +
  scale_fill_stepsn(colors = color,
                    breaks = breaks1)

for(i in which(est_diff1 == 1)){
  edge_plot1 = edge_plot1 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                      size = ((pvij1[i]- min(ch_edge1))/(max(ch_edge1) - min(ch_edge1)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=16, face="bold",hjust = 0.5),
          legend.title = element_blank()) +
    ggtitle(paste("Lung (T = ", T1, ")", sep=""))
}

# Esophageal cancer
ch_edge2 <- pvij2[which(est_diff2 == 1)]
edge_plot2 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group, fill = esophagus_SIR),
               color = "black",
               alpha = 0.6
  ) +
  scale_fill_stepsn(colors = color,
                    breaks = breaks2)
for(i in which(est_diff2 == 1)){
  edge_plot2 = edge_plot2 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                      size = ((pvij2[i]- min(ch_edge2))/(max(ch_edge2) - min(ch_edge2)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=16, face="bold",hjust = 0.5),
          legend.title = element_blank()) +
    ggtitle(paste("Esophageal (T = ", T2, ")", sep=""))
}

# Larynx cancer
ch_edge3 <- pvij3[which(est_diff3 == 1)]
edge_plot3 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group, fill = larynx_SIR),
               color = "black",
               alpha = 0.6
  ) +
  scale_fill_stepsn(colors = color,
                    breaks = breaks3)
for(i in which(est_diff3 == 1)){
  edge_plot3 = edge_plot3 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                      size = ((pvij3[i]- min(ch_edge3))/(max(ch_edge3) - min(ch_edge3)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=16, face="bold",hjust = 0.5),
          legend.title = element_blank()) +
    ggtitle(paste("Larynx (T = ", T3, ")", sep=""))
}

# Colorectal cancer
ch_edge4 <- pvij4[which(est_diff4 == 1)]
edge_plot4 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group, fill = colrect_SIR),
               color = "black",
               alpha = 0.6
  ) +
  scale_fill_stepsn(colors = color,
                    breaks = breaks4)
for(i in which(est_diff4 == 1)){
  edge_plot4 = edge_plot4 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                      size = ((pvij4[i]- min(ch_edge4))/(max(ch_edge4) - min(ch_edge4)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=16, face="bold",hjust = 0.5),
          legend.title = element_blank()) +
    ggtitle(paste("Colorectal (T = ", T4, ")", sep=""))
}

ggarrange(edge_plot1, edge_plot2, edge_plot3, edge_plot4, nrow = 2, ncol = 2) 


```

###### III. Shared difference boundary for each pair of cancers
We calculate posterior probabilities: $P(\phi_{id} \neq \phi_{jd}, \phi_{id'} \neq \phi_{jd'} | y),\quad i \sim j,\quad d \neq d'$
```{r}
vij_samplesc12 <- t(apply(cbind(phis1,phis2), 1, function(x){
  diff_bc1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]] & x[neighbor_list0[i,1]+58] != x[neighbor_list0[i,2]+58]){
      diff_bc1 <- c(diff_bc1, 1)
    }else{
      diff_bc1 <- c(diff_bc1, 0)
    }
  }
  return(diff_bc1)
}))

vij_samplesc13 <- t(apply(cbind(phis1,phis3), 1, function(x){
  diff_bc1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]] & x[neighbor_list0[i,1]+58] != x[neighbor_list0[i,2]+58]){
      diff_bc1 <- c(diff_bc1, 1)
    }else{
      diff_bc1 <- c(diff_bc1, 0)
    }
  }
  return(diff_bc1)
}))

vij_samplesc14 <- t(apply(cbind(phis1,phis4), 1, function(x){
  diff_bc1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]] & x[neighbor_list0[i,1]+58] != x[neighbor_list0[i,2]+58]){
      diff_bc1 <- c(diff_bc1, 1)
    }else{
      diff_bc1 <- c(diff_bc1, 0)
    }
  }
  return(diff_bc1)
}))

vij_samplesc23 <- t(apply(cbind(phis2,phis3), 1, function(x){
  diff_bc1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]] & x[neighbor_list0[i,1]+58] != x[neighbor_list0[i,2]+58]){
      diff_bc1 <- c(diff_bc1, 1)
    }else{
      diff_bc1 <- c(diff_bc1, 0)
    }
  }
  return(diff_bc1)
}))

vij_samplesc24 <- t(apply(cbind(phis2,phis4), 1, function(x){
  diff_bc1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]] & x[neighbor_list0[i,1]+58] != x[neighbor_list0[i,2]+58]){
      diff_bc1 <- c(diff_bc1, 1)
    }else{
      diff_bc1 <- c(diff_bc1, 0)
    }
  }
  return(diff_bc1)
}))

vij_samplesc34 <- t(apply(cbind(phis3,phis4), 1, function(x){
  diff_bc1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]] & x[neighbor_list0[i,1]+58] != x[neighbor_list0[i,2]+58]){
      diff_bc1 <- c(diff_bc1, 1)
    }else{
      diff_bc1 <- c(diff_bc1, 0)
    }
  }
  return(diff_bc1)
}))

# probabilities: P(phi_id != phi_jd, phi_id' != phi_jd'), i ~ j, d != d'
pvijc12 <- apply(vij_samplesc12, 2, mean)
pvijc13 <- apply(vij_samplesc13, 2, mean)
pvijc14 <- apply(vij_samplesc14, 2, mean)
pvijc23 <- apply(vij_samplesc23, 2, mean)
pvijc24 <- apply(vij_samplesc24, 2, mean)
pvijc34 <- apply(vij_samplesc34, 2, mean)

# Estimated FDR
thresholdc12 = sort(pvijc12, decreasing = TRUE)[T_edge1]
thresholdc13 = sort(pvijc13, decreasing = TRUE)[T_edge1]
thresholdc14 = sort(pvijc14, decreasing = TRUE)[T_edge1]
thresholdc23 = sort(pvijc23, decreasing = TRUE)[T_edge1]
thresholdc24 = sort(pvijc24, decreasing = TRUE)[T_edge1]
thresholdc34 = sort(pvijc34, decreasing = TRUE)[T_edge1]

FDR_estc12 <- rep(0, length(T_edge1))
FDR_estc13 <- rep(0, length(T_edge1))
FDR_estc14 <- rep(0, length(T_edge1))
FDR_estc23 <- rep(0, length(T_edge1))
FDR_estc24 <- rep(0, length(T_edge1))
FDR_estc34 <- rep(0, length(T_edge1))


for(i in 1:length(threshold1)){
  thc12 <- thresholdc12[i]
  est_diffc12 <- as.numeric(pvijc12 >= thc12)
  FDR_estc12[i] <- sum((1-pvijc12) * est_diffc12)  / sum(est_diffc12)
  
  thc13 <- thresholdc13[i]
  est_diffc13 <- as.numeric(pvijc13 >= thc13)
  FDR_estc13[i] <- sum((1-pvijc13) * est_diffc13)  / sum(est_diffc13)
  
  thc14 <- thresholdc14[i]
  est_diffc14 <- as.numeric(pvijc14 >= thc14)
  FDR_estc14[i] <- sum((1-pvijc14) * est_diffc14)  / sum(est_diffc14)
  
  thc23 <- thresholdc23[i]
  est_diffc23 <- as.numeric(pvijc23 >= thc23)
  FDR_estc23[i] <- sum((1-pvijc23) * est_diffc23)  / sum(est_diffc23)
  
  thc24 <- thresholdc24[i]
  est_diffc24 <- as.numeric(pvijc24 >= thc24)
  FDR_estc24[i] <- sum((1-pvijc24) * est_diffc24)  / sum(est_diffc24)
  
  thc34 <- thresholdc34[i]
  est_diffc34 <- as.numeric(pvijc34 >= thc34)
  FDR_estc34[i] <- sum((1-pvijc34) * est_diffc34)  / sum(est_diffc34)
}
```

Use the same threshold as individual cancer and identify shared boundaries
```{r}
alpha_nc = 0.1
T12c = sum(FDR_estc12<=alpha_nc)
T13c = sum(FDR_estc13<=alpha_nc)
T14c = sum(FDR_estc14<=alpha_nc)
T23c = sum(FDR_estc23<=alpha_nc)
T24c = sum(FDR_estc24<=alpha_nc)
T34c = sum(FDR_estc34<=alpha_nc)


est_diffc12 <- as.numeric(pvijc12 >= thresholdc12[T12c])
neighbor_list_diffc12 <- neighbor_list0[est_diffc12 == 1, ]

est_diffc13 <- as.numeric(pvijc13 >= thresholdc13[T13c])
neighbor_list_diffc13 <- neighbor_list0[est_diffc13 == 1, ]

est_diffc14 <- as.numeric(pvijc14 >= thresholdc14[T14c])
neighbor_list_diffc14 <- neighbor_list0[est_diffc14 == 1, ]

est_diffc23 <- as.numeric(pvijc23 >= thresholdc23[T23c])
neighbor_list_diffc23 <- neighbor_list0[est_diffc23 == 1, ]

est_diffc24 <- as.numeric(pvijc24 >= thresholdc24[T24c])
neighbor_list_diffc24 <- neighbor_list0[est_diffc24 == 1, ]

est_diffc34 <- as.numeric(pvijc34 >= thresholdc34[T34c])
neighbor_list_diffc34 <- neighbor_list0[est_diffc34 == 1, ]

ch_edge12 <- pvijc12[est_diffc12==1]
edge_plotc12 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group),
               color = "black",
               fill = "white"
  )
for(i in which(est_diffc12==1)){
  edge_plotc12 = edge_plotc12 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                          size = ((pvijc12[i]- min(ch_edge12))/(max(ch_edge12) - min(ch_edge12)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=14, face="bold",hjust = 0.5)) +
    ggtitle(paste("Lung, Esophageal (T = ", T12c, ")", sep=""))
}

ch_edge13 <- pvijc13[est_diffc13==1]
edge_plotc13 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group),
               color = "black",
               fill = "white"
  )
for(i in which(est_diffc13==1)){
  edge_plotc13 = edge_plotc13 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                          size = ((pvijc13[i]- min(ch_edge13))/(max(ch_edge13) - min(ch_edge13)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=14, face="bold",hjust = 0.5)) +
    ggtitle(paste("Lung, Layrnx (T = ", T13c, ")", sep=""))
}

ch_edge14 <- pvijc14[est_diffc14==1]
edge_plotc14 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group),
               color = "black",
               fill = "white"
  )
for(i in which(est_diffc14==1)){
  edge_plotc14 = edge_plotc14 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                          size = ((pvijc14[i]- min(ch_edge14))/(max(ch_edge14) - min(ch_edge14)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=14, face="bold",hjust = 0.5)) +
    ggtitle(paste("Lung, Colorectal (T = ", T14c, ")", sep=""))
}

ch_edge23 <- pvijc23[est_diffc23==1]
edge_plotc23 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group),
               color = "black",
               fill = "white"
  )
for(i in which(est_diffc23==1)){
  edge_plotc23 = edge_plotc23 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                          size = ((pvijc23[i]- min(ch_edge23))/(max(ch_edge23) - min(ch_edge23)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=14, face="bold",hjust = 0.5)) +
    ggtitle(paste("Esophageal, Layrnx (T = ", T23c, ")", sep=""))
}

ch_edge24 <- pvijc24[est_diffc24==1]
edge_plotc24 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group),
               color = "black",
               fill = "white"
  )
for(i in which(est_diffc24==1)){
  edge_plotc24 = edge_plotc24 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                          size = ((pvijc24[i]- min(ch_edge24))/(max(ch_edge24) - min(ch_edge24)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=14, face="bold",hjust = 0.5)) +
    ggtitle(paste("Esophageal, Colorectal (T = ", T24c, ")", sep=""))
}

ch_edge34 <- pvijc34[est_diffc34==1]
edge_plotc34 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group),
               color = "black",
               fill = "white"
  )
for(i in which(est_diffc34==1)){
  edge_plotc34 = edge_plotc34 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                          size = ((pvijc34[i]- min(ch_edge34))/(max(ch_edge34) - min(ch_edge34)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=14, face="bold",hjust = 0.5)) +
    ggtitle(paste("Larynx, Colorectal (T = ", T34c, ")", sep=""))
}

ggarrange(edge_plotc12, edge_plotc13, edge_plotc14, 
          edge_plotc23, edge_plotc24, edge_plotc34, nrow = 2, ncol = 3)
```

###### IV. Model fitting: D score in Table 4
```{r}
Y_rep1 <- matrix(0, nrow = n, ncol = 10000)
Y_rep2 <- matrix(0, nrow = n, ncol = 10000)
Y_rep3 <- matrix(0, nrow = n, ncol = 10000)
Y_rep4 <- matrix(0, nrow = n, ncol = 10000)
for(i in 1:10000){
  
  beta1 <- mcmc_samples$beta[i,1]
  beta2 <- mcmc_samples$beta[i,2]
  beta3 <- mcmc_samples$beta[i,3]
  beta4 <- mcmc_samples$beta[i,4]
  
  W1_mcmc = mcmc_samples$phi[i,1:58]
  W2_mcmc = mcmc_samples$phi[i,59:116]
  W3_mcmc = mcmc_samples$phi[i,117:174]
  W4_mcmc = mcmc_samples$phi[i,175:232]
  
  Y_rep1[,i] <- rpois(n, E1 * exp(X1[,1] * beta1 +  W1_mcmc)) / E1
  Y_rep2[,i] <- rpois(n, E2 * exp(X2[,1] * beta2 +  W2_mcmc)) / E2
  Y_rep3[,i] <- rpois(n, E3 * exp(X3[,1] * beta3 +  W3_mcmc)) / E3
  Y_rep4[,i] <- rpois(n, E4 * exp(X4[,1] * beta4 +  W4_mcmc)) / E4
  
}
mu_rep1 = rowMeans(Y_rep1)
mu_rep2 = rowMeans(Y_rep2)
mu_rep3 = rowMeans(Y_rep3)
mu_rep4 = rowMeans(Y_rep4)

var_rep1 = rowVars(Y_rep1)
var_rep2 = rowVars(Y_rep2)
var_rep3 = rowVars(Y_rep3)
var_rep4 = rowVars(Y_rep4)

G.latent1 = sum((Y1/E1 - mu_rep1)^2)
P.latent1 = sum(var_rep1)
D.latent1 = G.latent1 + P.latent1

G.latent2 = sum((Y2/E2 - mu_rep2)^2)
P.latent2 = sum(var_rep2)
D.latent2 = G.latent2 + P.latent2

G.latent3 = sum((Y3/E3 - mu_rep3)^2)
P.latent3 = sum(var_rep3)
D.latent3 = G.latent3 + P.latent3

G.latent4 = sum((Y4/E4 - mu_rep4)^2)
P.latent4 = sum(var_rep4)
D.latent4 = G.latent4 + P.latent4

D_DAGARind = c(D.latent1, D.latent2, D.latent3, D.latent4)
D_DAGARind
sum(D_DAGARind)
```

##### 3.3.2 CAR$_{ind}$ Model
```
sink("MCAR_ARDP_ind.txt")
cat("
    model
    {
    
    for (i in 1:k)
    {
    log(mu[i]) <- log(E[i]) + X[i,] %*% beta + phi[i]
    Y[i] ~ dpois(mu[i])
    phi[i] <- inprod(theta, u[i,])
    Fr[i] <- pnorm(r[i], 0, 1/invQ1[i,i])
    for(h in 2:H){
    u[i, h] <- ifelse(Fr[i] > sum_pi[h-1] && Fr[i] < sum_pi[h], 1, 0)
    }
    u[i,1] <- ifelse(sum(u[i, 2:H])==0, 1, 0)
    }
    
    for (i in (k+1):(2*k))
    {
    log(mu[i]) <- log(E[i]) + X[i,] %*% beta + phi[i]
    Y[i] ~ dpois(mu[i])
    phi[i] <- inprod(theta, u[i,])
    Fr[i] <- pnorm(r[i], 0, 1/invQ2[i-k,i-k])
    for(h in 2:H){
    u[i, h] <- ifelse(Fr[i] > sum_pi[h-1] && Fr[i] < sum_pi[h], 1, 0)
    }
    u[i,1] <- ifelse(sum(u[i, 2:H])==0, 1, 0)
    }
    
    for (i in (2*k+1):(3*k))
    {
    log(mu[i]) <- log(E[i]) + X[i,] %*% beta + phi[i]
    Y[i] ~ dpois(mu[i])
    phi[i] <- inprod(theta, u[i,])
    Fr[i] <- pnorm(r[i], 0, 1/invQ3[i-2*k,i-2*k])
    for(h in 2:H){
    u[i, h] <- ifelse(Fr[i] > sum_pi[h-1] && Fr[i] < sum_pi[h], 1, 0)
    }
    u[i,1] <- ifelse(sum(u[i, 2:H])==0, 1, 0)
    }
    
    for (i in (3*k+1):(4*k))
    {
    log(mu[i]) <- log(E[i]) + X[i,] %*% beta + phi[i]
    Y[i] ~ dpois(mu[i])
    phi[i] <- inprod(theta, u[i,])
    Fr[i] <- pnorm(r[i], 0, 1/invQ4[i-3*k,i-3*k])
    for(h in 2:H){
    u[i, h] <- ifelse(Fr[i] > sum_pi[h-1] && Fr[i] < sum_pi[h], 1, 0)
    }
    u[i,1] <- ifelse(sum(u[i, 2:H])==0, 1, 0)
    }
    
    
    pi[1] <- V[1]
    prod[1] <- 1 - pi[1]
    for (h in 2:(H-1)){
    prod[h] <- prod[h-1] * (1-V[h])
    pi[h] <- V[h] * prod[h-1]
    }
    pi[H] <- 1-sum(pi[1:(H-1)])
    
    for (h in 1:H){
    theta[h] ~ dnorm(0,taus)
    V[h] ~ dbeta(1,alpha)
    sum_pi[h] <- sum(pi[1:h])
    }
    
    Q1 <- tau1 * (D - rho1 * Minc)
    Q2 <- tau2 * (D - rho2 * Minc)
    Q3 <- tau3 * (D - rho3 * Minc)
    Q4 <- tau4 * (D - rho4 * Minc)
    
    r1[1:k] ~ dmnorm(rep(0, k), Q1)
    r2[1:k] ~ dmnorm(rep(0, k), Q2)
    r3[1:k] ~ dmnorm(rep(0, k), Q3)
    r4[1:k] ~ dmnorm(rep(0, k), Q4)
    r[1:k] <- r1[1:k]
    r[(k+1):(2*k)] <- r2[1:k]
    r[(2*k+1):(3*k)] <- r3[1:k]
    r[(3*k+1):(4*k)] <- r4[1:k]
    
    invQ1 <- inverse(Q1)
    invQ2 <- inverse(Q2)
    invQ3 <- inverse(Q3)
    invQ4 <- inverse(Q4)
    
    rho1 ~ dunif(0, 0.98)
    rho2 ~ dunif(0, 0.98)
    rho3 ~ dunif(0, 0.98)
    rho4 ~ dunif(0, 0.98)
    tau1 ~ dgamma(2, 0.1)
    tau2 ~ dgamma(2, 0.1)
    tau3 ~ dgamma(2, 0.1)
    tau4 ~ dgamma(2, 0.1)
    
    taus ~ dgamma(2,1)
    beta[1:ncolumn] ~ dmnorm(rep(0,ncolumn), (0.0001*I));
    }
    ", fill = TRUE)
sink()
```

```{r eval=FALSE}
model.data1 <- list(k = n,  I = diag(ncol(X)), X = X, Y = Y, E = E, 
                    alpha = 1, ncolumn = ncol(X),  H = 15,  D = D, Minc = Minc)
model.inits <- rep(list(list(rho1 = 0.1, rho2 = 0.1, rho3 = 0.1, rho4 = 0.1, tau1 = 1, tau2 = 1, tau3 = 1, tau4 = 1,  
                             taus = 1, beta = rep(0, ncol(X)))),2)
model.param <- c("beta", "rho1", "rho2", "rho3", "rho4", "tau1", "tau2", "tau3", "tau4", 
                 "taus", "phi", "u")
set.seed(123)
mcmc_samples <- jags(model.data1, model.inits, model.param, "MCAR_ARDP_ind.txt",
                 n.chains = 2, n.iter = 25000,n.burnin = 20000, n.thin = 1)
# Parameter estimates
sample.mcmc = mcmc_samples$BUGSoutput$sims.matrix[,c(246, 238:245)]
```

###### I. Obtain spatial randome effects in original spatial order
```{r}
phis <- mcmc_samples$BUGSoutput$sims.matrix[,6:237]
phis1 <- phis[,1:58][,order(final_perm)]
phis2 <- phis[,59:116][,order(final_perm)]
phis3 <- phis[,117:174][,order(final_perm)]
phis4 <- phis[,175:232][,order(final_perm)]
phis_origin <- cbind(phis1, phis2, phis3, phis4)
```

###### II. Difference boundaries for each cancer individually
We calculate posterior probabilities: $P(\phi_{id} \neq \phi_{jd} | y), \quad i \sim j$
```{r}
vij_samples1 <- t(apply(phis1, 1, function(x){
  diff_b1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]]){
      diff_b1 <- c(diff_b1, 1)
    }else{
      diff_b1 <- c(diff_b1, 0)
    }
  }
  return(diff_b1)
}))

vij_samples2 <- t(apply(phis2, 1, function(x){
  diff_b1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]]){
      diff_b1 <- c(diff_b1, 1)
    }else{
      diff_b1 <- c(diff_b1, 0)
    }
  }
  return(diff_b1)
}))

vij_samples3 <- t(apply(phis3, 1, function(x){
  diff_b1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]]){
      diff_b1 <- c(diff_b1, 1)
    }else{
      diff_b1 <- c(diff_b1, 0)
    }
  }
  return(diff_b1)
}))

vij_samples4 <- t(apply(phis4, 1, function(x){
  diff_b1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]]){
      diff_b1 <- c(diff_b1, 1)
    }else{
      diff_b1 <- c(diff_b1, 0)
    }
  }
  return(diff_b1)
}))

# probabilities: P(phi_id != phi_jd), i~j
pvij1 <- apply(vij_samples1, 2, mean)
pvij2 <- apply(vij_samples2, 2, mean)
pvij3 <- apply(vij_samples3, 2, mean)
pvij4 <- apply(vij_samples4, 2, mean)

names(pvij1) <- neighbor_name
names(pvij2) <- neighbor_name
names(pvij3) <- neighbor_name
names(pvij4) <- neighbor_name

```

####### a. Estimated FDR curves
```{r}
T_edge1 = seq(0, 135, 1)[-1]
threshold1 = sort(pvij1, decreasing = TRUE)[T_edge1]
threshold2 = sort(pvij2, decreasing = TRUE)[T_edge1]
threshold3 = sort(pvij3, decreasing = TRUE)[T_edge1]
threshold4 = sort(pvij4, decreasing = TRUE)[T_edge1]
FDR_est1 <- rep(0, length(T_edge1))
FDR_est2 <- rep(0, length(T_edge1))
FDR_est3 <- rep(0, length(T_edge1))
FDR_est4 <- rep(0, length(T_edge1))


for(i in 1:length(threshold1)){
  th1 <- threshold1[i]
  est_diff1 <- as.numeric(pvij1 >= th1)
  FDR_est1[i] <- sum((1-pvij1) * est_diff1)  / sum(est_diff1)
  
  th2 <- threshold2[i]
  est_diff2 <- as.numeric(pvij2 >= th2)
  FDR_est2[i] <- sum((1-pvij2) * est_diff2)  / sum(est_diff2)
  
  th3 <- threshold3[i]
  est_diff3 <- as.numeric(pvij3 >= th3)
  FDR_est3[i] <- sum((1-pvij3) * est_diff3)  / sum(est_diff3)
  
  th4 <- threshold4[i]
  est_diff4 <- as.numeric(pvij4 >= th4)
  FDR_est4[i] <- sum((1-pvij4) * est_diff4)  / sum(est_diff4)
}

plot(FDR_est1,type="l", lty = 1, xlab="Number of edges selected", ylab = "Estimated FDR", ylim=c(0,0.5))
lines(FDR_est2,type="l", lty = 2)
lines(FDR_est3,type="l", lty = 3)
lines(FDR_est4,type="l", lty = 4)
legend("topleft", legend=c("Lung", "Esophageal", "Layrnx", "Colorectal"),
       lty=1:4, cex=0.7)
```

####### b. Difference boundary detection
By setting FDR threshold as $0.1$, we get difference boundaries detected for each cancer individually.
```{r}
# Set threshold for FDR: 0.025
alpha_n = 0.1
T1 = sum(FDR_est1<=alpha_n)
T2 = sum(FDR_est2<=alpha_n)
T3 = sum(FDR_est3<=alpha_n)
T4 = sum(FDR_est4<=alpha_n)

# Difference boundary for each cancer with FDR <= 0.025
est_diff1 <- as.numeric(pvij1 >= threshold1[T1])

est_diff2 <- as.numeric(pvij2 >= threshold2[T2])

est_diff3 <- as.numeric(pvij3 >= threshold3[T3])

est_diff4 <- as.numeric(pvij4 >= threshold4[T4])

# Lung cancer
ch_edge1 <- pvij1[which(est_diff1 == 1)]
edge_plot1 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group, fill = lung_SIR),
               color = "black",
               alpha = 0.6
  ) +
  scale_fill_stepsn(colors = color,
                    breaks = breaks1)

for(i in which(est_diff1 == 1)){
  edge_plot1 = edge_plot1 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                      size = ((pvij1[i]- min(ch_edge1))/(max(ch_edge1) - min(ch_edge1)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=16, face="bold",hjust = 0.5),
          legend.title = element_blank()) +
    ggtitle(paste("Lung (T = ", T1, ")", sep=""))
}

# Esophageal cancer
ch_edge2 <- pvij2[which(est_diff2 == 1)]
edge_plot2 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group, fill = esophagus_SIR),
               color = "black",
               alpha = 0.6
  ) +
  scale_fill_stepsn(colors = color,
                    breaks = breaks2)
for(i in which(est_diff2 == 1)){
  edge_plot2 = edge_plot2 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                      size = ((pvij2[i]- min(ch_edge2))/(max(ch_edge2) - min(ch_edge2)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=16, face="bold",hjust = 0.5),
          legend.title = element_blank()) +
    ggtitle(paste("Esophageal (T = ", T2, ")", sep=""))
}

# Larynx cancer
ch_edge3 <- pvij3[which(est_diff3 == 1)]
edge_plot3 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group, fill = larynx_SIR),
               color = "black",
               alpha = 0.6
  ) +
  scale_fill_stepsn(colors = color,
                    breaks = breaks3)
for(i in which(est_diff3 == 1)){
  edge_plot3 = edge_plot3 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                      size = ((pvij3[i]- min(ch_edge3))/(max(ch_edge3) - min(ch_edge3)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=16, face="bold",hjust = 0.5),
          legend.title = element_blank()) +
    ggtitle(paste("Larynx (T = ", T3, ")", sep=""))
}

# Colorectal cancer
ch_edge4 <- pvij4[which(est_diff4 == 1)]
edge_plot4 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group, fill = colrect_SIR),
               color = "black",
               alpha = 0.6
  ) +
  scale_fill_stepsn(colors = color,
                    breaks = breaks4)
for(i in which(est_diff4 == 1)){
  edge_plot4 = edge_plot4 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                      size = ((pvij4[i]- min(ch_edge4))/(max(ch_edge4) - min(ch_edge4)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=16, face="bold",hjust = 0.5),
          legend.title = element_blank()) +
    ggtitle(paste("Colorectal (T = ", T4, ")", sep=""))
}
ggarrange(edge_plot1, edge_plot2, edge_plot3, edge_plot4, nrow = 2, ncol = 2) 


```

###### III. Shared difference boundary for each pair of cancers
We calculate posterior probabilities: $P(\phi_{id} \neq \phi_{jd}, \phi_{id'} \neq \phi_{jd'} | y),\quad i \sim j,\quad d \neq d'$
```{r}
vij_samplesc12 <- t(apply(cbind(phis1,phis2), 1, function(x){
  diff_bc1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]] & x[neighbor_list0[i,1]+58] != x[neighbor_list0[i,2]+58]){
      diff_bc1 <- c(diff_bc1, 1)
    }else{
      diff_bc1 <- c(diff_bc1, 0)
    }
  }
  return(diff_bc1)
}))

vij_samplesc13 <- t(apply(cbind(phis1,phis3), 1, function(x){
  diff_bc1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]] & x[neighbor_list0[i,1]+58] != x[neighbor_list0[i,2]+58]){
      diff_bc1 <- c(diff_bc1, 1)
    }else{
      diff_bc1 <- c(diff_bc1, 0)
    }
  }
  return(diff_bc1)
}))

vij_samplesc14 <- t(apply(cbind(phis1,phis4), 1, function(x){
  diff_bc1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]] & x[neighbor_list0[i,1]+58] != x[neighbor_list0[i,2]+58]){
      diff_bc1 <- c(diff_bc1, 1)
    }else{
      diff_bc1 <- c(diff_bc1, 0)
    }
  }
  return(diff_bc1)
}))

vij_samplesc23 <- t(apply(cbind(phis2,phis3), 1, function(x){
  diff_bc1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]] & x[neighbor_list0[i,1]+58] != x[neighbor_list0[i,2]+58]){
      diff_bc1 <- c(diff_bc1, 1)
    }else{
      diff_bc1 <- c(diff_bc1, 0)
    }
  }
  return(diff_bc1)
}))

vij_samplesc24 <- t(apply(cbind(phis2,phis4), 1, function(x){
  diff_bc1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]] & x[neighbor_list0[i,1]+58] != x[neighbor_list0[i,2]+58]){
      diff_bc1 <- c(diff_bc1, 1)
    }else{
      diff_bc1 <- c(diff_bc1, 0)
    }
  }
  return(diff_bc1)
}))

vij_samplesc34 <- t(apply(cbind(phis3,phis4), 1, function(x){
  diff_bc1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]] & x[neighbor_list0[i,1]+58] != x[neighbor_list0[i,2]+58]){
      diff_bc1 <- c(diff_bc1, 1)
    }else{
      diff_bc1 <- c(diff_bc1, 0)
    }
  }
  return(diff_bc1)
}))

# probabilities: P(phi_id != phi_jd, phi_id' != phi_jd'), i ~ j, d != d'
pvijc12 <- apply(vij_samplesc12, 2, mean)
pvijc13 <- apply(vij_samplesc13, 2, mean)
pvijc14 <- apply(vij_samplesc14, 2, mean)
pvijc23 <- apply(vij_samplesc23, 2, mean)
pvijc24 <- apply(vij_samplesc24, 2, mean)
pvijc34 <- apply(vij_samplesc34, 2, mean)

# Estimated FDR
thresholdc12 = sort(pvijc12, decreasing = TRUE)[T_edge1]
thresholdc13 = sort(pvijc13, decreasing = TRUE)[T_edge1]
thresholdc14 = sort(pvijc14, decreasing = TRUE)[T_edge1]
thresholdc23 = sort(pvijc23, decreasing = TRUE)[T_edge1]
thresholdc24 = sort(pvijc24, decreasing = TRUE)[T_edge1]
thresholdc34 = sort(pvijc34, decreasing = TRUE)[T_edge1]

FDR_estc12 <- rep(0, length(T_edge1))
FDR_estc13 <- rep(0, length(T_edge1))
FDR_estc14 <- rep(0, length(T_edge1))
FDR_estc23 <- rep(0, length(T_edge1))
FDR_estc24 <- rep(0, length(T_edge1))
FDR_estc34 <- rep(0, length(T_edge1))


for(i in 1:length(threshold1)){
  thc12 <- thresholdc12[i]
  est_diffc12 <- as.numeric(pvijc12 >= thc12)
  FDR_estc12[i] <- sum((1-pvijc12) * est_diffc12)  / sum(est_diffc12)
  
  thc13 <- thresholdc13[i]
  est_diffc13 <- as.numeric(pvijc13 >= thc13)
  FDR_estc13[i] <- sum((1-pvijc13) * est_diffc13)  / sum(est_diffc13)
  
  thc14 <- thresholdc14[i]
  est_diffc14 <- as.numeric(pvijc14 >= thc14)
  FDR_estc14[i] <- sum((1-pvijc14) * est_diffc14)  / sum(est_diffc14)
  
  thc23 <- thresholdc23[i]
  est_diffc23 <- as.numeric(pvijc23 >= thc23)
  FDR_estc23[i] <- sum((1-pvijc23) * est_diffc23)  / sum(est_diffc23)
  
  thc24 <- thresholdc24[i]
  est_diffc24 <- as.numeric(pvijc24 >= thc24)
  FDR_estc24[i] <- sum((1-pvijc24) * est_diffc24)  / sum(est_diffc24)
  
  thc34 <- thresholdc34[i]
  est_diffc34 <- as.numeric(pvijc34 >= thc34)
  FDR_estc34[i] <- sum((1-pvijc34) * est_diffc34)  / sum(est_diffc34)
}
```

Use the same threshold as individual cancer and identify shared boundaries
```{r}
alpha_nc = 0.1
T12c = sum(FDR_estc12<=alpha_nc)
T13c = sum(FDR_estc13<=alpha_nc)
T14c = sum(FDR_estc14<=alpha_nc)
T23c = sum(FDR_estc23<=alpha_nc)
T24c = sum(FDR_estc24<=alpha_nc)
T34c = sum(FDR_estc34<=alpha_nc)


est_diffc12 <- as.numeric(pvijc12 >= thresholdc12[T12c])
neighbor_list_diffc12 <- neighbor_list0[est_diffc12 == 1, ]

est_diffc13 <- as.numeric(pvijc13 >= thresholdc13[T13c])
neighbor_list_diffc13 <- neighbor_list0[est_diffc13 == 1, ]

est_diffc14 <- as.numeric(pvijc14 >= thresholdc14[T14c])
neighbor_list_diffc14 <- neighbor_list0[est_diffc14 == 1, ]

est_diffc23 <- as.numeric(pvijc23 >= thresholdc23[T23c])
neighbor_list_diffc23 <- neighbor_list0[est_diffc23 == 1, ]

est_diffc24 <- as.numeric(pvijc24 >= thresholdc24[T24c])
neighbor_list_diffc24 <- neighbor_list0[est_diffc24 == 1, ]

est_diffc34 <- as.numeric(pvijc34 >= thresholdc34[T34c])
neighbor_list_diffc34 <- neighbor_list0[est_diffc34 == 1, ]

ch_edge12 <- pvijc12[est_diffc12==1]
edge_plotc12 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group),
               color = "black",
               fill = "white"
  )
for(i in which(est_diffc12==1)){
  edge_plotc12 = edge_plotc12 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                          size = ((pvijc12[i]- min(ch_edge12))/(max(ch_edge12) - min(ch_edge12)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=14, face="bold",hjust = 0.5)) +
    ggtitle(paste("Lung, Esophageal (T = ", T12c, ")", sep=""))
}

ch_edge13 <- pvijc13[est_diffc13==1]
edge_plotc13 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group),
               color = "black",
               fill = "white"
  )
for(i in which(est_diffc13==1)){
  edge_plotc13 = edge_plotc13 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                          size = ((pvijc13[i]- min(ch_edge13))/(max(ch_edge13) - min(ch_edge13)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=14, face="bold",hjust = 0.5)) +
    ggtitle(paste("Lung, Layrnx (T = ", T13c, ")", sep=""))
}

ch_edge14 <- pvijc14[est_diffc14==1]
edge_plotc14 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group),
               color = "black",
               fill = "white"
  )
for(i in which(est_diffc14==1)){
  edge_plotc14 = edge_plotc14 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                          size = ((pvijc14[i]- min(ch_edge14))/(max(ch_edge14) - min(ch_edge14)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=14, face="bold",hjust = 0.5)) +
    ggtitle(paste("Lung, Colorectal (T = ", T14c, ")", sep=""))
}

ch_edge23 <- pvijc23[est_diffc23==1]
edge_plotc23 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group),
               color = "black",
               fill = "white"
  )
for(i in which(est_diffc23==1)){
  edge_plotc23 = edge_plotc23 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                          size = ((pvijc23[i]- min(ch_edge23))/(max(ch_edge23) - min(ch_edge23)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=14, face="bold",hjust = 0.5)) +
    ggtitle(paste("Esophageal, Layrnx (T = ", T23c, ")", sep=""))
}

ch_edge24 <- pvijc24[est_diffc24==1]
edge_plotc24 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group),
               color = "black",
               fill = "white"
  )
for(i in which(est_diffc24==1)){
  edge_plotc24 = edge_plotc24 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                          size = ((pvijc24[i]- min(ch_edge24))/(max(ch_edge24) - min(ch_edge24)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=14, face="bold",hjust = 0.5)) +
    ggtitle(paste("Esophageal, Colorectal (T = ", T24c, ")", sep=""))
}

ch_edge34 <- pvijc34[est_diffc34==1]
edge_plotc34 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group),
               color = "black",
               fill = "white"
  )
for(i in which(est_diffc34==1)){
  edge_plotc34 = edge_plotc34 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                          size = ((pvijc34[i]- min(ch_edge34))/(max(ch_edge34) - min(ch_edge34)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=14, face="bold",hjust = 0.5)) +
    ggtitle(paste("Larynx, Colorectal (T = ", T34c, ")", sep=""))
}


ggarrange(edge_plotc12, edge_plotc13, edge_plotc14, 
          edge_plotc23, edge_plotc24, edge_plotc34, nrow = 2, ncol = 3)
```

###### IV. Model fitting: D score in Table 4
```{r eval=FALSE}
Y_rep1 <- matrix(0, nrow = n, ncol = 10000)
Y_rep2 <- matrix(0, nrow = n, ncol = 10000)
Y_rep3 <- matrix(0, nrow = n, ncol = 10000)
Y_rep4 <- matrix(0, nrow = n, ncol = 10000)
for(i in 1:10000){
  
  beta1 <- mcmc_samples$beta[i,1]
  beta2 <- mcmc_samples$beta[i,2]
  beta3 <- mcmc_samples$beta[i,3]
  beta4 <- mcmc_samples$beta[i,4]
  
  W1_mcmc = mcmc_samples$phi[i,1:58]
  W2_mcmc = mcmc_samples$phi[i,59:116]
  W3_mcmc = mcmc_samples$phi[i,117:174]
  W4_mcmc = mcmc_samples$phi[i,175:232]
  
  Y_rep1[,i] <- rpois(n, E1 * exp(X1[,1] * beta1 +  W1_mcmc)) / E1
  Y_rep2[,i] <- rpois(n, E2 * exp(X2[,1] * beta2 +  W2_mcmc)) / E2
  Y_rep3[,i] <- rpois(n, E3 * exp(X3[,1] * beta3 +  W3_mcmc)) / E3
  Y_rep4[,i] <- rpois(n, E4 * exp(X4[,1] * beta4 +  W4_mcmc)) / E4
  
}
mu_rep1 = rowMeans(Y_rep1)
mu_rep2 = rowMeans(Y_rep2)
mu_rep3 = rowMeans(Y_rep3)
mu_rep4 = rowMeans(Y_rep4)

var_rep1 = rowVars(Y_rep1)
var_rep2 = rowVars(Y_rep2)
var_rep3 = rowVars(Y_rep3)
var_rep4 = rowVars(Y_rep4)

G.latent1 = sum((Y1/E1 - mu_rep1)^2)
P.latent1 = sum(var_rep1)
D.latent1 = G.latent1 + P.latent1

G.latent2 = sum((Y2/E2 - mu_rep2)^2)
P.latent2 = sum(var_rep2)
D.latent2 = G.latent2 + P.latent2

G.latent3 = sum((Y3/E3 - mu_rep3)^2)
P.latent3 = sum(var_rep3)
D.latent3 = G.latent3 + P.latent3

G.latent4 = sum((Y4/E4 - mu_rep4)^2)
P.latent4 = sum(var_rep4)
D.latent4 = G.latent4 + P.latent4

D_CARind = c(D.latent1, D.latent2, D.latent3, D.latent4)
D_CARind
sum(D_CARind)
```

### 4. Difference boundary detection accounting for covariates using MDAGAR
Accounting for covariates would change difference boundaries detected for each cancer. Here we include three covariates sequentially: (1)smoking; (2) smoking and age; (3) smoking, age and unemployed

#### 4.1 Account for smoking
```{r}
# Construct covariates matrix
X = as.matrix(bdiag(bdiag(X1[,1:2], X2[,1:2]), bdiag(X3[,1:2],X4[,1:2])))

RDP_joint_diff_smoke<-function(y=Y, x1=as.matrix(X1[,1:2]), x2=as.matrix(X2[,1:2]), x3 = as.matrix(X3[,1:2]), x4 = as.matrix(X4[,1:2]), 
                          X = X, E = E, Minc, alpha=1, q = 4, n.atoms=15, runs=10000, burn=1000){
  #y:       data
  #x:       covariates
  #n.atoms: number of atoms in the mixture dist.
  #theta:      the theta's (iid from baseline) in the model
  #alpha:     v~beta(1,alpha)
  #u:       the index indicator of spatial random effect 
  #rho:     sptatial autocorrelation parameter in DAGAR
  #Minc:       0-1 adjacency matrix 
  #V_r:     covariance matrix of joint MDAGAR
  #Q:       presicion matrix of DAGAR
  #r:       random effects following DAGAR
  #F_r:     Marginal CDF of r
  #taued:   presicion in Baseline of DP for disease d   
  #taus:    precision for theta
  
  
  
  
  nq<-length(y)
  n <- nq / q
  p<-ncol(X)
  y1 <- y[1:n]
  y2 <- y[(n+1):(2*n)]
  y3 <- y[(2*n+1):(3*n)]
  y4 <- y[(3*n+1):(4*n)]
  
  E1 <- E[1:n]
  E2 <- E[(n+1):(2*n)]
  E3 <- E[(2*n+1):(3*n)]
  E4 <- E[(3*n+1):(4*n)]
  
  sigmasq_beta = 10000
  keepbeta=matrix(0,runs,p)
  keepphi=matrix(0,runs,nq)
  keeptheta=matrix(0,runs,n.atoms)
  keepA=matrix(0,runs,10)
  keepu=matrix(0,runs,nq)
  keeprho1=keeprho2=keeprho3=keeprho4=keeptaus = rep(0,runs)
  keepv=matrix(0,runs,n.atoms)
  keepr=matrix(0,runs,nq)
  keepFr=matrix(0,runs,nq)
  
  
  #initial values
  
  theta=rep(0,n.atoms)
  
  beta1<-rep(0,ncol(x1))
  beta2<-rep(0,ncol(x2))
  beta3<-rep(0,ncol(x3))
  beta4<-rep(0,ncol(x4))
  beta = c(beta1, beta2, beta3, beta4)
  
  taus = 1
  
  c=2
  d=0.1
  d2=1
  v<-rep(.1,n.atoms)
  probs=makeprobs(v)
  
  #A matrix
  rho=c(0.5, 0.5, 0.5, 0.5)
  eta = log(rho/(1-rho))
  A = matrix(0, q, q)
  for(i in 1:q){
    for(j in 1:i){
      if(j == i){
        A[i, j] = exp(rnorm(1))
      }else{
        A[i, j] = rnorm(1)
      }
    }
  }
  
  Q = Dinv_new(Rho = rho, n, cn, ns, udnei, q)
  invQ1 = solve(Q[[1]])
  invQ2 = solve(Q[[2]])
  invQ3 = solve(Q[[3]])
  invQ4 = solve(Q[[4]])
  
  kprod = kronecker(A, diag(n))
  invQ = as.matrix(bdiag(bdiag(invQ1, invQ2), bdiag(invQ3, invQ4)))
  
  Vr = as.matrix(forceSymmetric(kprod %*% invQ %*% t(kprod)))
  r = rmvnorm(1, rep(0, nq), Vr)
  
  
  F_r=pnorm(r,0,sqrt(diag(Vr)))
  u=makeu(F_r,probs)
  phi <- theta[u]
  
  nu = 2
  R = 0.1 * diag(q)
  
  
  acceptr=acceptv=acceptrho=acceptA=0
  acceptbeta1 = acceptbeta2=acceptbeta3=acceptbeta4 = 0
  accepttheta = 0
  
  count<-afterburn<-0
  burn = burn + 1

  for(iter in 1:runs){
    
    if(iter %% 100 == 0){
      print(iter)
      print(acceptA/(iter-1)) 
      print(acceptr/nq/(iter-1))
      print(acceptrho/(iter-1)) 
      print(acceptv/n.atoms/(iter-1))
      print(accepttheta/n.atoms/(iter-1))
      print(acceptbeta1/(iter-1))
      print(acceptbeta2/(iter-1))
      print(acceptbeta3/(iter-1))
      print(acceptbeta4/(iter-1))
    }
    ######################
    ###   update beta  ###
    ###################### 
    # update beta (intercept only model)
    
    pro_beta1=rmvnorm(1,beta1,0.000002*diag(ncol(x1)))
    MHrate1=sum(-E1*exp(pro_beta1%*%t(x1)+theta[u][1:n])+y1*(pro_beta1%*%t(x1)+theta[u][1:n]))-
      sum(-E1*exp(beta1%*%t(x1)+theta[u][1:n])+y1*(beta1%*%t(x1)+theta[u][1:n]))  
    
    if(runif(1,0,1)<exp(MHrate1)){
      beta1<-pro_beta1
      acceptbeta1=acceptbeta1+1
    } 
    
    pro_beta2=rmvnorm(1,beta2,0.00001*diag(ncol(x2)))
    MHrate2=sum(-E2*exp(pro_beta2%*%t(x2)+theta[u][(n+1):(2*n)])+y2*(pro_beta2%*%t(x2)+theta[u][(n+1):(2*n)]))-
      sum(-E2*exp(beta2%*%t(x2)+theta[u][(n+1):(2*n)])+y2*(beta2%*%t(x2)+theta[u][(n+1):(2*n)]))  
    
    if(runif(1,0,1)<exp(MHrate2)){
      beta2<-pro_beta2
      acceptbeta2=acceptbeta2+1
    } 
    
    pro_beta3=rmvnorm(1,beta3,0.00004*diag(ncol(x3)))
    MHrate3=sum(-E3*exp(pro_beta3%*%t(x3)+theta[u][(2*n+1):(3*n)])+y3*(pro_beta3%*%t(x3)+theta[u][(2*n+1):(3*n)]))-
      sum(-E3*exp(beta3%*%t(x3)+theta[u][(2*n+1):(3*n)])+y3*(beta3%*%t(x3)+theta[u][(2*n+1):(3*n)]))  
    
    if(runif(1,0,1)<exp(MHrate3)){
      beta3<-pro_beta3
      acceptbeta3=acceptbeta3+1
    } 
    
    pro_beta4=rmvnorm(1,beta4,0.000001*diag(ncol(x4)))
    MHrate4=sum(-E4*exp(pro_beta4%*%t(x4)+theta[u][(3*n+1):(4*n)])+y4*(pro_beta4%*%t(x4)+theta[u][(3*n+1):(4*n)]))-
      sum(-E4*exp(beta4%*%t(x4)+theta[u][(3*n+1):(4*n)])+y4*(beta4%*%t(x4)+theta[u][(3*n+1):(4*n)]))  
    
    if(runif(1,0,1)<exp(MHrate4)){
      beta4<-pro_beta4
      acceptbeta4=acceptbeta4+1
    } 
    
    beta <- c(beta1, beta2, beta3, beta4)
    
    
    #########################
    ###   update theta    ###
    ######################### 
    
    u1 = u[1:n]
    u2 = u[(n+1):(2*n)]
    u3 = u[(2*n+1):(3*n)]
    u4 = u[(3*n+1):(4*n)]
    
    
    
    for (j in 1:n.atoms){
      
      count[j]=sum(u==j)
      
      if (count[j]==0){
        theta[j]=rnorm(1,0,sqrt(1/taus))
      }else{
        pro_theta=rnorm(1,theta[j],0.06)
        
        MHrate<-sum(-(E[u==j])*exp(X[u==j,] %*% beta+pro_theta)+(y[u==j])*(X[u==j,] %*% beta+pro_theta))-taus/2*pro_theta^2-
          sum(-(E[u==j])*exp(X[u==j,] %*% beta+theta[j])+(y[u==j])*(X[u==j,] %*% beta+theta[j]))+taus/2*theta[j]^2
        if(runif(1,0,1)<exp(MHrate)){
          theta[j]<-pro_theta
          accepttheta=accepttheta+1
        } 
      }
    }
    
    
    
    ######################
    ###   update r     ###
    ###################### 
    
    for (k in 1:nq){
      
      pro_r=r;pro_Fr=F_r;pro_u=u
      pro_r[k]=rnorm(1,r[k],1.8)
      pro_Fr[k]=pnorm(pro_r[k],0,sqrt(Vr[k,k]))			
      pro_u[k]=makeu(pro_Fr[k],probs)
      
      MH=dmvnorm(pro_r,mean=rep(0, nq),sigma=Vr,log=T)+dpois(y[k],E[k]*exp(as.numeric(X[k,]%*%beta)+theta[pro_u[k]]),log=T)-
        dmvnorm(r,mean=rep(0, nq),sigma=Vr,log=T)-dpois(y[k],E[k]*exp(as.numeric(X[k,]%*%beta)+theta[u[k]]),log=T) 
      if(runif(1,0,1)<exp(MH)){
        r[k]=pro_r[k]
        F_r[k]=pro_Fr[k]
        u[k]=pro_u[k]
        acceptr=acceptr+1
      }
    }
    
    
    ######################
    ###   update   v   ###
    ###################### 
    
    
    #pro_v=v
    for (j in 1:n.atoms){
      #0.9
      pro_v=v
      pro_v[j]<-rnorm(1,v[j],0.04)
      while (pro_v[j]<0 || pro_v[j]>1) {pro_v[j]<-rnorm(1,v[j],0.04)}
      #if(pro_v[j]>0 & pro_v[j]<1){
      pro_probs=makeprobs(pro_v)
      pro_u=makeu(F_r,pro_probs)
      
      MH=log(dbeta(pro_v[j],1,alpha))+sum(dpois(y,E*exp(X%*%beta+theta[pro_u]),log=T))-
        log(dbeta(v[j],1,alpha))-sum(dpois(y,E*exp(X%*%beta+theta[u]),log=T))
      
      
      if(runif(1,0,1)<exp(MH)){
        v[j]=pro_v[j]
        probs=pro_probs
        u=pro_u
        acceptv=acceptv+1	
      }
      # }
    }
    
    
    ######################
    ###   update taus  ###
    ###################### 
    
    
    taus=rgamma(1,shape=1/2*n.atoms+c,rate=sum(theta^2)/2+d2)
    
    
    ######################
    ###   update rho  ###
    ###################### 
    
    
    pro_eta1 = rnorm(1, eta[1], 0.5)
    pro_eta2 = rnorm(1, eta[2], 0.5)
    pro_eta3 = rnorm(1, eta[3], 0.5)
    pro_eta4 = rnorm(1, eta[4], 0.5)
    pro_eta = c(pro_eta1, pro_eta2, pro_eta3, pro_eta4)
    pro_rho = exp(pro_eta)/(1+exp(pro_eta))
    pro_Q = Dinv_new(Rho = pro_rho, n, cn, ns, udnei, q)
    pro_invQ1 = solve(pro_Q[[1]])
    pro_invQ2 = solve(pro_Q[[2]])
    pro_invQ3 = solve(pro_Q[[3]])
    pro_invQ4 = solve(pro_Q[[4]])
    
    kprod = kronecker(A, diag(n))
    pro_invQ = as.matrix(bdiag(bdiag(pro_invQ1, pro_invQ2), bdiag(pro_invQ3, pro_invQ4)))
    pro_Vr = kprod %*% pro_invQ %*% t(kprod)
    
    MH = dmvnorm(r,mean=rep(0, nq),sigma=as.matrix(forceSymmetric(pro_Vr)),log=T) + 
      log(pro_rho[1]) + log(1-pro_rho[1]) + log(pro_rho[2]) + log(1-pro_rho[2]) +
      log(pro_rho[3]) + log(1-pro_rho[3]) + log(pro_rho[4]) + log(1-pro_rho[4]) - 
      dmvnorm(r,mean=rep(0, nq),sigma=Vr,log=T) - log(rho[1]) - log(1-rho[1]) - log(rho[2]) - log(1-rho[2]) -
      log(rho[3]) - log(1-rho[3]) - log(rho[4]) - log(1-rho[4])
    
    if(runif(1,0,1)<exp(MH)){
      eta = pro_eta
      rho = pro_rho
      Vr = as.matrix(forceSymmetric(pro_Vr))
      acceptrho=acceptrho+1	
    }
    
    ######################
    ###   update A  ###
    ###################### 
    #0.06
    pro_A = matrix(0, q, q)
    for(i in 1:q){
      for(j in 1:i){
        if(j == i){
          pro_A[i, j] = exp(rnorm(1, log(A[i, j]), 0.04))
          #pro_A[i, j] = rlnorm(1, log(A[i, j]), 0.1)
        }else{
          pro_A[i, j] = rnorm(1, A[i, j], 0.03)
        }
      }
    }
    
    Q = Dinv_new(Rho = rho, n, cn, ns, udnei, q=4)
    invQ1 = solve(Q[[1]])
    invQ2 = solve(Q[[2]])
    invQ3 = solve(Q[[3]])
    invQ4 = solve(Q[[4]])
    Sigma =  A %*% t(A)
    
    pro_kprod = kronecker(pro_A, diag(n))
    invQ = as.matrix(bdiag(bdiag(invQ1, invQ2), bdiag(invQ3, invQ4)))
    pro_Vr = pro_kprod %*% as.matrix(forceSymmetric(invQ)) %*% t(pro_kprod)
    pro_Sigma =  pro_A %*% t(pro_A)
    
    lpA = -(nu+4)/2*logdet(Sigma) - 1/2*sum(diag(nu*R%*%solve(Sigma))) + log(Jacob_A(A)) + sum(log(diag(A)))
    pro_lpA = -(nu+4)/2*logdet(pro_Sigma) - 1/2*sum(diag(nu*R%*%solve(pro_Sigma))) + log(Jacob_A(pro_A)) + sum(log(diag(pro_A)))
    MH = dmvnorm(r,mean=rep(0, nq),sigma=as.matrix(forceSymmetric(pro_Vr)),log=T) + pro_lpA - 
      dmvnorm(r,mean=rep(0, nq),sigma=Vr,log=T) - lpA
    
    if(runif(1,0,1)<exp(MH)){
      A = pro_A
      Vr = as.matrix(forceSymmetric(pro_Vr))
      acceptA=acceptA+1	
    }
    
    #kprod = kronecker(A, diag(n))
    #Vr = as.matrix(forceSymmetric(kprod %*% invQ %*% t(kprod)))
    
    ######################
    ### record samples ###
    ###################### 
    
    keeptheta[iter,] = theta
    keepphi[iter,] = theta[u]
    
    keepA[iter,] = as.vector(A)[-c(5,9,10,13:15)]
    
    keepbeta[iter,]=beta
    keeptaus[iter]=taus
    keeprho1[iter]=rho[1]
    keeprho2[iter]=rho[2]
    keeprho3[iter]=rho[3]
    keeprho4[iter]=rho[4]
    keepu[iter,]=u
    keepv[iter,]=v
    keepr[iter,]=r
    keepFr[iter,]=F_r
    
    #	cat("iteration = ", i, "acceptance rate of r = ", acceptr/(n*i),"acceptance rate of v = ",acceptv/(n.atoms*i), "\n")
    
  }
  
  list(beta=keepbeta[burn:runs,], A = keepA[burn:runs,],phi=keepphi[burn:runs,],theta=keeptheta[burn:runs,],u=keepu[burn:runs,],
       v=keepv[burn:runs,],r=keepr[burn:runs,],Fr=keepFr[burn:runs,], taus=keeptaus[burn:runs],
       rho1=keeprho1[burn:runs], rho2=keeprho2[burn:runs], rho3=keeprho3[burn:runs], rho4=keeprho4[burn:runs])
}
set.seed(123)
mcmc_samples=ARDP_joint_diff_smoke(y=Y, x1=as.matrix(X1[,1:2]), x2=as.matrix(X2[,1:2]), x3 = as.matrix(X3[,1:2]), x4 = as.matrix(X4[,1:2]), 
                             X = X, E=E, Minc, alpha=1, q = 4, n.atoms=15, runs=30000, burn=20000)

sample.mcmc = cbind(mcmc_samples$taus, mcmc_samples$rho1, mcmc_samples$rho2, mcmc_samples$rho3, mcmc_samples$rho4)
```

##### 4.1.1 Coefficients (covariates slopes) estimates in Table 5
```{r}
#Summary function
mysummary = function(invector) {
  c(mean(invector), quantile(invector, .025), quantile(invector,.975))
}
#Coefficients
cov_effect = round(t(apply(mcmc_samples$beta, 2, mysummary)),3)
beta_est = paste(cov_effect[,1], " (", cov_effect[,2], ", ", cov_effect[,3], ")", sep="")
beta_table = data.frame(t(matrix(beta_est, ncol = 2, byrow = TRUE)))
colnames(beta_table) = c("Lung", "Esophageal", "Larynx", "Colorectal")
rownames(beta_table) = c("Intercept", "Smoking")
kable(beta_table)

# Estimates for $rho_d$
round(mysummary(mcmc_samples$rho1), 3)
round(mysummary(mcmc_samples$rho2), 3)
round(mysummary(mcmc_samples$rho3), 3)
round(mysummary(mcmc_samples$rho4), 3)
```

##### 4.1.2 Difference boundaries for each cancer individually

Figure 10 (a): Difference boundaries (highlighted in red) detected by MDAGAR after accounting for smoking for four cancers
```{r, fig.height=7, fig.width=10}
phis <- mcmc_samples$phi
phis1 <- phis[,1:58][,order(final_perm)]
phis2 <- phis[,59:116][,order(final_perm)]
phis3 <- phis[,117:174][,order(final_perm)]
phis4 <- phis[,175:232][,order(final_perm)]
phis_origin <- cbind(phis1, phis2, phis3, phis4)

vij_samples1 <- t(apply(phis1, 1, function(x){
  diff_b1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]]){
      diff_b1 <- c(diff_b1, 1)
    }else{
      diff_b1 <- c(diff_b1, 0)
    }
  }
  return(diff_b1)
}))



vij_samples2 <- t(apply(phis2, 1, function(x){
  diff_b1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]]){
      diff_b1 <- c(diff_b1, 1)
    }else{
      diff_b1 <- c(diff_b1, 0)
    }
  }
  return(diff_b1)
}))

vij_samples3 <- t(apply(phis3, 1, function(x){
  diff_b1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]]){
      diff_b1 <- c(diff_b1, 1)
    }else{
      diff_b1 <- c(diff_b1, 0)
    }
  }
  return(diff_b1)
}))



vij_samples4 <- t(apply(phis4, 1, function(x){
  diff_b1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]]){
      diff_b1 <- c(diff_b1, 1)
    }else{
      diff_b1 <- c(diff_b1, 0)
    }
  }
  return(diff_b1)
}))

#probabilities
pvij1 <- apply(vij_samples1, 2, mean)
pvij2 <- apply(vij_samples2, 2, mean)
pvij3 <- apply(vij_samples3, 2, mean)
pvij4 <- apply(vij_samples4, 2, mean)

names(pvij1) <- neighbor_name
names(pvij2) <- neighbor_name
names(pvij3) <- neighbor_name
names(pvij4) <- neighbor_name

T_edge1 = seq(0, 135, 1)[-1]
threshold1 = sort(pvij1, decreasing = TRUE)[T_edge1]
threshold2 = sort(pvij2, decreasing = TRUE)[T_edge1]
threshold3 = sort(pvij3, decreasing = TRUE)[T_edge1]
threshold4 = sort(pvij4, decreasing = TRUE)[T_edge1]
FDR_est1 <- rep(0, length(T_edge1))
FDR_est2 <- rep(0, length(T_edge1))
FDR_est3 <- rep(0, length(T_edge1))
FDR_est4 <- rep(0, length(T_edge1))


for(i in 1:length(threshold1)){
  th1 <- threshold1[i]
  est_diff1 <- as.numeric(pvij1 >= th1)
  FDR_est1[i] <- sum((1-pvij1) * est_diff1)  / sum(est_diff1)
  
  th2 <- threshold2[i]
  est_diff2 <- as.numeric(pvij2 >= th2)
  FDR_est2[i] <- sum((1-pvij2) * est_diff2)  / sum(est_diff2)
  
  th3 <- threshold3[i]
  est_diff3 <- as.numeric(pvij3 >= th3)
  FDR_est3[i] <- sum((1-pvij3) * est_diff3)  / sum(est_diff3)
  
  th4 <- threshold4[i]
  est_diff4 <- as.numeric(pvij4 >= th4)
  FDR_est4[i] <- sum((1-pvij4) * est_diff4)  / sum(est_diff4)
}

# Use the same threshold as MDAGAR and identify difference boundary for each cancer individually
alpha_n = 0.05
T1 = sum(FDR_est1<=alpha_n)
T2 = sum(FDR_est2<=alpha_n)
T3 = sum(FDR_est3<=alpha_n)
T4 = sum(FDR_est4<=alpha_n)

est_diff1 <- as.numeric(pvij1 >= threshold1[T1])
est_diff2 <- as.numeric(pvij2 >= threshold2[T2])
est_diff3 <- as.numeric(pvij3 >= threshold3[T3])
est_diff4 <- as.numeric(pvij4 >= threshold4[T4])

ch_edge1 <- pvij1[which(est_diff1 == 1)]
edge_plot1 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group, fill = lung_SIR),
               #color = ifelse(ca.poly.df$zero_pos == 1, "white", "black"),
               #alpha = ifelse(ca.poly.df$zero_pos == 1, 0.5, 1),
               color = "black",
               alpha = 0.6
               #fill = "white"
               #linetype = ifelse(ca.poly.df$zero_pos == 1, "dotted", "solid")
  ) +
  scale_fill_stepsn(colors = color,
                    breaks = breaks1)

for(i in which(est_diff1 == 1)){
  edge_plot1 = edge_plot1 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                      size = ((pvij1[i]- min(ch_edge1))/(max(ch_edge1) - min(ch_edge1)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=16, face="bold",hjust = 0.5),
          legend.title = element_blank()) +
    ggtitle(paste("Lung (T = ", T1, ")", sep=""))
}


ch_edge2 <- pvij2[which(est_diff2 == 1)]
edge_plot2 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group, fill = esophagus_SIR),
               #color = ifelse(ca.poly.df$zero_pos == 1, "white", "black"),
               #alpha = ifelse(ca.poly.df$zero_pos == 1, 0.5, 1),
               color = "black",
               alpha = 0.6
               #fill = "white"
               #linetype = ifelse(ca.poly.df$zero_pos == 1, "dotted", "solid")
  ) +
  scale_fill_stepsn(colors = color,
                    breaks = breaks2)
for(i in which(est_diff2 == 1)){
  edge_plot2 = edge_plot2 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                      size = ((pvij2[i]- min(ch_edge2))/(max(ch_edge2) - min(ch_edge2)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=16, face="bold",hjust = 0.5),
          legend.title = element_blank()) +
    ggtitle(paste("Esophageal (T = ", T2, ")", sep=""))
}

ch_edge3 <- pvij3[which(est_diff3 == 1)]
edge_plot3 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group, fill = larynx_SIR),
               #color = ifelse(ca.poly.df$zero_pos == 1, "white", "black"),
               #alpha = ifelse(ca.poly.df$zero_pos == 1, 0.5, 1),
               color = "black",
               alpha = 0.6
               #fill = "white"
               #linetype = ifelse(ca.poly.df$zero_pos == 1, "dotted", "solid")
  ) +
  scale_fill_stepsn(colors = color,
                    breaks = breaks3)
for(i in which(est_diff3 == 1)){
  edge_plot3 = edge_plot3 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                      size = ((pvij3[i]- min(ch_edge3))/(max(ch_edge3) - min(ch_edge3)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=16, face="bold",hjust = 0.5),
          legend.title = element_blank()) +
    ggtitle(paste("Larynx (T = ", T3, ")", sep=""))
}

ch_edge4 <- pvij4[which(est_diff4 == 1)]
edge_plot4 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group, fill = colrect_SIR),
               #color = ifelse(ca.poly.df$zero_pos == 1, "white", "black"),
               #alpha = ifelse(ca.poly.df$zero_pos == 1, 0.5, 1),
               color = "black",
               alpha = 0.6
               #fill = "white"
               #linetype = ifelse(ca.poly.df$zero_pos == 1, "dotted", "solid")
  ) +
  scale_fill_stepsn(colors = color,
                    breaks = breaks4)
for(i in which(est_diff4 == 1)){
  edge_plot4 = edge_plot4 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                      size = ((pvij4[i]- min(ch_edge4))/(max(ch_edge4) - min(ch_edge4)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=16, face="bold",hjust = 0.5),
          legend.title = element_blank()) +
    ggtitle(paste("Colorectal (T = ", T4, ")", sep=""))
}
ggarrange(edge_plot1, edge_plot2, edge_plot3, edge_plot4, nrow = 2, ncol = 2) 
```

#### 4.2 Account for smoking and unemployed
```{r}
# Construct covariates matrix
X = as.matrix(bdiag(bdiag(X1[,c(1,2,7)], X2[,c(1,2,7)]), bdiag(X3[,c(1,2,7)],X4[,c(1,2,7)])))

ARDP_joint_diff<-function(y=Y, x1=as.matrix(X1[,c(1,2,7)]), x2=as.matrix(X2[,c(1,2,7)]), x3 = as.matrix(X3[,c(1,2,7)]), x4 = as.matrix(X4[,c(1,2,7)]),
                          X = X, E = E, Minc, alpha=1, q = 4, n.atoms=15, runs=10000, burn=1000){
  #y:       data
  #x:       covariates
  #n.atoms: number of atoms in the mixture dist.
  #theta:      the theta's (iid from baseline) in the model
  #alpha:     v~beta(1,alpha)
  #u:       the index indicator of spatial random effect
  #rho:     sptatial autocorrelation parameter in DAGAR
  #Minc:       0-1 adjacency matrix
  #V_r:     covariance matrix of joint MDAGAR
  #Q:       presicion matrix of DAGAR
  #r:       random effects following DAGAR
  #F_r:     Marginal CDF of r
  #taued:   presicion in Baseline of DP for disease d
  #taus:    precision for theta
  
  
  
  
  nq<-length(y)
  n <- nq / q
  p<-ncol(X)
  y1 <- y[1:n]
  y2 <- y[(n+1):(2*n)]
  y3 <- y[(2*n+1):(3*n)]
  y4 <- y[(3*n+1):(4*n)]
  
  E1 <- E[1:n]
  E2 <- E[(n+1):(2*n)]
  E3 <- E[(2*n+1):(3*n)]
  E4 <- E[(3*n+1):(4*n)]
  
  sigmasq_beta = 10000
  keepbeta=matrix(0,runs,p)
  keepphi=matrix(0,runs,nq)
  keeptheta=matrix(0,runs,n.atoms)
  keepA=matrix(0,runs,10)
  keepu=matrix(0,runs,nq)
  keeprho1=keeprho2=keeprho3=keeprho4=keeptaus = rep(0,runs)
  keepv=matrix(0,runs,n.atoms)
  keepr=matrix(0,runs,nq)
  keepFr=matrix(0,runs,nq)
  
  
  #initial values
  
  theta=rep(0,n.atoms)
  
  beta1<-rep(0,ncol(x1))
  beta2<-rep(0,ncol(x2))
  beta3<-rep(0,ncol(x3))
  beta4<-rep(0,ncol(x4))
  beta = c(beta1, beta2, beta3, beta4)
  
  taus = 1
  
  c=2
  d=0.1
  d2=1
  v<-rep(.1,n.atoms)
  probs=makeprobs(v)
  
  #A matrix
  rho=c(0.5, 0.5, 0.5, 0.5)
  eta = log(rho/(1-rho))
  A = matrix(0, q, q)
  for(i in 1:q){
    for(j in 1:i){
      if(j == i){
        A[i, j] = exp(rnorm(1))
      }else{
        A[i, j] = rnorm(1)
      }
    }
  }
  #Sigma = A%*%t(A)
  
  Q = Dinv_new(Rho = rho, n, cn, ns, udnei, q)
  invQ1 = solve(Q[[1]])
  invQ2 = solve(Q[[2]])
  invQ3 = solve(Q[[3]])
  invQ4 = solve(Q[[4]])
  
  #f1 = rmvnorm(1, rep(0, n), invQ1)
  #f2 = rmvnorm(1, rep(0, n), invQ2)
  #f = c(f1, f2)
  kprod = kronecker(A, diag(n))
  invQ = as.matrix(bdiag(bdiag(invQ1, invQ2), bdiag(invQ3, invQ4)))
  #r = kprod %*% f
  
  Vr = as.matrix(forceSymmetric(kprod %*% invQ %*% t(kprod)))
  r = rmvnorm(1, rep(0, nq), Vr)
  
  
  F_r=pnorm(r,0,sqrt(diag(Vr)))
  u=makeu(F_r,probs)
  phi <- theta[u]
  
  nu = 2
  R = 0.1 * diag(q)
  
  
  acceptr=acceptv=acceptrho=acceptA=0
  acceptbeta1 = acceptbeta2=acceptbeta3=acceptbeta4 = 0
  accepttheta = 0
  
  #sumtp<-summu<-summu2<-rep(0,n)
  count<-afterburn<-0
  burn = burn + 1
  #iter = 1
  #runs = 1000
  for(iter in 1:runs){
    
    if(iter %% 100 == 0){
      print(iter)
      print(acceptA/(iter-1))
      print(acceptr/nq/(iter-1))
      print(acceptrho/(iter-1))
      print(acceptv/n.atoms/(iter-1))
      print(accepttheta/n.atoms/(iter-1))
      print(acceptbeta1/(iter-1))
      print(acceptbeta2/(iter-1))
      print(acceptbeta3/(iter-1))
      print(acceptbeta4/(iter-1))
    }
    ######################
    ###   update beta  ###
    ######################
    # update beta (intercept only model)
    
    pro_beta1=rmvnorm(1,beta1,0.0000009*diag(ncol(x1)))
    MHrate1=sum(-E1*exp(pro_beta1%*%t(x1)+theta[u][1:n])+y1*(pro_beta1%*%t(x1)+theta[u][1:n]))-
      sum(-E1*exp(beta1%*%t(x1)+theta[u][1:n])+y1*(beta1%*%t(x1)+theta[u][1:n]))
    
    if(runif(1,0,1)<exp(MHrate1)){
      beta1<-pro_beta1
      acceptbeta1=acceptbeta1+1
    }
    
    pro_beta2=rmvnorm(1,beta2,0.000012*diag(ncol(x2)))
    MHrate2=sum(-E2*exp(pro_beta2%*%t(x2)+theta[u][(n+1):(2*n)])+y2*(pro_beta2%*%t(x2)+theta[u][(n+1):(2*n)]))-
      sum(-E2*exp(beta2%*%t(x2)+theta[u][(n+1):(2*n)])+y2*(beta2%*%t(x2)+theta[u][(n+1):(2*n)]))
    
    if(runif(1,0,1)<exp(MHrate2)){
      beta2<-pro_beta2
      acceptbeta2=acceptbeta2+1
    }
    
    pro_beta3=rmvnorm(1,beta3,0.000022*diag(ncol(x3)))
    MHrate3=sum(-E3*exp(pro_beta3%*%t(x3)+theta[u][(2*n+1):(3*n)])+y3*(pro_beta3%*%t(x3)+theta[u][(2*n+1):(3*n)]))-
      sum(-E3*exp(beta3%*%t(x3)+theta[u][(2*n+1):(3*n)])+y3*(beta3%*%t(x3)+theta[u][(2*n+1):(3*n)]))
    
    if(runif(1,0,1)<exp(MHrate3)){
      beta3<-pro_beta3
      acceptbeta3=acceptbeta3+1
    }
    
    pro_beta4=rmvnorm(1,beta4,0.0000012*diag(ncol(x4)))
    MHrate4=sum(-E4*exp(pro_beta4%*%t(x4)+theta[u][(3*n+1):(4*n)])+y4*(pro_beta4%*%t(x4)+theta[u][(3*n+1):(4*n)]))-
      sum(-E4*exp(beta4%*%t(x4)+theta[u][(3*n+1):(4*n)])+y4*(beta4%*%t(x4)+theta[u][(3*n+1):(4*n)]))
    
    if(runif(1,0,1)<exp(MHrate4)){
      beta4<-pro_beta4
      acceptbeta4=acceptbeta4+1
    }
    
    beta <- c(beta1, beta2, beta3, beta4)
    
    
    #########################
    ###   update theta    ###
    #########################
    
    u1 = u[1:n]
    u2 = u[(n+1):(2*n)]
    u3 = u[(2*n+1):(3*n)]
    u4 = u[(3*n+1):(4*n)]
    
    
    
    for (j in 1:n.atoms){
      
      count[j]=sum(u==j)
      
      if (count[j]==0){
        theta[j]=rnorm(1,0,sqrt(1/taus))
      }else{
        pro_theta=rnorm(1,theta[j],0.06)
        
        MHrate<-sum(-(E[u==j])*exp(X[u==j,] %*% beta+pro_theta)+(y[u==j])*(X[u==j,] %*% beta+pro_theta))-taus/2*pro_theta^2-
          sum(-(E[u==j])*exp(X[u==j,] %*% beta+theta[j])+(y[u==j])*(X[u==j,] %*% beta+theta[j]))+taus/2*theta[j]^2
        if(runif(1,0,1)<exp(MHrate)){
          theta[j]<-pro_theta
          accepttheta=accepttheta+1
        }
      }
    }
    
    
    
    ######################
    ###   update r     ###
    ######################
    
    #pro_r=t(rmvnorm(1,r,sigma=covcar2,method="chol"))
    
    
    #tauvec = c(rep(taue1, n), rep(taue2, n), rep(taue3, n), rep(taue4, n))
    
    #pro_r = rmvnorm(1, r, 0.01*diag(nq))
    
    for (k in 1:nq){
      
      pro_r=r;pro_Fr=F_r;pro_u=u
      pro_r[k]=rnorm(1,r[k],1.8)
      pro_Fr[k]=pnorm(pro_r[k],0,sqrt(Vr[k,k]))
      pro_u[k]=makeu(pro_Fr[k],probs)
      
      MH=dmvnorm(pro_r,mean=rep(0, nq),sigma=Vr,log=T)+dpois(y[k],E[k]*exp(as.numeric(X[k,]%*%beta)+theta[pro_u[k]]),log=T)-
        dmvnorm(r,mean=rep(0, nq),sigma=Vr,log=T)-dpois(y[k],E[k]*exp(as.numeric(X[k,]%*%beta)+theta[u[k]]),log=T)
      #MH=dmvnorm(pro_r,mean=rep(0, nq),sigma=Vr,log=T)+dmvnorm(t(as.matrix(y1)),mean=as.vector(x1%*%t(beta1))+theta[pro_u[1:n]],sigma=1/taue1*diag(n),log=T)+
      # dmvnorm(t(as.matrix(y2)),mean=as.vector(x2%*%t(beta2))+theta[pro_u[(n+1):(2*n)]],sigma=1/taue2*diag(n),log=T)-
      #dmvnorm(r,mean=rep(0, nq),sigma=Vr,log=T)-dmvnorm(t(as.matrix(y1)),mean=as.vector(x1%*%t(beta1))+theta[u[1:n]],sigma=1/taue1*diag(n),log=T)-
      #dmvnorm(t(as.matrix(y2)),mean=as.vector(x2%*%t(beta2))+theta[u[(n+1):(2*n)]],sigma=1/taue2*diag(n),log=T)
      if(runif(1,0,1)<exp(MH)){
        r[k]=pro_r[k]
        F_r[k]=pro_Fr[k]
        u[k]=pro_u[k]
        acceptr=acceptr+1
      }
    }
    
    
    ######################
    ###   update   v   ###
    ######################
    
    
    #pro_v=v
    for (j in 1:n.atoms){
      #0.9
      pro_v=v
      pro_v[j]<-rnorm(1,v[j],0.05)
      while (pro_v[j]<0 || pro_v[j]>1) {pro_v[j]<-rnorm(1,v[j],0.05)}
      #if(pro_v[j]>0 & pro_v[j]<1){
      pro_probs=makeprobs(pro_v)
      pro_u=makeu(F_r,pro_probs)
      
      MH=log(dbeta(pro_v[j],1,alpha))+sum(dpois(y,E*exp(X%*%beta+theta[pro_u]),log=T))-
        log(dbeta(v[j],1,alpha))-sum(dpois(y,E*exp(X%*%beta+theta[u]),log=T))
      
      
      if(runif(1,0,1)<exp(MH)){
        v[j]=pro_v[j]
        probs=pro_probs
        u=pro_u
        acceptv=acceptv+1
      }
      # }
    }
    
    
    ######################
    ###   update taus  ###
    ######################
    
    
    taus=rgamma(1,shape=1/2*n.atoms+c,rate=sum(theta^2)/2+d2)
    
    
    ######################
    ###   update rho  ###
    ######################
    
    
    pro_eta1 = rnorm(1, eta[1], 0.5)
    pro_eta2 = rnorm(1, eta[2], 0.5)
    pro_eta3 = rnorm(1, eta[3], 0.5)
    pro_eta4 = rnorm(1, eta[4], 0.5)
    pro_eta = c(pro_eta1, pro_eta2, pro_eta3, pro_eta4)
    pro_rho = exp(pro_eta)/(1+exp(pro_eta))
    pro_Q = Dinv_new(Rho = pro_rho, n, cn, ns, udnei, q)
    pro_invQ1 = solve(pro_Q[[1]])
    pro_invQ2 = solve(pro_Q[[2]])
    pro_invQ3 = solve(pro_Q[[3]])
    pro_invQ4 = solve(pro_Q[[4]])
    
    kprod = kronecker(A, diag(n))
    pro_invQ = as.matrix(bdiag(bdiag(pro_invQ1, pro_invQ2), bdiag(pro_invQ3, pro_invQ4)))
    pro_Vr = kprod %*% pro_invQ %*% t(kprod)
    
    MH = dmvnorm(r,mean=rep(0, nq),sigma=as.matrix(forceSymmetric(pro_Vr)),log=T) +
      log(pro_rho[1]) + log(1-pro_rho[1]) + log(pro_rho[2]) + log(1-pro_rho[2]) +
      log(pro_rho[3]) + log(1-pro_rho[3]) + log(pro_rho[4]) + log(1-pro_rho[4]) -
      dmvnorm(r,mean=rep(0, nq),sigma=Vr,log=T) - log(rho[1]) - log(1-rho[1]) - log(rho[2]) - log(1-rho[2]) -
      log(rho[3]) - log(1-rho[3]) - log(rho[4]) - log(1-rho[4])
    
    if(runif(1,0,1)<exp(MH)){
      eta = pro_eta
      rho = pro_rho
      Vr = as.matrix(forceSymmetric(pro_Vr))
      acceptrho=acceptrho+1
    }
    
    ######################
    ###   update A  ###
    ######################
    #0.06
    pro_A = matrix(0, q, q)
    for(i in 1:q){
      for(j in 1:i){
        if(j == i){
          pro_A[i, j] = exp(rnorm(1, log(A[i, j]), 0.03))
          #pro_A[i, j] = rlnorm(1, log(A[i, j]), 0.1)
        }else{
          pro_A[i, j] = rnorm(1, A[i, j], 0.03)
        }
      }
    }
    
    Q = Dinv_new(Rho = rho, n, cn, ns, udnei, q=4)
    invQ1 = solve(Q[[1]])
    invQ2 = solve(Q[[2]])
    invQ3 = solve(Q[[3]])
    invQ4 = solve(Q[[4]])
    Sigma =  A %*% t(A)
    
    pro_kprod = kronecker(pro_A, diag(n))
    invQ = as.matrix(bdiag(bdiag(invQ1, invQ2), bdiag(invQ3, invQ4)))
    pro_Vr = pro_kprod %*% as.matrix(forceSymmetric(invQ)) %*% t(pro_kprod)
    pro_Sigma =  pro_A %*% t(pro_A)
    
    lpA = -(nu+4)/2*logdet(Sigma) - 1/2*sum(diag(nu*R%*%solve(Sigma))) + log(Jacob_A(A)) + sum(log(diag(A)))
    pro_lpA = -(nu+4)/2*logdet(pro_Sigma) - 1/2*sum(diag(nu*R%*%solve(pro_Sigma))) + log(Jacob_A(pro_A)) + sum(log(diag(pro_A)))
    MH = dmvnorm(r,mean=rep(0, nq),sigma=as.matrix(forceSymmetric(pro_Vr)),log=T) + pro_lpA -
      dmvnorm(r,mean=rep(0, nq),sigma=Vr,log=T) - lpA
    
    if(runif(1,0,1)<exp(MH)){
      A = pro_A
      Vr = as.matrix(forceSymmetric(pro_Vr))
      acceptA=acceptA+1
    }
    
    #kprod = kronecker(A, diag(n))
    #Vr = as.matrix(forceSymmetric(kprod %*% invQ %*% t(kprod)))
    
    ######################
    ### record samples ###
    ######################
    
    keeptheta[iter,] = theta
    keepphi[iter,] = theta[u]
    
    keepA[iter,] = as.vector(A)[-c(5,9,10,13:15)]
    
    keepbeta[iter,]=beta
    keeptaus[iter]=taus
    keeprho1[iter]=rho[1]
    keeprho2[iter]=rho[2]
    keeprho3[iter]=rho[3]
    keeprho4[iter]=rho[4]
    keepu[iter,]=u
    keepv[iter,]=v
    keepr[iter,]=r
    keepFr[iter,]=F_r
    
    #    cat("iteration = ", i, "acceptance rate of r = ", acceptr/(n*i),"acc1eptance rate of v = ",acceptv/(n.atoms*i), "\n")
    
  }
  
  list(beta=keepbeta[burn:runs,], A = keepA[burn:runs,],phi=keepphi[burn:runs,],theta=keeptheta[burn:runs,],u=keepu[burn:runs,],
       v=keepv[burn:runs,],r=keepr[burn:runs,],Fr=keepFr[burn:runs,], taus=keeptaus[burn:runs],
       rho1=keeprho1[burn:runs], rho2=keeprho2[burn:runs], rho3=keeprho3[burn:runs], rho4=keeprho4[burn:runs])
}
set.seed(123)
mcmc_samples=ARDP_joint_diff(y=Y, x1=as.matrix(X1[,c(1,2,7)]), x2=as.matrix(X2[,c(1,2,7)]), x3 = as.matrix(X3[,c(1,2,7)]), x4 = as.matrix(X4[,c(1,2,7)]),
                             X = X, E=E, Minc, alpha=1, q = 4, n.atoms=15, runs=30000, burn=20000)


sample.mcmc = cbind(mcmc_samples$taus, mcmc_samples$rho1, mcmc_samples$rho2, mcmc_samples$rho3, mcmc_samples$rho4)
```

##### 4.2.1 Coefficients (covariates slopes) estimates in Table 5
```{r}
#Summary function
mysummary = function(invector) {
  c(mean(invector), quantile(invector, .025), quantile(invector,.975))
}
#Coefficients
cov_effect = round(t(apply(mcmc_samples$beta, 2, mysummary)),3)
beta_est = paste(cov_effect[,1], " (", cov_effect[,2], ", ", cov_effect[,3], ")", sep="")
beta_table = data.frame(t(matrix(beta_est, ncol = 3, byrow = TRUE)))
colnames(beta_table) = c("Lung", "Esophageal", "Larynx", "Colorectal")
rownames(beta_table) = c("Intercept", "Smoking", "Age")
kable(beta_table)

# Estimates for $rho_d$
round(mysummary(mcmc_samples$rho1), 3)
round(mysummary(mcmc_samples$rho2), 3)
round(mysummary(mcmc_samples$rho3), 3)
round(mysummary(mcmc_samples$rho4), 3)
```

##### 4.2.2 Difference boundaries for each cancer individually

Figure 10 (b): Difference boundaries (highlighted in red) detected by MDAGAR after accounting for smoking and age for four cancers
```{r, fig.height=7, fig.width=10}
phis <- mcmc_samples$phi
phis1 <- phis[,1:58][,order(final_perm)]
phis2 <- phis[,59:116][,order(final_perm)]
phis3 <- phis[,117:174][,order(final_perm)]
phis4 <- phis[,175:232][,order(final_perm)]
phis_origin <- cbind(phis1, phis2, phis3, phis4)

vij_samples1 <- t(apply(phis1, 1, function(x){
  diff_b1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]]){
      diff_b1 <- c(diff_b1, 1)
    }else{
      diff_b1 <- c(diff_b1, 0)
    }
  }
  return(diff_b1)
}))



vij_samples2 <- t(apply(phis2, 1, function(x){
  diff_b1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]]){
      diff_b1 <- c(diff_b1, 1)
    }else{
      diff_b1 <- c(diff_b1, 0)
    }
  }
  return(diff_b1)
}))

vij_samples3 <- t(apply(phis3, 1, function(x){
  diff_b1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]]){
      diff_b1 <- c(diff_b1, 1)
    }else{
      diff_b1 <- c(diff_b1, 0)
    }
  }
  return(diff_b1)
}))



vij_samples4 <- t(apply(phis4, 1, function(x){
  diff_b1 <- NULL
  for(i in 1:nrow(neighbor_list0)){
    if(x[neighbor_list0[i,1]] != x[neighbor_list0[i,2]]){
      diff_b1 <- c(diff_b1, 1)
    }else{
      diff_b1 <- c(diff_b1, 0)
    }
  }
  return(diff_b1)
}))

#probabilities
pvij1 <- apply(vij_samples1, 2, mean)
pvij2 <- apply(vij_samples2, 2, mean)
pvij3 <- apply(vij_samples3, 2, mean)
pvij4 <- apply(vij_samples4, 2, mean)

names(pvij1) <- neighbor_name
names(pvij2) <- neighbor_name
names(pvij3) <- neighbor_name
names(pvij4) <- neighbor_name

T_edge1 = seq(0, 135, 1)[-1]
threshold1 = sort(pvij1, decreasing = TRUE)[T_edge1]
threshold2 = sort(pvij2, decreasing = TRUE)[T_edge1]
threshold3 = sort(pvij3, decreasing = TRUE)[T_edge1]
threshold4 = sort(pvij4, decreasing = TRUE)[T_edge1]
FDR_est1 <- rep(0, length(T_edge1))
FDR_est2 <- rep(0, length(T_edge1))
FDR_est3 <- rep(0, length(T_edge1))
FDR_est4 <- rep(0, length(T_edge1))


for(i in 1:length(threshold1)){
  th1 <- threshold1[i]
  est_diff1 <- as.numeric(pvij1 >= th1)
  FDR_est1[i] <- sum((1-pvij1) * est_diff1)  / sum(est_diff1)
  
  th2 <- threshold2[i]
  est_diff2 <- as.numeric(pvij2 >= th2)
  FDR_est2[i] <- sum((1-pvij2) * est_diff2)  / sum(est_diff2)
  
  th3 <- threshold3[i]
  est_diff3 <- as.numeric(pvij3 >= th3)
  FDR_est3[i] <- sum((1-pvij3) * est_diff3)  / sum(est_diff3)
  
  th4 <- threshold4[i]
  est_diff4 <- as.numeric(pvij4 >= th4)
  FDR_est4[i] <- sum((1-pvij4) * est_diff4)  / sum(est_diff4)
}

# Use the same threshold as MDAGAR and identify difference boundary for each cancer individually
alpha_n = 0.05
T1 = sum(FDR_est1<=alpha_n)
T2 = sum(FDR_est2<=alpha_n)
T3 = sum(FDR_est3<=alpha_n)
T4 = sum(FDR_est4<=alpha_n)

est_diff1 <- as.numeric(pvij1 >= threshold1[T1])
est_diff2 <- as.numeric(pvij2 >= threshold2[T2])
est_diff3 <- as.numeric(pvij3 >= threshold3[T3])
est_diff4 <- as.numeric(pvij4 >= threshold4[T4])

ch_edge1 <- pvij1[which(est_diff1 == 1)]
edge_plot1 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group, fill = lung_SIR),
               #color = ifelse(ca.poly.df$zero_pos == 1, "white", "black"),
               #alpha = ifelse(ca.poly.df$zero_pos == 1, 0.5, 1),
               color = "black",
               alpha = 0.6
               #fill = "white"
               #linetype = ifelse(ca.poly.df$zero_pos == 1, "dotted", "solid")
  ) +
  scale_fill_stepsn(colors = color,
                    breaks = breaks1)

for(i in which(est_diff1 == 1)){
  edge_plot1 = edge_plot1 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                      size = ((pvij1[i]- min(ch_edge1))/(max(ch_edge1) - min(ch_edge1)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=16, face="bold",hjust = 0.5),
          legend.title = element_blank()) +
    ggtitle(paste("Lung (T = ", T1, ")", sep=""))
}

ch_edge2 <- pvij2[which(est_diff2 == 1)]
edge_plot2 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group, fill = esophagus_SIR),
               #color = ifelse(ca.poly.df$zero_pos == 1, "white", "black"),
               #alpha = ifelse(ca.poly.df$zero_pos == 1, 0.5, 1),
               color = "black",
               alpha = 0.6
               #fill = "white"
               #linetype = ifelse(ca.poly.df$zero_pos == 1, "dotted", "solid")
  ) +
  scale_fill_stepsn(colors = color,
                    breaks = breaks2)
for(i in which(est_diff2 == 1)){
  edge_plot2 = edge_plot2 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                      size = ((pvij2[i]- min(ch_edge2))/(max(ch_edge2) - min(ch_edge2)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=16, face="bold",hjust = 0.5),
          legend.title = element_blank()) +
    ggtitle(paste("Esophageal (T = ", T2, ")", sep=""))
}

ch_edge3 <- pvij3[which(est_diff3 == 1)]
edge_plot3 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group, fill = larynx_SIR),
               #color = ifelse(ca.poly.df$zero_pos == 1, "white", "black"),
               #alpha = ifelse(ca.poly.df$zero_pos == 1, 0.5, 1),
               color = "black",
               alpha = 0.6
               #fill = "white"
               #linetype = ifelse(ca.poly.df$zero_pos == 1, "dotted", "solid")
  ) +
  scale_fill_stepsn(colors = color,
                    breaks = breaks3)
for(i in which(est_diff3 == 1)){
  edge_plot3 = edge_plot3 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                      size = ((pvij3[i]- min(ch_edge3))/(max(ch_edge3) - min(ch_edge3)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=16, face="bold",hjust = 0.5),
          legend.title = element_blank()) +
    ggtitle(paste("Larynx (T = ", T3, ")", sep=""))
}

ch_edge4 <- pvij4[which(est_diff4 == 1)]
edge_plot4 <- ggplot() +
  geom_polygon(data = ca.poly.df,  aes(long, lat, group = group, fill = colrect_SIR),
               #color = ifelse(ca.poly.df$zero_pos == 1, "white", "black"),
               #alpha = ifelse(ca.poly.df$zero_pos == 1, 0.5, 1),
               color = "black",
               alpha = 0.6
               #fill = "white"
               #linetype = ifelse(ca.poly.df$zero_pos == 1, "dotted", "solid")
  ) +
  scale_fill_stepsn(colors = color,
                    breaks = breaks4)
for(i in which(est_diff4 == 1)){
  edge_plot4 = edge_plot4 + geom_path(aes_string(x = path[[i]][,1], y = path[[i]][,2]), color = "red",
                                      size = ((pvij4[i]- min(ch_edge4))/(max(ch_edge4) - min(ch_edge4)))*0.5 + 0.5) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(size=16, face="bold",hjust = 0.5),
          legend.title = element_blank()) +
    ggtitle(paste("Colorectal (T = ", T4, ")", sep=""))
}


ggarrange(edge_plot1, edge_plot2, edge_plot3, edge_plot4, nrow = 2, ncol = 2) 
```



